---
title: "Formulating experimental feeds"
author: Tormey Reimer
date: today
---

# Introduction

The purpose of this markdown is to formulate all the feeds that will be used in the analysis. This needs to be done before 

```{r setup, include=F, cache=F}
#| code-summary: R setup

library(tidyverse)
library(arrow)
library(sf)
library(terra)
library(furrr)
library(future)
library(tictoc)
library(fs)
library(conflicted)
library(readxl)
library(units)
library(qs)
library(here)
conflicts_prefer(dplyr::filter(), dplyr::select(), .quiet = T)

here("src") %>% 
  list.files(pattern = "\\.R$", full.names = TRUE) %>% 
  walk(source)
```

# Import ingredients

There is no check or correction here to ensure that protein + lipid + ash + carb = 1 in the incoming ingredient data. This needs to be checked manually prior to import. 

``` {r import-ingredients}
#| code-summary: Get ingredient details (PLC composition, digestibility)

ingreds <- file.path(bigdata_path, "all_ingredients.xlsx") %>% 
  read_excel(sheet = "all_ingredients") %>% 
  mutate(ingredient = as.factor(ingredient))

ingred_nms <- levels(ingreds$ingredient)
```

# Import feeds

Unlike ingredients, the feeds do not need to be checked for sum = 1 prior to import. That check is done here.

``` {r import-feeds}
#| code-summary: Feed compositions

feed_inputs <- file.path(input_feed_profile_path, "all_feeds.xlsx") %>% 
  read_excel(sheet = "all_feeds")
  
feed_inputs <- feed_inputs %>% 
  pivot_longer(names_to = "feed", values_to = "proportion", cols = !contains(c("ingredient", "category"))) %>% 
  mutate(
    feed = as.factor(feed),
    ingredient = as.factor(ingredient),
    proportion = case_when(is.na(proportion) ~ 0, T ~ proportion)
  ) %>% 
  left_join(ingreds, by = "ingredient")
```

Check that ingredient proportions sum to 1.

```{r check-proportions}
#| code-summary: Composition proportions

feed_inputs <- feed_inputs %>% 
  filter(!is.na(proportion) & proportion != 0) %>% 
  group_by(feed) %>% 
  mutate(
    sum = sum(proportion),
    proportion = proportion/sum
  ) %>% 
  ungroup() %>% 
  select(-sum, -ash)
```

# Prepare for pipeline

The following chunk allows as many ingredients and feeds to be added to the incoming data as needed, it will still be formatted into a form that the `targets` pipelines can use. 

```{r formulate-feeds}
#| code-summary: Format feeds as lists to give to targets

feed_types <- levels(feed_inputs$feed)

feed_params_all <- purrr::map(feed_types, function(ft) {
  # Filter by specified feed
  df <- feed_inputs %>% filter(feed == ft) 

  list(
    Proteins = df %>% 
      select(ingredient, proportion, contains("protein")) %>% 
      rename(
        macro = protein, 
        mean_digest = mean_protein_digestibility,
        min_digest = min_protein_digestibility,
        max_digest = max_protein_digestibility
      ),
    Carbohydrates = df %>% 
      select(ingredient, proportion, contains("carb")) %>% 
      rename(
        macro = carb, 
        mean_digest = mean_carb_digestibility,
        min_digest = min_carb_digestibility,
        max_digest = max_carb_digestibility
      ),
    Lipids = df %>% 
      select(ingredient, proportion, contains("lipid")) %>% 
      rename(
        macro = lipid, 
        mean_digest = mean_lipid_digestibility,
        min_digest = min_lipid_digestibility,
        max_digest = max_lipid_digestibility
      )
  )
}) %>% 
  setNames(feed_types)
```

This next chunk creates the mean, minimum and maximum digestibility variants of each of the feeds. 

```{r min-max-values}

make_mean_digest <- function(df) {
  df %>% 
    mutate(digest = mean_digest) %>% 
    select(ingredient, proportion, macro, digest)
}
make_min_digest <- function(df) {
  df %>% 
    mutate(digest = min_digest) %>% 
    select(ingredient, proportion, macro, digest)
}
make_max_digest <- function(df) {
  df %>% 
    mutate(digest = max_digest) %>% 
    select(ingredient, proportion, macro, digest)
}

feed_params <- list(
  "marine_dominant_biomar"     = feed_params_all[["marine_dominant_biomar"]] %>% lapply(make_mean_digest),
  "marine_dominant_biomar_min" = feed_params_all[["marine_dominant_biomar"]] %>% lapply(make_min_digest),
  "marine_dominant_biomar_max" = feed_params_all[["marine_dominant_biomar"]] %>% lapply(make_max_digest),
  "animal_inclusive_biomar"     = feed_params_all[["animal_inclusive_biomar"]] %>% lapply(make_mean_digest),
  "animal_inclusive_biomar_min" = feed_params_all[["animal_inclusive_biomar"]] %>% lapply(make_min_digest),
  "animal_inclusive_biomar_max" = feed_params_all[["animal_inclusive_biomar"]] %>% lapply(make_max_digest),
  "novel_inclusive_biomar"     = feed_params_all[["novel_inclusive_biomar"]] %>% lapply(make_mean_digest),
  "novel_inclusive_biomar_min" = feed_params_all[["novel_inclusive_biomar"]] %>% lapply(make_min_digest),
  "novel_inclusive_biomar_max" = feed_params_all[["novel_inclusive_biomar"]] %>% lapply(make_max_digest)
  )
```

```{r save}
#| code-summary: Save feed params file

qsave(feed_params, file.path(output_species_data_path, "feed_params_all.qs"))

feed_params_MD <- feed_params[1:3]
feed_params_PD <- feed_params[7:9]
feed_params_AD <- feed_params[4:6]

names(feed_params_PD) <- c("plant_dominant_biomar", "plant_dominant_biomar_min", "plant_dominant_biomar_max")
names(feed_params_AD) <- c("animal_dominant_biomar", "animal_dominant_biomar_min", "animal_dominant_biomar_max")

qsave(feed_params_MD, file.path(output_species_data_path, "feed_params_MD.qs"))
qsave(feed_params_PD, file.path(output_species_data_path, "feed_params_PD.qs"))
qsave(feed_params_AD, file.path(output_species_data_path, "feed_params_AD.qs"))
```

# For exploring C/N

```{r temporary-CN-exploring}
feed_params_MD <- qread(file.path(output_species_data_path, "feed_params_MD.qs"))
feed_params_PD <- qread(file.path(output_species_data_path, "feed_params_PD.qs"))
feed_params_AD <- qread(file.path(output_species_data_path, "feed_params_AD.qs"))

# Look at the most important lipid ingredients
rbind(
  feed_params_MD[[1]][["Lipids"]] %>% mutate(total = proportion * macro),
  feed_params_PD[[1]][["Lipids"]] %>% mutate(total = proportion * macro),
  feed_params_AD[[1]][["Lipids"]] %>% mutate(total = proportion * macro)
  ) %>% 
    group_by(ingredient) %>% 
    reframe(total = sum(total)) %>% 
    arrange(-total)

# These are the top ten lipid-providing ingredients in the feeds.
# canola-oil
# fish-oil-forage
# poultry-oil
# algal-oil
# fish-oil-trimmings
# fishmeal-forage
# poultry-meal
# fishmeal-trimmings
# wheat
# soy-protein-concentrate
```

These are the top ten lipid-providing ingredients in the feeds:

   ingredient                total
   <fct>                     <dbl>
 1 canola-oil              0.579  
 2 fish-oil-forage         0.138  
 3 poultry-oil             0.120  
 4 algal-oil               0.0880 
 5 fish-oil-trimmings      0.0460 
 6 fishmeal-forage         0.0341 
 7 poultry-meal            0.0164 
 8 fishmeal-trimmings      0.0102 
 9 wheat                   0.00884
10 soy-protein-concentrate 0.00792

# Feed information

```{r}
proteins <- feed_params %>% 
  lapply(function(ls) {
    ls$Proteins %>% 
      mutate(prop_macro = proportion * macro) %>% 
      pull(prop_macro) %>% sum()
  }) %>% 
    as.data.frame() %>% 
    pivot_longer(cols = names(feed_params), names_to = "feed", values_to = "protein")

lipids <- feed_params %>% 
  lapply(function(ls) {
    ls$Lipids %>% 
      mutate(prop_macro = proportion * macro) %>% 
      pull(prop_macro) %>% sum()
  }) %>% 
    as.data.frame() %>% 
    pivot_longer(cols = names(feed_params), names_to = "feed", values_to = "lipid")

carbs <- feed_params %>% 
  lapply(function(ls) {
    ls$Carbohydrates %>% 
      mutate(prop_macro = proportion * macro) %>% 
      pull(prop_macro) %>% sum()
  }) %>% 
    as.data.frame() %>% 
    pivot_longer(cols = names(feed_params), names_to = "feed", values_to = "carbohydrate")

proteins %>% 
  left_join(lipids, by = "feed") %>% 
  left_join(carbs, by = "feed") %>% 
  mutate(total = protein+lipid+carbohydrate)
```

```{r}
proteins <- feed_params %>% 
  lapply(function(ls) {
    ls$Proteins %>% 
      filter(macro > 0) %>% 
      mutate(prop_digest = proportion * digest / sum(proportion)) %>% 
      pull(prop_digest) %>% sum()
  }) %>% 
    as.data.frame() %>% 
    pivot_longer(cols = names(feed_params), names_to = "feed", values_to = "AD_protein")

lipids <- feed_params %>% 
  lapply(function(ls) {
    ls$Lipids %>% 
      filter(macro > 0) %>% 
      mutate(prop_digest = proportion * digest / sum(proportion)) %>% 
      pull(prop_digest) %>% sum()
  }) %>% 
    as.data.frame() %>% 
    pivot_longer(cols = names(feed_params), names_to = "feed", values_to = "AD_lipid")

carbs <- feed_params %>% 
  lapply(function(ls) {
    ls$Carbohydrates %>% 
      filter(macro > 0) %>% 
      mutate(prop_digest = proportion * digest / sum(proportion)) %>% 
      pull(prop_digest) %>% sum()
  }) %>% 
    as.data.frame() %>% 
    pivot_longer(cols = names(feed_params), names_to = "feed", values_to = "AD_carbohydrate")

proteins %>% 
  left_join(lipids, by = "feed") %>% 
  left_join(carbs, by = "feed")
```

# Comparing feeds

The three "biomar" feeds were designed to be isoenergetic.

```{r feed-stats}
#| code-summary: Calculate energy stats about the feeds

feed_params[[1]] %>% 
  lapply(function(ls) {
    ls %>% mutate(total = proportion * macro * digest) %>% 
      pull(total) %>% sum()
  }) %>% as.data.frame() %>% 
    mutate(
      Proteins = Proteins * 17,
      Carbohydrates = Carbohydrates * 37.6,
      Lipids = Lipids * 16.8,
      total = Proteins + Carbohydrates + Lipids
    )


feed_stats <- 
  map2(feed_params, names(feed_params), function(ls, feedname) {
    map2(ls, names(ls), function(df, macroname) {
      df %>% mutate(component = macroname)
    }) %>% bind_rows() %>% 
      mutate(feed = feedname)
  }) %>% 
  bind_rows() %>% 
  mutate(
    ing_macro = proportion * macro,
    digest_macro = ing_macro * digest,
    E = case_when(
      component == "Proteins" ~ digest_macro * 17,
      component == "Lipids" ~ digest_macro * 37.6,
      T ~ digest_macro * 16.8
    )
  ) %>% 
  group_by(feed) %>% 
  mutate(total_E = sum(E)) %>% 
  ungroup() %>% 
  group_by(feed, component, total_E) %>% 
  reframe(comp_E = sum(E)) %>% 
  mutate(prop_E = comp_E/total_E)

feed_stats %>% 
  group_by(feed) %>% 
  reframe(total_E = mean(total_E))

feed_stats %>% 
  filter(component == "Proteins")
```

```{r feed-macro-digestibility}
#| code-summary: Code to generate plot
#| eval: false
#| fig-cap: Total protein, carbohydrate and lipid composition of each feed, with the undigestible portion shown as the lighter bars on top. 

df <- merge(
  feed_stats %>% 
    select(feed, group, contains("total_")) %>% 
    pivot_longer(
      cols = contains("total"),
      names_to = "macro", names_prefix = "total_",
      values_to = "total"
    ),
  feed_stats %>% 
    select(feed, group, contains("AD_")) %>% 
    pivot_longer(
      cols = contains("AD"),
      names_to = "macro", names_prefix = "AD_",
      values_to = "AD"
    ),
  by = c("feed", "group", "macro")
  ) %>% 
  mutate(feed = as.factor(feed)) 
  
df %>% 
  filter(feed %in% c("animal_inclusive_biomar", "marine_dominant_biomar", "novel_inclusive_biomar"))

df %>% 
  ggplot(aes(x = macro, fill = feed)) +
  geom_col(aes(y = total), position = position_dodge(), alpha = 0.5, colour = "black") +
  geom_col(aes(y = AD), position = position_dodge(), alpha = 0.75, colour = "black") +
  theme_classic() +
  facet_wrap(facets = vars(group))
```

```{r feed-energy}
#| code-summary: Code to generate plot
#| eval: false
#| fig-cap: The proportion of energy derived from protein, carbohydrate and lipid within each feed.

feed_stats %>% 
  mutate(
    E_protein = E_protein/E,
    E_carb = E_carb/E,
    E_lipid = E_lipid/E
  ) %>% 
  select(feed, group, contains("E_")) %>% 
  pivot_longer(
    cols = contains("E_"),
    names_to = "macro", names_prefix = "E_",
    values_to = "value"
  ) %>% 
  ggplot(aes(x = macro, y = value, fill = feed)) +
  geom_col(position = position_dodge(), colour = "black") +
  theme_classic() +
  facet_wrap(facets = vars(group))
```
