---
title: "7: How many runs per farm?"
subtitle: "Local to global salmon aquaculture waste modelling"
author:
  - "Tormey Reimer"
date: last-modified
---

::: {style="display: flex; justify-content: space-between;"}
[← Previous: Assigning production](06_assigning%20production.qmd)

[Next: Run Atlantic salmon model →](08_run_atlantic_salmon_model.qmd)
:::

# Overview {#sec-overview}

Originally, each farm was run with 5000 Monte-Carlo simulation runs. This takes quite a long time (4-5 minutes per farm and feed). But is this necessary?

```{r setup}
#| cache: false
#| echo: false
#| warning: false
#| message: false
#| output: false

library(tidyverse)
library(arrow)
library(sf)
library(terra)
library(purrr)
library(furrr)
library(future)
library(tictoc)
library(fs)
library(conflicted)
library(readxl)
library(units)
library(qs)
library(here)
library(targets)
conflicted::conflicts_prefer(dplyr::filter(), dplyr::select(), .quiet = T)

overwrite <- F # Make sure to set this to FALSE for quarto rendering
```

```{r source-scripts}
#| cache: false
#| echo: false
#| warning: false
#| message: false
#| output: false
#| code-summary: "Source R scripts in use"

here("src") %>%
  list.files(pattern = "\\.R$", full.names = TRUE) %>%
  str_subset("map", negate = T) %>%
  walk(source)
```

# Set up data for MC testing {#sec-setup-data}

Only use a small but representative sample of farms (n = 50) across the mean temperature spectrum.

```{r set-up-data}
#| code-summary: "Set up data"

farm_ts_data <- file.path(input_farm_coords_path, "farm_temporal_data.qs") %>% qread()
farm_static_data <- file.path(input_farm_coords_path, "farm_static_data_w_stocking.qs") %>% qread()

feed_params <- file.path(output_species_data_path, "feed_params_PD.qs") %>% qread()
feed_params <- feed_params[["plant_dominant_biomar"]]

all_params <- c(
  file.path(output_species_data_path, "species_params.qs") %>% qread(),
  file.path(output_species_data_path, "pop_params.qs") %>% qread()
)

# Get a small sample of farms
samp <- farm_ts_data %>% 
  group_by(farm_id) %>% 
  reframe(mean = mean(temp_c)) %>% 
  mutate(temp_bin = ntile(mean, 10)) %>%
  group_by(temp_bin) %>%
  slice_sample(n = 5) %>%
  ungroup() %>%
  pull(farm_id)

farm_ts_data <- farm_ts_data %>% 
  rename(farm_ID = farm_id) %>% 
  filter(farm_ID %in% samp) %>% 
  group_by(farm_ID) %>% 
  group_split()

farm_static_data <- farm_static_data %>% 
  rename(farm_ID = farm_id) %>% 
  filter(farm_ID %in% samp) %>% 
  group_by(farm_ID) %>% 
  group_split()
```

```{r add-population}
#| code-summary: "Add population structure to farm time-series data"

# For testing:
# ts <- farm_ts_data[[1]]
# static <- farm_static_data[[1]]

farm_ts_data_2 <- purrr::map2(
  farm_static_data, 
  farm_ts_data, 
  function(static, ts) {
    times <- c(t_start = static$t_start, t_end = static$t_end, dt = 1)

    cbind(
      ts %>% 
        filter(day %in% seq(static$t_start, static$t_end, 1)) %>% 
        arrange(day),
      data.frame(Npop = generate_pop(
          harvest_n = static$harvest_n,
          mort = all_params['mortmyt'],
          times = times
      ))
    )
  },
  .progress = T
)
```

```{r define-model-runs}
#| code-summary: "Set up different farm runs as functions, define multisessions"

plan(multisession, workers = parallelly::availableCores()-2)

# For testing:
# ts <- farm_ts_data_2[[1]]
# static <- farm_static_data[[1]]

# This is uniform
farm_run_singular <- function() {
  furrr::future_map2_dfr(
    farm_ts_data_2, 
    farm_static_data, 
    function(ts, static) {
      fg <- uni_farm_growth(
        pop_params = all_params,
        species_params = all_params,
        water_temp = ts$temp_c,
        feed_params = feed_params,
        times = c(t_start = static$t_start, t_end = static$t_end, dt = 1),
        N_pop = ts$Npop
      )

      # mat <- fg[[1]]
      fg %>%
         purrr::map_dfr(function(mat) {
          mat %>% 
            as.data.frame() %>% 
            rename(t = V1, mean = V2, sd = 0) %>% 
            mutate(
              farm_ID = as.integer(static$farm_ID),
              t = as.integer(t)
            )
          }, .id = "measure"
        ) %>% 
        mutate(measure = as.factor(measure))
  },
  .options = furrr_options(seed = T),
  .progress = T
  )
}

# This has the MC runs in it
# n <- 500
farm_run_MC = function(n) {
  furrr::future_map2_dfr(
    farm_ts_data_2, 
    farm_static_data, 
    function(ts, static) {
      fg <- farm_growth(
        pop_params = all_params,
        species_params = all_params,
        water_temp = ts$temp_c,
        feed_params = feed_params,
        times = c(t_start = static$t_start, t_end = static$t_end, dt = 1),
        N_pop = ts$Npop,
        nruns = n
      )
      fg %>% 
        purrr::map_dfr(function(mat) {
          mat %>% 
            as.data.frame() %>% 
            rename(t = V1, mean = V2, sd = V3) %>% 
            mutate(
              farm_ID = as.integer(static$farm_ID),
              t = as.integer(t)
            )
          }, 
          .id = "measure"
        ) %>% 
        mutate(measure = as.factor(measure))
    },
  .options = furrr_options(seed = T),
  .progress = T
  )
}
```

# Save runs {#sec-save-runs}

These take quite a long time (a couple hours) so best not to do this again if you don't have to.

```{r save-nruns-5000}
#| code-summary: "Run the model with 5000 nruns (default) and save"

res_5000 <- farm_run_MC(5000) 

res_5000 <- res_5000 %>% 
  mutate(nruns = as.factor(5000)) %>% 
  filter(measure == "biomass_stat")

qsave(res_5000, file.path(output_sens_data_path, "res_nruns_5000.qs"))
```

```{r save-nruns-1}
#| code-summary: "Run the model with nrun = 1"

res_1 <- farm_run_singular() 

res_1 <- res_1 %>% 
  mutate(nruns = as.factor(1), sd = 0) %>% 
  relocate(sd, .after = mean) %>% 
  filter(measure == "biomass_stat")

qsave(res_1, file.path(output_sens_data_path, "res_nruns_1.qs"))
```

```{r save-nruns-other}
#| code-summary: "Run the model with other numbers of nruns and save"

for (i in c(100, 250, 500, 1000, 2500)) {
  fnm <- file.path(output_sens_data_path, paste0("res_nruns_", i, ".qs"))
  print(fnm)
  res <- farm_run_MC(i)
  res <- res %>% 
    mutate(nruns = as.factor(i)) %>% 
    filter(measure == "biomass_stat")
  qsave(res, fnm)
}
```

```{r read-in-all-runs}
#| code-summary: Read in previously saved run data

res_1 <- qread(file.path(output_sens_data_path, "res_nruns_1.qs"))
res_100 <- qread(file.path(output_sens_data_path, "res_nruns_100.qs"))
res_250 <- qread(file.path(output_sens_data_path, "res_nruns_250.qs"))
res_500 <- qread(file.path(output_sens_data_path, "res_nruns_500.qs"))
res_1000 <- qread(file.path(output_sens_data_path, "res_nruns_1000.qs"))
res_2500 <- qread(file.path(output_sens_data_path, "res_nruns_2500.qs"))
res_5000 <- qread(file.path(output_sens_data_path, "res_nruns_5000.qs"))
```

# Difference between uniform run and 5000 nruns {#sec-diff-uniform}

How does the MC run approach using 5000 runs (the default) differ from only using a uniform fish weight/maximum ingestion?

```{r fig-1-nruns-weight}
#| code-summary: "Plot individual weight timeseries comparing 1 run vs 5000 Monte-Carlo runs"
#| fig-width: 8.5
#| fig-cap: "Individual weight at each timestep, when the farm has been run with 5000 Monte-Carlo runs (with a range of fish weights/maximum ingestion values) and when only mean fish weight/maximum ingestion was used."

ggplot(
  rbind(res_1, res_5000),
  aes(x = t, y = mean, ymin = mean-sd, ymax = mean+sd, colour = nruns, fill = nruns)
) +
  geom_line() +
  geom_ribbon(alpha = 0.25) +
  facet_wrap(facets = vars(farm_ID), nrow = 5) +
  theme_classic()
```

```{r fig-1-nruns-weight-diff}
#| code-summary: "Plot percentage difference in individual weight between 1 run and 5000 Monte-Carlo runs"
#| fig-width: 8.5
#| fig-cap: "Percentage difference in individual weight at each timestep between a single run and 5000 Monte-Carlo runs."

df_diff_1 <- rbind(res_1, res_5000) %>% 
  mutate(sd = sd/mean) %>% 
  pivot_wider(
    names_from = "nruns", names_sep = ".", 
    values_from = c(mean, sd)
    ) %>% 
    mutate(
      mean_diff = (mean.1 - mean.5000)/mean.5000,
      sd_diff = mean_diff*sqrt(sd.1^2 + sd.5000^2)
    )

ggplot(df_diff_1, aes(x = t, y = 100*mean_diff, ymin = mean_diff-sd_diff, ymax = mean_diff+sd_diff)) +
  geom_line() +
  facet_wrap(facets = vars(farm_ID), nrow = 5) +
  theme_classic()
```

## Difference between 100 and 5000 nruns {#sec-diff-100}

```{r fig-100-nruns-weight}
#| fig-width: 8.5
#| fig-cap: Individual weight at each timestep, when the farm has been run with 5000 Monte-Carlo runs vs 100 Monte-Carlo runs.

ggplot(
  rbind(res_100, res_5000),
  aes(x = t, y = mean, ymin = mean-sd, ymax = mean+sd, colour = nruns, fill = nruns)
) +
  geom_line() +
  geom_ribbon(alpha = 0.25) +
  facet_wrap(facets = vars(farm_ID), nrow = 5) +
  theme_classic()
```

```{r fig-100-nruns-weight-diff}
#| fig-width: 8.5
#| fig-cap: Difference in individual weight at each timestep when the farm has been run with 5000 vs 100 Monte-Carlo runs.

df_diff_100 <- rbind(res_100, res_5000) %>% 
  mutate(sd = sd/mean) %>% 
  pivot_wider(
    names_from = "nruns", names_sep = ".", 
    values_from = c(mean, sd)
    ) %>% 
    mutate(
      mean_diff = (mean.100 - mean.5000)/mean.5000,
      sd_diff = mean_diff*sqrt(sd.100^2 + sd.5000^2)
    )

ggplot(df_diff_100, aes(x = t, y = 100*mean_diff, ymin = mean_diff-sd_diff, ymax = mean_diff+sd_diff)) +
  geom_line() +
  facet_wrap(facets = vars(farm_ID), nrow = 5) +
  theme_classic()
```

## Difference between 250 and 5000 nruns {#sec-diff-250}

```{r fig-250-nruns-weight}
#| fig-width: 8.5
#| fig-cap: Individual weight at each timestep, when the farm has been run with 5000 Monte-Carlo runs vs 250 Monte-Carlo runs.

ggplot(
  rbind(res_250, res_5000),
  aes(x = t, y = mean, ymin = mean-sd, ymax = mean+sd, colour = nruns, fill = nruns)
) +
  geom_line() +
  geom_ribbon(alpha = 0.25) +
  facet_wrap(facets = vars(farm_ID), nrow = 5) +
  theme_classic()
```

```{r fig-250-nruns-weight-diff}
#| fig-width: 8.5
#| fig-cap: Difference in individual weight at each timestep when the farm has been run with 5000 vs 250 Monte-Carlo runs.

df_diff_250 <- rbind(res_250, res_5000) %>% 
  mutate(sd = sd/mean) %>% 
  pivot_wider(
    names_from = "nruns", names_sep = ".", 
    values_from = c(mean, sd)
    ) %>% 
    mutate(
      mean_diff = (mean.250 - mean.5000)/mean.5000,
      sd_diff = mean_diff*sqrt(sd.250^2 + sd.5000^2)
    )

ggplot(df_diff_250, aes(x = t, y = 100*mean_diff, ymin = mean_diff-sd_diff, ymax = mean_diff+sd_diff)) +
  geom_line() +
  facet_wrap(facets = vars(farm_ID), nrow = 5) +
  theme_classic()
```

## Difference between 500 and 5000 nruns {#sec-diff-500}

```{r fig-500-nruns-weight}
#| fig-width: 8.5
#| fig-cap: Individual weight at each timestep, when the farm has been run with 5000 Monte-Carlo runs vs 500 Monte-Carlo runs.

ggplot(
  rbind(res_500, res_5000),
  aes(x = t, y = mean, ymin = mean-sd, ymax = mean+sd, colour = nruns, fill = nruns)
) +
  geom_line() +
  geom_ribbon(alpha = 0.25) +
  facet_wrap(facets = vars(farm_ID), nrow = 5) +
  theme_classic()
```

```{r fig-500-nruns-weight-diff}
#| fig-width: 8.5
#| fig-cap: Difference in individual weight at each timestep when the farm has been run with 5000 vs 500 Monte-Carlo runs.

df_diff_500 <- rbind(res_500, res_5000) %>% 
  mutate(sd = sd/mean) %>% 
  pivot_wider(
    names_from = "nruns", names_sep = ".", 
    values_from = c(mean, sd)
    ) %>% 
    mutate(
      mean_diff = (mean.500 - mean.5000)/mean.5000,
      sd_diff = mean_diff*sqrt(sd.500^2 + sd.5000^2)
    )

ggplot(df_diff_500, aes(x = t, y = 100*mean_diff, ymin = mean_diff-sd_diff, ymax = mean_diff+sd_diff)) +
  geom_line() +
  facet_wrap(facets = vars(farm_ID), nrow = 5) +
  theme_classic()
```

## Difference between 1000 and 5000 nruns {#sec-diff-1000}

```{r fig-1000-nruns-weight}
#| fig-width: 8.5
#| fig-cap: Individual weight at each timestep, when the farm has been run with 5000 Monte-Carlo runs vs 1000 Monte-Carlo runs.

ggplot(
  rbind(res_1000, res_5000),
  aes(x = t, y = mean, ymin = mean-sd, ymax = mean+sd, colour = nruns, fill = nruns)
) +
  geom_line() +
  geom_ribbon(alpha = 0.25) +
  facet_wrap(facets = vars(farm_ID), nrow = 5) +
  theme_classic()
```

```{r fig-1000-nruns-weight-diff}
#| fig-width: 8.5
#| fig-cap: Difference in individual weight at each timestep when the farm has been run with 5000 vs 1000 Monte-Carlo runs.

df_diff_1000 <- rbind(res_1000, res_5000) %>% 
  mutate(sd = sd/mean) %>% 
  pivot_wider(
    names_from = "nruns", names_sep = ".", 
    values_from = c(mean, sd)
    ) %>% 
    mutate(
      mean_diff = (mean.1000 - mean.5000)/mean.5000,
      sd_diff = mean_diff*sqrt(sd.1000^2 + sd.5000^2)
    )

ggplot(df_diff_1000, aes(x = t, y = 100*mean_diff, ymin = mean_diff-sd_diff, ymax = mean_diff+sd_diff)) +
  geom_line() +
  facet_wrap(facets = vars(farm_ID), nrow = 5) +
  theme_classic()
```

## Difference between 2500 and 5000 nruns {#sec-diff-2500}

```{r fig-2500-nruns-weight}
#| fig-width: 8.5
#| fig-cap: Individual weight at each timestep, when the farm has been run with 5000 Monte-Carlo runs vs 2500 Monte-Carlo runs.

ggplot(
  rbind(res_2500, res_5000),
  aes(x = t, y = mean, ymin = mean-sd, ymax = mean+sd, colour = nruns, fill = nruns)
) +
  geom_line() +
  geom_ribbon(alpha = 0.25) +
  facet_wrap(facets = vars(farm_ID), nrow = 5) +
  theme_classic()
```

```{r fig-2500-nruns-weight-diff}
#| fig-width: 8.5
#| fig-cap: Difference in individual weight at each timestep when the farm has been run with 5000 vs 2500 Monte-Carlo runs.

df_diff_2500 <- rbind(res_2500, res_5000) %>% 
  mutate(sd = sd/mean) %>% 
  pivot_wider(
    names_from = "nruns", names_sep = ".", 
    values_from = c(mean, sd)
    ) %>% 
    mutate(
      mean_diff = (mean.2500 - mean.5000)/mean.5000,
      sd_diff = mean_diff*sqrt(sd.2500^2 + sd.5000^2)
    )

ggplot(df_diff_2500, aes(x = t, y = 100*mean_diff, ymin = mean_diff-sd_diff, ymax = mean_diff+sd_diff)) +
  geom_line() +
  facet_wrap(facets = vars(farm_ID), nrow = 5) +
  theme_classic()
```

# Summary {#sec-summary}

```{r summary}
#| code-summary: "Summarise maximum difference from 5000-run baseline across all nrun settings"
df_diff <- data.frame(
  nruns = c(1, 100, 250, 500, 1000, 2500),
  max_diff = c(
    max(abs(df_diff_1$mean_diff)),
    max(abs(df_diff_100$mean_diff)),
    max(abs(df_diff_250$mean_diff)),
    max(abs(df_diff_500$mean_diff)),
    max(abs(df_diff_1000$mean_diff)),
    max(abs(df_diff_2500$mean_diff))
  )
)

ggplot(df_diff, aes(x = as.factor(nruns), y = 100*max_diff)) +
  geom_col() + theme_classic()
```

::: {style="display: flex; justify-content: space-between;"}
[← Previous: Assigning production](06_assigning%20production.qmd)

[Next: Run Atlantic salmon model →](08_run_atlantic_salmon_model.qmd)
:::
