---
title: "Process data from Atlantic salmon model runs"
author: 
  - "Tormey Reimer"
date: last-modified
code-fold: true
execute:
  warning: false
  message: false
---

# Introduction

```{r setup}
#| cache: false
#| include: false

library(tidyverse)
library(arrow)
library(sf)
library(terra)
library(magrittr)
library(furrr)
library(future)
library(tictoc)
library(fs)
library(conflicted)
library(readxl)
library(units)
library(qs)
library(here)
library(scales)
library(targets)
library(ggforce)

conflicted::conflicts_prefer(
  dplyr::filter(), 
  dplyr::select(), 
  scales::rescale(), 
  .quiet = T
)

here("src") %>% 
  list.files(pattern = "\\.R$", full.names = TRUE) %>% 
  walk(source)

remove_unit("g_fish")
remove_unit("kg_fish")
remove_unit("t_fish")
install_unit("g_fish")
install_unit("kg_fish", "1000 g_fish")
install_unit("t_fish", "1000 kg_fish")
```

# Get outputs

These are the outputs that I want to view and analyse.

## Total (summed) values

The total values describe selected output variables (measures) summed over the entire period of a salmon growing period (548 days). The data were generated from a suite of model runs which cover several dimensions:

- 2697 unique farms
- 3 different feed groups, representing different formulations of salmon feed: MD (marine-dominant), AD (animal-dominant) and PD (plant-dominant)
- 3 different feed types representing levels of digestibility of each of the three different feeds: mean digestibility, minimum digestibility, and maximum digestibility

The output values are expressed per biomass of salmon produced.

```{r total-values}
#| code-summary: Get total values data. These data give the total (summed) values of each response metric over 1 complete production period.

overwrite <- F

total_vals <- file.path(data_analysis_path, "total_vals.qs")
total_props <- file.path(data_analysis_path, "total_props.qs")

if (any(!file.exists(c(total_vals, total_props))) | overwrite) {

  total_vals <- file.path(output_path, "total_ends.qs") %>% 
    qread() %>% 
    mutate(
      feed_group = factor(
        str_remove(feed, "_min|_max"), 
        levels = orig_feeds,
        labels = short_feeds
        ),
      feed_type = case_when(
        str_detect(feed, "_min") ~ "min",
        str_detect(feed, "_max") ~ "max",
        T ~ "mean"
        ),
      feed_type = factor(feed_type, levels = c("min", "mean", "max"), labels = c("Minimum", "Mean", "Maximum")),
      mean = set_units(mean, "g"),
      sd = set_units(sd, "g"),
      mean_biomass = set_units(mean_biomass, "g_fish"),
      sd_biomass = set_units(sd_biomass, "g_fish"),
      mean_per_biom = set_units(mean_per_biom, "g g_fish-1"),
      sd_per_biom = set_units(sd_per_biom, "g g_fish-1")
    ) %>% 
    select(-feed)

  total_vals_1 <- total_vals %>% 
    filter(measure %in% c("food_prov_stat", "P_excr_stat", "P_uneat_stat", "total_excr_stat", "total_uneat_stat")) %>% 
    mutate(measure = droplevels(measure)) %>% 
    select(farm_ID, measure, mean_per_biom, feed_group, feed_type) %>% 
    pivot_wider(
      names_from = measure,
      values_from = mean_per_biom
    ) %>% 
    mutate(
      total_waste_stat = total_excr_stat + total_uneat_stat,
      total_P_lost_stat = P_excr_stat + P_uneat_stat
    ) %>% 
    pivot_longer(
      cols = contains("stat"),
      names_to = "measure", names_transform = as.factor,
      values_to = "mean_per_biom"
    )

  total_vals_2 <- total_vals %>% 
    filter(measure %in% c("food_prov_stat", "P_excr_stat", "P_uneat_stat", "total_excr_stat", "total_uneat_stat")) %>% 
    mutate(measure = droplevels(measure)) %>% 
    select(farm_ID, measure, mean, feed_group, feed_type) %>% 
    pivot_wider(
      names_from = measure, 
      values_from = mean
    ) %>% 
    mutate(
      total_waste_stat = total_excr_stat + total_uneat_stat,
      total_P_lost_stat = P_excr_stat + P_uneat_stat
    ) %>% 
    pivot_longer(
      cols = contains("stat"),
      names_to = "measure", names_transform = as.factor,
      values_to = "mean"
    )

  total_vals <- total_vals_1 %>% 
    left_join(total_vals_2, by = c("farm_ID", "feed_group", "feed_type", "measure"))

  total_props <- total_vals %>% 
    pivot_wider(
      names_from = measure, names_sep = ".",
      values_from = c(mean, mean_per_biom)
    ) %>% 
    mutate(
      prop_excr_stat = mean.total_excr_stat/mean.total_waste_stat, # The proportion of total waste (feed+faeces) that comes from excreted faeces
      prop_P_stat = mean.P_excr_stat/mean.total_P_lost_stat, # The proportion of total protein lost (feed+faeces) that is lost through excretion
      prop_uneat_stat = mean.total_uneat_stat/mean.food_prov_stat, # The proportion of feed provided that is lost to the environment
      prop_comp_excr_P_stat = mean.P_excr_stat/mean.total_excr_stat, # The composition of faeces, ie how much of faeces is protein
    ) %>% 
    select(farm_ID, feed_group, feed_type, contains("prop"))
    
  total_props <- total_props %>% 
    select(farm_ID, feed_group, feed_type, contains("prop")) %>% 
    pivot_longer(
      cols = contains("prop"), 
      names_to = c("measure"), names_transform = as.factor,
      values_to = "mean_prop", values_transform = drop_units
    )

  qsave(total_vals, file.path(data_analysis_path, "total_vals.qs"))
  qsave(total_props, file.path(data_analysis_path, "total_props.qs"))
}

total_vals <- file.path(data_analysis_path, "total_vals.qs") %>% 
  qread() %>% 
  mutate(feed_group = factor(feed_group, levels = c("MD","PD","AD")))
total_props <- file.path(data_analysis_path, "total_props.qs") %>% 
  qread() %>% 
  mutate(feed_group = factor(feed_group, levels = c("MD","PD","AD")))
```

The structure of these data are as follows:

```{r}
total_vals %>% 
  str()
```

```{r}
total_props %>% 
  str()
```

## Daily values

The daily values describe selected output variables (measures) daily over one year of farm operations, which includes 2 overlapping cohorts of salmon. The data were generated from the same suite of model runs as the total values, so the only difference between the structure of these data and the previous data is that they also include a calendar DOY (`t`) and a production day (`prod_t`). The output values are expressed per biomass of salmon produced.

``` {r daily-values}
#| code-summary: Get daily values data. These data give the daily values of each response metric over 1 calendar year (365 days), which encompasses 2 overlapping production cohorts.

overwrite <- F

cleanup_daily_vals <- function(df) {
  df %>% 
    mutate(
      feed_group = factor(
        str_remove(feed, "_min|_max"), 
        levels = orig_feeds,
        labels = short_feeds
        ),
      feed_group = factor(feed_group, levels = c("MD","PD","AD")),
      feed_type = case_when(
        str_detect(feed, "_min") ~ "min",
        str_detect(feed, "_max") ~ "max",
        T ~ "mean"
        ),
      feed_type = factor(
        feed_type, 
        levels = c("min", "mean", "max"), 
        labels = c("Minimum", "Mean", "Maximum")
      ),
      mean_per_biom = set_units(mean_per_biom, "g g_fish-1"),
      sd_per_biom = set_units(sd_per_biom, "g g_fish-1"),
      mean = set_units(mean, "g"),
      sd = set_units(sd, "g")
    ) %>% 
    select(-c(feed, mean_biomass, sd_biomass, cohorts))
}

daily_vals.food_prov_stat <- file.path(data_analysis_path, "daily_vals.food_prov_stat.qs")
daily_vals.P_excr_stat <- file.path(data_analysis_path, "daily_vals.P_excr_stat.qs")
daily_vals.P_uneat_stat <- file.path(data_analysis_path, "daily_vals.P_uneat_stat.qs")
daily_vals.total_excr_stat <- file.path(data_analysis_path, "daily_vals.total_excr_stat.qs")
daily_vals.total_P_lost_stat <- file.path(data_analysis_path, "daily_vals.total_P_lost_stat.qs")
daily_vals.total_uneat_stat <- file.path(data_analysis_path, "daily_vals.total_uneat_stat.qs")
daily_vals.total_waste_stat <- file.path(data_analysis_path, "daily_vals.total_waste_stat.qs")
daily_props <- file.path(data_analysis_path, "daily_props.qs")

if (any(!file.exists(c(daily_vals.food_prov_stat, daily_vals.P_excr_stat, daily_vals.P_uneat_stat, daily_vals.total_excr_stat, daily_vals.total_P_lost_stat, daily_vals.total_uneat_stat, daily_vals.total_waste_stat, daily_props))) | overwrite) {

  daily_files <- c(
    file.path(output_path, "cohort_results_food_prov_stat.qs"),
    file.path(output_path, "cohort_results_P_excr_stat.qs"),
    file.path(output_path, "cohort_results_P_uneat_stat.qs"),
    file.path(output_path, "cohort_results_total_uneat_stat.qs"),
    file.path(output_path, "cohort_results_total_excr_stat.qs")
  )

  daily_vals <- map_dfr(daily_files, function(filename) {
    qread(filename) %>% 
      cleanup_daily_vals()
  }) %>% 
    pivot_wider(
      names_from = measure, names_sep = ".",
      values_from = c(mean_per_biom, sd_per_biom, mean, sd)
    ) %>% 
    mutate(
      mean_per_biom.total_waste_stat = mean_per_biom.total_excr_stat + mean_per_biom.total_uneat_stat,
      mean_per_biom.total_P_lost_stat = mean_per_biom.P_excr_stat + mean_per_biom.P_uneat_stat,
      mean.total_waste_stat = mean.total_excr_stat + mean.total_uneat_stat,
      mean.total_P_lost_stat = mean.P_excr_stat + mean.P_uneat_stat,
      sd_per_biom.total_waste_stat = mean_per_biom.total_waste_stat * 
        ((sd_per_biom.total_excr_stat/mean_per_biom.total_excr_stat) + 
         (sd_per_biom.total_uneat_stat/mean_per_biom.total_uneat_stat)),
      sd_per_biom.total_P_lost_stat = mean_per_biom.total_P_lost_stat * 
        ((sd_per_biom.P_excr_stat/mean_per_biom.P_excr_stat) + 
         (sd_per_biom.P_uneat_stat/mean_per_biom.P_uneat_stat)),
      sd.total_excr_stat = mean.total_excr_stat * 
        ((sd.total_excr_stat/mean.total_excr_stat) + 
         (sd.total_uneat_stat/mean.total_uneat_stat)),
      sd.total_P_lost_stat = mean.total_P_lost_stat * 
        ((sd.P_excr_stat/mean.P_excr_stat) + 
         (sd.P_uneat_stat/mean.P_uneat_stat)),
    ) %>% 
    pivot_longer(
      cols = contains("stat"),
      names_to = c(".value", "measure"), 
      names_transform = list(measure = as.factor),
      names_sep = "\\.",
    )

  daily_vals %>% 
    filter(measure == "food_prov_stat") %>% mutate(measure = droplevels(measure)) %>% 
    qsave(., daily_vals.food_prov_stat)

  daily_vals %>% 
    filter(measure == "P_excr_stat") %>% mutate(measure = droplevels(measure)) %>% 
    qsave(., daily_vals.P_excr_stat)

  daily_vals %>% 
    filter(measure == "P_uneat_stat") %>% mutate(measure = droplevels(measure)) %>% 
    qsave(., daily_vals.P_uneat_stat)

  daily_vals %>% 
    filter(measure == "total_excr_stat") %>% mutate(measure = droplevels(measure)) %>% 
    qsave(., daily_vals.total_excr_stat)

  daily_vals %>% 
    filter(measure == "total_P_lost_stat") %>% mutate(measure = droplevels(measure)) %>% 
    qsave(., daily_vals.total_P_lost_stat)

  daily_vals %>% 
    filter(measure == "total_uneat_stat") %>% mutate(measure = droplevels(measure)) %>% 
    qsave(., daily_vals.total_uneat_stat)

  daily_vals %>% 
    filter(measure == "total_waste_stat") %>% mutate(measure = droplevels(measure)) %>% 
    qsave(., daily_vals.total_waste_stat)

  daily_props <- daily_vals %>% 
    pivot_wider(
      names_from = measure, names_sep = ".",
      values_from = c(mean_per_biom, sd_per_biom, mean, sd)
    ) %>% 
    mutate(
      mean_per_biom.prop_excr_stat = mean_per_biom.total_excr_stat/mean_per_biom.total_waste_stat, # The proportion of total waste (feed+faeces) that comes from excreted faeces
      mean_per_biom.prop_P_stat = mean_per_biom.P_excr_stat/mean_per_biom.total_P_lost_stat, # The proportion of total protein lost (feed+faeces) that is lost through excretion
      mean_per_biom.prop_uneat_stat = mean_per_biom.total_uneat_stat/mean_per_biom.food_prov_stat, # The proportion of feed provided that is lost to the environment
      mean.prop_excr_stat = mean.total_excr_stat/mean.total_waste_stat, 
      mean.prop_P_stat = mean.P_excr_stat/mean.total_P_lost_stat, 
      mean.prop_uneat_stat = mean.total_uneat_stat/mean.food_prov_stat, 
      sd_per_biom.prop_excr_stat = mean_per_biom.prop_excr_stat * 
        ((sd_per_biom.total_excr_stat/mean_per_biom.total_excr_stat) + 
        (sd_per_biom.total_waste_stat/mean_per_biom.total_waste_stat)), 
      sd_per_biom.prop_P_stat = mean_per_biom.prop_P_stat * 
        ((sd_per_biom.P_excr_stat/mean_per_biom.P_excr_stat) + 
        (sd_per_biom.total_P_lost_stat/mean_per_biom.total_P_lost_stat)), 
      sd_per_biom.prop_uneat_stat = mean_per_biom.prop_uneat_stat * 
        ((sd_per_biom.total_uneat_stat/mean_per_biom.total_uneat_stat) + 
        (sd_per_biom.food_prov_stat/mean_per_biom.food_prov_stat)), 
      sd.prop_excr_stat = mean.prop_excr_stat * 
        ((sd.total_excr_stat/mean.total_excr_stat) + 
        (sd.total_waste_stat/mean.total_waste_stat)), 
      sd.prop_P_stat = mean.prop_P_stat * 
        ((sd.P_excr_stat/mean.P_excr_stat) + 
        (sd.total_P_lost_stat/mean.total_P_lost_stat)), 
      sd.prop_uneat_stat = mean.prop_uneat_stat * 
        ((sd.total_uneat_stat/mean.total_uneat_stat) + 
        (sd.food_prov_stat/mean.food_prov_stat)), 
    ) %>% 
    select(farm_ID, feed_group, feed_type, prod_t, t, contains("prop")) %>% 
    pivot_longer(
      cols = contains("stat"),
      names_to = c(".value", "measure"), 
      names_transform = list(measure = as.factor),
      names_sep = "\\.",
    )

    qsave(daily_props, file.path(data_analysis_path, "daily_props.qs"))
    rm(daily_vals, daily_props)
}
```

The structure of these data are as follows:

```{r}
daily_vals.total_P_lost_stat %>% 
  qread() %>% 
  str()
```

```{r}
daily_props %>% 
  qread() %>% 
  str()
```

# Total values

The "total values" refers to the value summed across a single production run, with no overlapping cohorts. This is where the mean/biomass metric is most useful, as it refers to the mean amount required to produce the required biomass within the production period.

## Feed provided and total waste

```{r fig-total-food-provision}
#| fig-cap: "Total food provided per unit biomass produced across feed groups (MD shown in purple, PD in green, AD in orange) and three digestibility levels (minimum, mean, and maximum)."

total_vals %>%
  filter(measure == "food_prov_stat") %>%
  mutate(mean_per_biom = set_units(mean_per_biom, "kg t_fish-1")) %>% 
  ggplot(aes(x = feed_type, y = drop_units(mean_per_biom), fill = feed_group)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = feed_pal) +
  labs(
    x = "Feed digesbility level",
    y = expression("Food provided (kg"~t[fish]^{-1}*")")
  ) +
  prettyplot_300() +
  theme(legend.position = "none")
```

```{r fig-total-waste}
#| fig-cap: "Total waste (excreted faeces + uneaten feed) per unit biomass produced across feed groups (MD shown in purple, PD in green, AD in orange) and three digestibility levels (minimum, mean, and maximum)."

total_vals %>%
  filter(measure == "total_waste_stat") %>%
  mutate(mean_per_biom = set_units(mean_per_biom, "kg t_fish-1")) %>% 
  ggplot(aes(x = feed_type, y = drop_units(mean_per_biom), fill = feed_group)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = feed_pal) +
  labs(
    x = "Feed digestibility level",
    y = expression("Total waste (kg"~t[fish]^{-1}*")")
  ) +
  prettyplot_300() +
  theme(legend.position = "none")
```

```{r fig-waste-components}
#| fig-cap: "Components of waste: excreted faeces (red) vs uneaten feed (blue) across feed groups (MD, PD, AD). Each panel shows a different digestibility level."

total_vals %>%
  filter(measure %in% c("total_excr_stat", "total_uneat_stat")) %>%
  mutate(mean_per_biom = set_units(mean_per_biom, "kg t_fish-1")) %>% 
  ggplot(aes(x = feed_group, y = mean_per_biom, fill = measure)) +
  geom_boxplot(alpha = 0.7, position = "dodge") +
  facet_wrap(~ feed_type, labeller = label_both) +
  scale_fill_manual(
    values = c("total_excr_stat" = "#c73535ff", "total_uneat_stat" = "#4682B4"),
    labels = c("Excreted", "Uneaten")
  ) +
  labs(
    x = "Feed",
    y = expression("Waste component (kg"~t[fish]^{-1}*")"),
    fill = "Waste type"
  ) +
  prettyplot_300() +
  theme(legend.position = "right")
```

```{r fig-prop-uneaten}
#| fig-cap: "Proportion of provided feed that is uneaten (lost to environment) across feed groups (MD shown in purple, PD in green, AD in orange) and digestibility levels (minimum, mean, and maximum). Points show median values while error bars show range from minimum to maximum."

total_props %>%
  filter(measure == "prop_uneat_stat") %>%
  group_by(feed_group, feed_type, measure) %>%
  summarise(
    median_val = median(mean_prop),
    min_val = min(mean_prop),
    max_val = max(mean_prop),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = feed_type, y = median_val, ymin = min_val, ymax = max_val, color = feed_group)) +
  geom_point(position = position_dodge(width = 0.35), size = 2.5) +
  geom_errorbar(position = position_dodge(width = 0.35), width = 0.25) +
  scale_color_manual(values = feed_pal) +
  scale_y_continuous(
    breaks = seq(0, 0.05, 0.0025),
    limits = c(0.0275, 0.045),
    labels = scales::percent_format()
  ) +
  labs(
    x = "Feed digestibility level",
    y = "Proportion of feed uneaten (%)"
  ) +
  prettyplot_300() +
  theme(legend.position = "none")
```

## Protein loss to the environment

```{r fig-protein-losses}
#| fig-cap: "Total protein losses (excreted + uneaten) per unit biomass produced across feed groups (MD shown in purple, PD in green, AD in orange) and three digestibility levels (minimum, mean, and maximum)."

total_vals %>%
  filter(measure == "total_P_lost_stat") %>%
  mutate(mean_per_biom = set_units(mean_per_biom, "kg t_fish-1")) %>% 
  ggplot(aes(x = feed_type, y = drop_units(mean_per_biom), fill = feed_group)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = feed_pal) +
  labs(
    x = "Feed digestibility levels",
    y = expression("Total protein lost (kg"~t[fish]^{-1}*")")
  ) +
  prettyplot_300() +
  theme(legend.position = "none")
```

```{r fig-protein-components}
#| fig-cap: "Components of protein loss: excreted protein (red) vs uneaten protein (blue) across feed groups (MD, PD, AD). Each panel shows a different digestibility level (minimum, mean, and maximum)."

total_vals %>%
  filter(measure %in% c("P_excr_stat", "P_uneat_stat")) %>%
  mutate(mean_per_biom = set_units(mean_per_biom, "kg t_fish-1") %>% drop_units()) %>% 
  ggplot(aes(x = feed_group, y = mean_per_biom, fill = measure)) +
  geom_boxplot(alpha = 0.7, position = "dodge") +
  facet_wrap(~ feed_type, labeller = label_both) +
  scale_fill_manual(
    values = c("P_excr_stat" = "#c73535ff", "P_uneat_stat" = "#4682B4"),
    labels = c("P excreted", "P uneaten")
  ) +
  labs(
    x = "Feed group",
    y = expression("Total protein lost (kg"~t[fish]^{-1}*")"),
    fill = "Loss type"
  ) +
  prettyplot_300() +
  theme(legend.position = "right")
```

## Waste composition

```{r fig-prop-excr-vs-total-waste}
#| fig-cap: "Proportion of total and protein waste that is lost through excreted faeces (vs uneaten feed) across feeds (MD shown in purple, PD in green, AD in orange) and digestibility levels. Points show median values while error bars show range from minimum to maximum."

total_props %>%
  filter(measure %in% c("prop_excr_stat", "prop_P_stat")) %>%
  group_by(feed_group, feed_type, measure) %>%
  summarise(
    median_val = median(mean_prop),
    min_val = min(mean_prop),
    max_val = max(mean_prop),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = measure, y = median_val, ymin = min_val, ymax = max_val, color = feed_group)) +
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_errorbar(position = position_dodge(width = 0.5), width = 0.3) +
  scale_color_manual(values = feed_pal) +
  scale_y_continuous(
    breaks = seq(0.6, 0.9, 0.05),
    limits = c(0.6, 0.9),
    labels = scales::percent_format()
  ) +
  scale_x_discrete(
    labels = c("Total", "Protein")
  ) +
  facet_wrap(~feed_type) +
  labs(
    x = "Feed digestibility level",
    y = "Proportion of waste from excretion (%)"
  ) +
  prettyplot_300() +
  theme(legend.position = "none")
```

```{r protein-composition-of-faeces}
#| code-summary: Get the protein composition of excreted faeces. These are consistent across farms (no impact of temperature) and time (no impact of fish size).

total_props %>% 
  filter(measure == "prop_comp_excr_P_stat") %>%
  group_by(feed_group, feed_type, measure) %>%
  summarise(
    median_val = median(mean_prop),
    .groups = "drop"
  )
```

## Comparison across measures

```{r fig-all-metrics-heatmap}
#| fig-cap: "Heatmap showing median values of all metrics across feed groups (MD, PD, AD) and digestibility levels (minimum, mean, and maximum). Values have been scaled between -1 and 1 within each metric; blue indicates lower values, red indicates higher values, and white indicates mid-range values."

total_vals %>%
  group_by(feed_group, feed_type, measure) %>%
  summarise(median_val = median(mean_per_biom), .groups = "drop") %>%
  group_by(measure) %>%
  mutate(scaled_val = rescale(drop_units(median_val), c(-1,1))) %>%
  ungroup() %>%
  mutate(
    measure_label = factor(
      measure,
      levels = c("food_prov_stat", "total_waste_stat", "total_excr_stat", "total_uneat_stat", "total_P_lost_stat", "P_excr_stat", "P_uneat_stat"),
      labels = c("Food provided", "Total waste", "Excreted faeces", "Uneaten feed", "Total P lost", "P excreted", "P uneaten"))
  ) %>%
  ggplot(aes(x = feed_group, y = measure_label, fill = scaled_val)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = sprintf("%.2f", scaled_val)), size = 3) +
  facet_wrap(~feed_type, nrow = 1) +
  scale_fill_gradient2(low = "#4682B4", mid = "white", high = "#c73535ff", midpoint = 0) +
  labs(
    x = "Feed digestibility level",
    y = "Metric",
    fill = "Scaled\nvalue"
  ) +
  prettyplot_300() +
  theme(
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

# Daily values

## Temporal patterns over a production year

```{r fig-daily-waste-total}
#| fig-cap: "Mean daily total waste production across all farms over one production cycle (548 days). Feed groups shown as: MD = marine-dominant (purple), PD = plant-dominant (green), AD = animal-dominant (orange). Digestibility levels shown by line types: solid = mean, dotted = minimum, dashed = maximum."

daily_vals.total_waste_stat %>%
  qread() %>% 
  group_by(prod_t, feed_group, feed_type) %>%
  summarise(
    mean_val = mean(mean),
    .groups = "drop"
  ) %>%
  mutate(mean_val = mean_val %>% set_units("kg") %>% drop_units()) %>% 
  ggplot(aes(x = prod_t, y = mean_val, color = feed_group, linetype = feed_type)) +
  geom_line(linewidth = 0.8) +
  scale_color_manual(values = feed_pal) +
  scale_linetype_manual(values = level_lines) +
  labs(
    x = "Production day",
    y = expression("Mean total waste (kg)"),
    color = "Feed group",
    linetype = "Digestibility"
  ) +
  prettyplot_300() +
  theme(legend.position = "right")
```

```{r fig-daily-waste-components}
#| fig-cap: "Daily patterns of waste components (excreted faeces in top panel, uneaten feed in bottom panel) for mean digestibility. Feed groups shown as: MD = marine-dominant (purple), PD = plant-dominant (green), AD = animal-dominant (orange)."

rbind(
  daily_vals.total_excr_stat %>% qread(),
  daily_vals.total_uneat_stat %>% qread()
) %>%
  filter(feed_type == "Mean") %>%
  group_by(prod_t, feed_group, measure) %>%
  summarise(
    mean_val = mean(mean),
    .groups = "drop"
  ) %>%
  mutate(
    measure_label = factor(measure,
                          levels = c("total_excr_stat", "total_uneat_stat"),
                          labels = c("Excreted", "Uneaten")),
    mean_val = mean_val %>% set_units("kg") %>% drop_units()
  ) %>%
  ggplot(aes(x = prod_t, y = mean_val, color = feed_group)) +
  geom_line(linewidth = 1) +
  facet_wrap(~ measure_label, ncol = 1, scales = "free_y") +
  scale_color_manual(values = feed_pal) +
  labs(
    x = "Production day",
    y = expression("Mean waste component (kg)"),
    color = "Feed group"
  ) +
  prettyplot_300() +
  theme(legend.position = "right")
```

```{r fig-daily-waste-digestibility-comparison}
#| fig-cap: "Effect of digestibility on daily waste production patterns within each feed group (MD, PD, AD shown in separate panels). Digestibility levels shown by color: Minimum, Mean, and Maximum."

daily_vals.total_waste_stat %>%
  qread() %>% 
  group_by(prod_t, feed_group, feed_type) %>%
  summarise(
    mean_val = mean(mean),
    .groups = "drop"
  ) %>%
  mutate(mean_val = mean_val %>% set_units("kg") %>% drop_units()) %>% 
  ggplot(aes(x = prod_t, y = mean_val, color = feed_type)) +
  geom_line(linewidth = 1) +
  facet_wrap(~ feed_group, ncol = 1) +
  scale_color_manual(values = level_pal) +
  labs(
    x = "Production day",
    y = expression("Mean total waste (kg)"),
    color = "Digestibility"
  ) +
  prettyplot_300() +
  theme(legend.position = "right")
```

```{r fig-daily-protein-losses}
#| fig-cap: "Mean daily protein losses over production year (365 days). Feed groups shown as: MD = marine-dominant (purple), PD = plant-dominant (green), AD = animal-dominant (orange). Digestibility levels shown by line types: solid = mean, dotted = minimum, dashed = maximum."

daily_vals.total_P_lost_stat %>%
  qread() %>% 
  group_by(prod_t, feed_group, feed_type) %>%
  summarise(
    mean_val = mean(mean),
    .groups = "drop"
  ) %>%
  mutate(mean_val = mean_val %>% set_units("kg") %>% drop_units()) %>% 
  ggplot(aes(x = prod_t, y = mean_val, color = feed_group, linetype = feed_type)) +
  geom_line(linewidth = 0.8) +
  scale_color_manual(values = feed_pal) +
  scale_linetype_manual(values = level_lines) +
  labs(
    x = "Production day",
    y = expression("Mean protein loss (kg)"),
    color = "Feed group",
    linetype = "Digestibility"
  ) +
  prettyplot_300()
```

```{r fig-daily-protein-components}
#| fig-cap: "Daily patterns of protein loss components (protein excreted in top panel, protein uneaten in bottom panel) for mean digestibility. Feed groups shown as: MD = marine-dominant (purple), PD = plant-dominant (green), AD = animal-dominant (orange)."

rbind(
  daily_vals.P_excr_stat %>% qread(),
  daily_vals.P_uneat_stat %>% qread()
 ) %>%
  filter(feed_type == "Mean") %>%
  group_by(prod_t, feed_group, measure) %>%
  summarise(
    mean_val = mean(mean),
    .groups = "drop"
  ) %>%
  mutate(
    measure_label = factor(measure,
                          levels = c("P_excr_stat", "P_uneat_stat"),
                          labels = c("P excreted", "P uneaten")),
    mean_val = mean_val %>% set_units("kg") %>% drop_units()
  ) %>%
  ggplot(aes(x = prod_t, y = mean_val, color = feed_group)) +
  geom_line(linewidth = 1) +
  facet_wrap(~ measure_label, ncol = 1, scales = "free_y") +
  scale_color_manual(values = feed_pal) +
  labs(
    x = "Production day",
    y = expression("Mean protein loss (kg)"),
    color = "Feed group"
  ) +
  prettyplot_300()
```

# Daily proportions

```{r fig-daily-prop-excr}
#| fig-cap: "Daily proportion of waste from excretion (vs uneaten feed) over production year (365 days). Feed groups shown as: MD = marine-dominant (purple), PD = plant-dominant (green), AD = animal-dominant (orange). Digestibility levels shown by line types: solid = mean, dotted = minimum, dashed = maximum."

daily_props %>%
  qread() %>% 
  filter(measure == "prop_excr_stat") %>%
  group_by(prod_t, feed_group, feed_type) %>%
  summarise(
    mean_val = mean(mean),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = prod_t, y = drop_units(mean_val), color = feed_group, linetype = feed_type)) +
  geom_line(linewidth = 0.8) +
  scale_color_manual(values = feed_pal) +
  scale_linetype_manual(values = level_lines) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    x = "Production day",
    y = "Proportion of waste from excretion (%)",
    color = "Feed group",
    linetype = "Digestibility"
  ) +
  prettyplot_300()
```

```{r fig-daily-prop-protein-excr}
#| fig-cap: "Daily proportion of protein lost via excretion (vs uneaten) over production year (365 days). Feed groups shown as: MD = marine-dominant (purple), PD = plant-dominant (green), AD = animal-dominant (orange). Digestibility levels shown by line types: solid = mean, dotted = minimum, dashed = maximum."

daily_props %>%
  qread() %>% 
  filter(measure == "prop_P_stat") %>%
  group_by(prod_t, feed_group, feed_type) %>%
  summarise(
    mean_val = mean(mean),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = prod_t, y = drop_units(mean_val), color = feed_group, linetype = feed_type)) +
  geom_line(linewidth = 0.8) +
  scale_color_manual(values = feed_pal) +
  scale_linetype_manual(values = level_lines) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    x = "Production day",
    y = "Proportion of protein lost via excretion (%)",
    color = "Feed group",
    linetype = "Digestibility"
  ) +
  prettyplot_300()
```

# Feed group comparison summary

## Pairwise comparisons of feed groups

```{r fig-feed-group-differences}
#| fig-cap: "Distribution of differences in waste metrics between feed groups relative to MD. Panels show different digestibility levels. Green boxes show PD vs MD comparison, orange boxes show AD vs MD comparison. Positive values indicate higher waste in the alternative feed group compared to MD."

total_vals %>%
  mutate(
    measure = factor(measure,
    levels = c(
      "food_prov_stat", "total_uneat_stat", "P_uneat_stat", 
      "total_waste_stat", "total_P_lost_stat", 
      "total_excr_stat", "P_excr_stat"
    ),
    labels = c(
      "Feed provided", "Total uneaten", "Protein uneaten", 
      "Total waste", "Total P lost", 
      "Total excreted", "P excreted"
      )
    )
  ) %>% 
  filter(measure %in% c("P excreted", "Total excreted", "Total P lost", "Total waste")) %>% 
  select(farm_ID, feed_group, mean_per_biom, feed_type, measure) %>%
  pivot_wider(names_from = c(feed_group, feed_type), values_from = mean_per_biom) %>%
  drop_na() %>%
  mutate(
    ADvsMD_Mean = AD_Mean - MD_Mean, 
    PDvsMD_Mean = PD_Mean - MD_Mean,
    ADvsMD_Maximum = AD_Maximum - MD_Maximum, 
    PDvsMD_Maximum = PD_Maximum - MD_Maximum,
    ADvsMD_Minimum = AD_Minimum - MD_Minimum, 
    PDvsMD_Minimum = PD_Minimum - MD_Minimum
  ) %>%
  select(farm_ID, measure, contains("vs")) %>%
  pivot_longer(cols = contains("vs"), names_to = c("comparison", "feed_type"), names_sep = "_", values_to = "difference") %>% 
  mutate(
    difference = difference %>% set_units("kg t_fish-1") %>% drop_units(),
    comparison = factor(comparison, levels = c("PDvsMD", "ADvsMD"), labels = c("PD vs MD", "AD vs MD"))
  ) %>% 
  ggplot(aes(x = feed_type, y = difference, fill = comparison)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = c("#1B9E77", "#D95F02")) +
  labs(
    x = "Feed digestibility",
    y = expression("Difference in total waste (kg"~t[fish]^{-1}*")")
  ) +
  facet_grid(rows = vars(measure)) +
  prettyplot_300() +
  theme(legend.position = "none")
```

```{r fig-feed-group-percent-differences-MD}
#| fig-cap: "Percent differences between feed groups relative to MD feed across three digestibility levels (minimum, mean, and maximum). Green bars show PD vs MD comparison, orange bars show AD vs MD comparison. Error bars show the range from minimum to maximum difference across all farms."

total_vals %>%
  select(farm_ID, feed_group, feed_type, measure, mean_per_biom) %>%
  pivot_wider(
    names_from = c(feed_group, feed_type), 
    values_from = mean_per_biom
  ) %>%
  drop_na() %>%
  mutate(
    `ADvsMD_Mean` = drop_units((AD_Mean - MD_Mean) / MD_Mean * 100),
    `PDvsMD_Mean` = drop_units((PD_Mean - MD_Mean) / MD_Mean * 100),
    `ADvsMD_Minimum` = drop_units((AD_Minimum - MD_Minimum) / MD_Minimum * 100),
    `PDvsMD_Minimum` = drop_units((PD_Minimum - MD_Minimum) / MD_Minimum * 100),
    `ADvsMD_Maximum` = drop_units((AD_Maximum - MD_Maximum) / MD_Maximum * 100),
    `PDvsMD_Maximum` = drop_units((PD_Maximum - MD_Maximum) / MD_Maximum * 100)
  ) %>%
  select(farm_ID, measure, contains("vs")) %>%
  pivot_longer(
    cols = contains("vs"), 
    names_to = c("comparison", "feed_type"), names_sep = "_",
    values_to = "percent_diff"
  ) %>%
  mutate(
    measure = factor(measure,
      levels = c(
        "food_prov_stat", 
        "total_waste_stat", "total_excr_stat", "total_uneat_stat", 
        "total_P_lost_stat", "P_excr_stat", "P_uneat_stat"
      ),
      labels = c(
        "Food provided", 
        "Total waste", "Excreted waste", "Uneaten feed", 
        "Total protein lost", "Protein excreted", "Protein uneaten"
      )),
    comparison = factor(comparison, levels = c("PDvsMD", "ADvsMD"))
  ) %>%
  group_by(comparison, feed_type, measure) %>% 
  reframe(
    mean = mean(percent_diff),
    min = min(percent_diff),
    max = max(percent_diff)
  ) %>% 
  filter(measure %in% c("Total waste", "Total protein lost", "Excreted waste", "Uneaten feed", "Protein excreted", "Protein uneaten")) %>% 
  ggplot(aes(x = feed_type, y = mean, ymin = min, ymax = max, fill = comparison)) +
  geom_col(position = position_dodge(), width = 0.9, alpha = 0.7, colour = "black") +
  geom_errorbar(position = position_dodge(0.9), width = 0.3, colour = "black") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  scale_fill_manual(values = c("ADvsMD" = "#D95F02", "PDvsMD" = "#1B9E77")) +
  facet_wrap(facets = vars(measure), ncol = 3) +
  labs(
    x = "Feed digestibility",
    y = "Percent difference from MD (%)",
    fill = "Comparison"
  ) +
  prettyplot_300()
```

```{r fig-feed-group-percent-differences-PD}
#| fig-cap: "Percent differences between feed groups relative to PD feed across three digestibility levels (minimum, mean, and maximum). Purple bars show MD vs PD comparison, orange bars show AD vs PD comparison. Error bars show the range from minimum to maximum difference across all farms."

total_vals %>%
  select(farm_ID, feed_group, feed_type, measure, mean_per_biom) %>%
  pivot_wider(
    names_from = c(feed_group, feed_type), 
    values_from = mean_per_biom
  ) %>%
  drop_na() %>%
  mutate(
    `ADvsPD_Mean` = drop_units((AD_Mean - PD_Mean) / PD_Mean * 100),
    `MDvsPD_Mean` = drop_units((MD_Mean - PD_Mean) / PD_Mean * 100),
    `ADvsPD_Minimum` = drop_units((AD_Minimum - PD_Minimum) / PD_Minimum * 100),
    `MDvsPD_Minimum` = drop_units((MD_Minimum - PD_Minimum) / PD_Minimum * 100),
    `ADvsPD_Maximum` = drop_units((AD_Maximum - PD_Maximum) / PD_Maximum * 100),
    `MDvsPD_Maximum` = drop_units((MD_Maximum - PD_Maximum) / PD_Maximum * 100)
  ) %>%
  select(farm_ID, measure, contains("vs")) %>%
  pivot_longer(
    cols = contains("vs"), 
    names_to = c("comparison", "feed_type"), names_sep = "_",
    values_to = "percent_diff"
  ) %>%
  mutate(
    measure = factor(measure,
      levels = c(
        "food_prov_stat", 
        "total_waste_stat", "total_excr_stat", "total_uneat_stat", 
        "total_P_lost_stat", "P_excr_stat", "P_uneat_stat"
      ),
      labels = c(
        "Food provided", 
        "Total waste", "Excreted waste", "Uneaten feed", 
        "Total protein lost", "Protein excreted", "Protein uneaten"
      )),
    comparison = factor(comparison, levels = c("MDvsPD", "ADvsPD"))
  ) %>%
  group_by(comparison, feed_type, measure) %>% 
  reframe(
    mean = mean(percent_diff),
    min = min(percent_diff),
    max = max(percent_diff)
  ) %>% 
  filter(measure %in% c("Total waste", "Total protein lost", "Excreted waste", "Uneaten feed", "Protein excreted", "Protein uneaten")) %>% 
  ggplot(aes(x = feed_type, y = mean, ymin = min, ymax = max, fill = comparison)) +
  geom_col(position = position_dodge(), width = 0.95, alpha = 0.7, colour = "black") +
  geom_errorbar(position = position_dodge(0.95), width = 0.3, colour = "black") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  scale_fill_manual(values = c("ADvsPD" = "#D95F02", "MDvsPD" = "#7570B3")) +
  facet_wrap(facets = vars(measure), ncol = 3) +
  labs(
    x = "Feed digestibility",
    y = "Percent difference from MD (%)",
    fill = "Comparison"
  ) +
  prettyplot_300()
```

# Farm-level variability

## Distribution of outcomes across farms

```{r fig-farm-variability-total-waste}
#| fig-cap: "Distribution of total waste across farms for each combination of feed group (MD = purple, PD = green, AD = orange) and digestibility level (rows show minimum, mean, and maximum). Histograms show the variability in outcomes across the 2697 farms."

total_vals %>%
  filter(measure == "total_waste_stat") %>%
  mutate(mean_per_biom = mean_per_biom %>% set_units("kg t_fish-1") %>% drop_units()) %>% 
  ggplot(aes(x = mean_per_biom, fill = feed_group)) +
  geom_histogram(bins = 50, alpha = 0.7, position = "identity") +
  facet_grid(feed_type ~ feed_group) +
  scale_fill_manual(values = feed_pal) +
  labs(
    x = expression("Total waste (kg"~t[fish]^{-1}*")"),
    y = "Number of farms"
  ) +
  prettyplot_300() +
  theme(legend.position = "none")
```

# Total values: Spatial results

```{r location-globals-farm}
#| code-summary: Get locations of farms for mapping

farm_locs <- file.path(output_farm_data_path, "farm_geometry.qs") %>% 
  qread() %>% 
  rename(farm_ID = farm_id) %>%
  mutate(
    longitude = st_coordinates(geometry)[, 1],
    latitude = st_coordinates(geometry)[, 2]
  ) %>% 
  mutate(
    country = as.factor(country),
    region = case_when(
      country %in% c("Canada", "United States of America") & longitude < -100 ~ "North America (WC)", 
      country %in% c("Canada", "United States of America") & longitude < -50 ~ "North America (EC)", 
      country %in% c("Ireland", "United Kingdom") ~ "UK & Ireland",
      country %in% c("Iceland", "Faroe Islands") ~ "Iceland & FI",
      country  == "Norway" ~ "Norway",
      country == "Australia" ~ "Australia",
      country == "Chile" ~ "Chile", 
      country == "Russian Federation" ~ NA
    ) %>% as.factor()
  )
```

## Global base (plant-dominant, mean digestibility)

```{r basemap-data}
#| code-summary: Get data for base map visualisation (plant-dominant, mean digestibility)

data_basemap <- total_vals %>% 
  filter(feed_type == "Mean", feed_group == "PD") %>% 
  merge(farm_locs, by = "farm_ID") %>% 
  mutate(mean = mean %>% set_units("kg") %>% drop_units())
```

```{r fig-map-total-P-lost-pd-mean}
#| fig-cap: Map showing spatial distribution of total protein losses for PD feed at mean digestibility.

p_bigmap_robinson +
  geom_sf(
    data = data_basemap %>% 
      filter(measure == "total_P_lost_stat") %>% 
      st_as_sf(crs = 4326) %>% 
      st_transform(crs = "+proj=robin"), 
    aes(color = mean, geometry = geometry), 
    size = 2
  ) +
  coord_sf() +
  scale_color_viridis_c(option = "turbo", breaks = seq(55, 90, 5), limits = c(55, 90)) + 
  guides(color = guide_colourbar(
    title = expression("Total excretion (kg/t salmon)"),
    direction = "horizontal", position = "bottom",
    label.position = "bottom", title.position = "top",
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 1,
    barheight = 1.5, barwidth = 20,
  )) +
  labs(y = "Latitude", x = "Longitude") +
  prettyplot_300() +
  theme(legend.position = "bottom") +
  remove_map_axes()
```

```{r fig-map-total-P-lost-pd-mean-patchwork}
#| fig-cap: Patchework map showing spatial distribution of total protein losses for PD feed at mean digestibility.

map1 <- p_bigmap_mercator +
  geom_sf(
    data = data_basemap %>% 
      filter(measure == "total_P_lost_stat"), 
    aes(color = mean_per_biom, geometry = geometry), 
    size = 2
  ) +
  coord_sf() +
  scale_color_viridis_c(option = "turbo", breaks = seq(55, 90, 5), limits = c(55, 90)) + 
  guides(color = guide_colourbar(
    title = expression("Total excretion (kg/t salmon)"),
    direction = "horizontal", 
    label.position = "bottom", title.position = "top",
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 1,
    barheight = 1.5, barwidth = 20,
  )) +
  labs(y = "Latitude", x = "Longitude") +
  prettyplot_300()

patchwork_mercator_1(map1)
patchwork_mercator_2(map1)
```

```{r fig-total-P-lost-farm-location}
#| fig-cap: Total feed lost vs various aspects of farm location (latitude, temperature).

tarstore <- here("targets/_targets_farmruns")
farm_temp_data <- targets::tar_read(farm_temp_data_chunked, store = tarstore) %>% 
  group_by(farm_ID) %>% 
  reframe(
    mean_temp = mean(temp_c),
    median_temp = median(temp_c),
    min_temp = min(temp_c),
    max_temp = max(temp_c)
  ) %>% 
  left_join(
    total_vals,
    by = "farm_ID"
  ) %>% 
  left_join(
    farm_locs,
    by = "farm_ID"
  ) %>%
  mutate(mean_per_biom = set_units(mean_per_biom, "kg t_fish-1") %>% drop_units()) %>% 
  filter(measure == "total_P_lost_stat" & feed_type == "Mean")
  
farm_temp_data %>% 
  filter(feed_group == "PD") %>% 
  ggplot(aes(x = mean_temp, y = mean_per_biom, colour = region)) +
  geom_point(shape = 1, size = 3.5) +
  scale_x_continuous(breaks = seq(0, 20, 2), limits = c(6, 18)) +
  scale_color_brewer(palette = "Dark2") + 
  prettyplot_300() +
  theme(legend.position = "right") +
  labs(
    x = expression("Mean farm temperature ("*degree*"C)"), 
    y = expression("Total protein lost (kg"~t[fish]^{-1}*")")
  ) +
  ggtitle("Total P lost per farm by mean farm temperature")

farm_temp_data %>% 
  ggplot(aes(x = mean_temp, y = mean_per_biom, colour = feed_group)) +
  geom_point(size = 2.5) +
  scale_color_manual(values = feed_pal) +
  scale_x_continuous(breaks = seq(0, 20, 2), limits = c(6, 18)) +
  scale_color_brewer(palette = "Dark2") + 
  prettyplot_300() +
  labs(
    x = expression("Mean farm temperature ("*degree*"C)"), 
    y = expression("Total protein lost (kg"~t[fish]^{-1}*")")
  ) +
  ggtitle("Total P lost per farm by mean farm temperature")

farm_temp_data %>% 
  distinct(farm_ID, mean_temp, latitude, country, region) %>% 
  ggplot(aes(x = latitude, y = mean_temp, colour = region)) +
  geom_point(size = 2.5) +
  scale_color_brewer(palette = "Dark2") + 
  prettyplot_300() +
  labs(
    x = expression("Farm latitude ("*degree*"E)"), 
    y = expression("Mean farm temperature ("*degree*"C)")
  ) +
  theme(legend.position = "right") +
  ggtitle("Mean farm temperature by farm latitude")

# farm_temp_data %>% 
#   ggplot(aes(x = median_temp, y = mean_per_biom, colour = feed_group)) +
#   geom_point() +
#   scale_color_manual(values = feed_pal) +
#   prettyplot_300() +
#   labs(
#     x = expression("Median farm temperature ("*degree*"C)"), 
#     y = expression("Total protein lost (kg"~t[fish]^{-1}*")")
#   ) +
#   ggtitle("Total P lost per farm by median farm temperature")

# farm_temp_data %>% 
#   ggplot(aes(x = min_temp, y = mean_per_biom, colour = feed_group)) +
#   geom_point() +
#   scale_color_manual(values = feed_pal) +
#   prettyplot_300() +
#   labs(
#     x = expression("Minimum farm temperature ("*degree*"C)"), 
#     y = expression("Total protein lost (kg"~t[fish]^{-1}*")")
#   ) +
#   ggtitle("Total P lost per farm by minimum farm temperature")

# farm_temp_data %>% 
#   ggplot(aes(x = max_temp, y = mean_per_biom, colour = feed_group)) +
#   geom_point() +
#   scale_color_manual(values = feed_pal) +
#   prettyplot_300() +
#   labs(
#     x = expression("Maximum farm temperature ("*degree*"C)"), 
#     y = expression("Total protein lost (kg"~t[fish]^{-1}*")")
#   ) +
#   ggtitle("Total P lost per farm by maximum farm temperature")

# farm_temp_data %>% 
#   ggplot(aes(x = latitude, y = mean_per_biom, colour = feed_group)) +
#   geom_point() +
#   scale_color_manual(values = feed_pal) +
#   prettyplot_300() +
#   labs(
#     x = expression("Farm latitude ("*degree*"E)"), 
#     y = expression("Total protein lost (kg"~t[fish]^{-1}*")")
#   ) +
#   ggtitle("Total P lost per farm by mean farm latitude")
```

## Chile as case study?

```{r fig-mapping-total-P-lost-chile}
#| fig-cap: Total protein lost across all feeds and digestibility levels in Chile.
#| fig-height: 9.5

data_formap <- total_vals %>% 
  filter(measure == "total_P_lost_stat") %>% 
  merge(farm_locs, by = "farm_ID") %>% 
  mutate(mean_per_biom = mean_per_biom %>% set_units("kg t_fish-1") %>% drop_units())

m1 <- p_bigmap_mercator +
  geom_sf(
    data = data_formap, 
    aes(color = mean_per_biom, geometry = geometry), 
    size = 2
  ) +
  coord_sf() +
  scale_color_viridis_c(option = "turbo", breaks = seq(0, 215, 20)) + 
  scale_x_continuous(breaks = seq(5, -95, -10)) +
  guides(color = guide_colourbar(
    title = expression(atop("Total protein", "lost (kg"~t[fish]^{-1}*")")),
    direction = "vertical", 
    label.position = "right", title.position = "top",
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 1,
    barheight = 20, barwidth = 1.5
  )) +
  facet_grid(
    rows = vars(feed_type), 
    cols = vars(feed_group)
    ) +
  labs(y = "Latitude", x = "Longitude") +
  prettyplot_300() +
  theme(legend.position = "right")

m1_insets <- get_insets_mercator(m1)
m1_insets[[4]]
```

# Daily values: Spatial results

In order to map daily results spatially, I have chosen to represent 3 time periods: the trough day (the day when outputs are lowest), the peak day (the day when outputs are highest) and the median day (the day where outputs are closest to the median value). 

```{r peak-trough-median days}
#| code-summary: Get production days where excretion (and basically everything else) are at their minimum, maximum, and median.

# pd = peak days
# td = trough days

dtw <- daily_vals.total_waste_stat %>% qread()

pds <- dtw$prod_t[dtw$mean == max(dtw$mean)]
tds <- dtw$prod_t[dtw$mean == min(dtw$mean)]
mds <- dtw$prod_t[(dtw$mean - median(dtw$mean)) == min(abs(dtw$mean - median(dtw$mean)))]

rm(dtw)

print(str_c("First peak day: ", min(pds)))
print(str_c("First trough day: ", min(tds)))
print(str_c("First median day: ", min(mds)))
```

```{r fig-daily-total-waste-stat-country}
#| fig-height: 6
#| fig-cap: 
#|  - "Total waste production on the median day across all countries, comparing feed groups (MD, PD, AD) and digestibility levels (minimum = light blue, mean = blue, maximum = dark blue). Each panel shows a different country."
#|  - "Total waste production on the peak day across all countries, comparing feed groups (MD, PD, AD) and digestibility levels (minimum = light blue, mean = blue, maximum = dark blue). Each panel shows a different country."
#|  - "Total waste production on the trough day across all countries, comparing feed groups (MD, PD, AD) and digestibility levels (minimum = light blue, mean = blue, maximum = dark blue). Each panel shows a different country."

daily_vals.total_waste_stat %>% 
  qread() %>% 
  filter(prod_t == min(mds)) %>% 
  mutate(mean = mean %>% set_units("kg") %>% drop_units()) %>% 
  left_join(farm_locs, by = "farm_ID") %>% 
  ggplot(aes(x = feed_group, y = mean, fill = feed_type)) +
  geom_boxplot() +
  scale_fill_manual(values = level_pal) +
  facet_wrap(facets = vars(country)) +
  prettyplot_300()

daily_vals.total_waste_stat %>% 
  qread() %>% 
  filter(prod_t == min(pds)) %>% 
  mutate(mean = mean %>% set_units("kg") %>% drop_units()) %>% 
  left_join(farm_locs, by = "farm_ID") %>% 
  ggplot(aes(x = feed_group, y = mean, fill = feed_type)) +
  geom_boxplot() +
  scale_fill_manual(values = level_pal) +
  facet_wrap(facets = vars(country)) +
  prettyplot_300()

daily_vals.total_waste_stat %>% 
  qread() %>% 
  filter(prod_t == min(tds)) %>% 
  mutate(mean = mean %>% set_units("kg") %>% drop_units()) %>% 
  left_join(farm_locs, by = "farm_ID") %>% 
  ggplot(aes(x = feed_group, y = mean, fill = feed_type)) +
  geom_boxplot() +
  scale_fill_manual(values = level_pal) +
  facet_wrap(facets = vars(country)) +
  prettyplot_300()
```

```{r fig-daily-total-P-lost-stat-median}
#| fig-height: 6
#| fig-cap: 
#|  - "Total protein lost on the median day across all countries, comparing feed groups (MD, PD, AD) and digestibility levels (minimum = light blue, mean = blue, maximum = dark blue). Each panel shows a different country."
#|  - "Total protein lost on the peak day across all countries, comparing feed groups (MD, PD, AD) and digestibility levels (minimum = light blue, mean = blue, maximum = dark blue). Each panel shows a different country."
#|  - "Total protein lost on the trough day across all countries, comparing feed groups (MD, PD, AD) and digestibility levels (minimum = light blue, mean = blue, maximum = dark blue). Each panel shows a different country."

daily_vals.total_P_lost_stat %>% 
  qread() %>% 
  filter(prod_t == min(mds)) %>% 
  mutate(mean = mean %>% set_units("kg") %>% drop_units()) %>% 
  left_join(farm_locs, by = "farm_ID") %>% 
  ggplot(aes(x = feed_group, y = mean, fill = feed_type)) +
  geom_boxplot() +
  scale_fill_manual(values = level_pal) +
  facet_wrap(facets = vars(country)) +
  prettyplot_300()

daily_vals.total_P_lost_stat %>% 
  qread() %>% 
  filter(prod_t == min(pds)) %>% 
  mutate(mean = mean %>% set_units("kg") %>% drop_units()) %>% 
  left_join(farm_locs, by = "farm_ID") %>% 
  ggplot(aes(x = feed_group, y = mean, fill = feed_type)) +
  geom_boxplot() +
  scale_fill_manual(values = level_pal) +
  facet_wrap(facets = vars(country)) +
  prettyplot_300()

daily_vals.total_P_lost_stat %>% 
  qread() %>% 
  filter(prod_t == min(tds)) %>% 
  mutate(mean = mean %>% set_units("kg") %>% drop_units()) %>% 
  left_join(farm_locs, by = "farm_ID") %>% 
  ggplot(aes(x = feed_group, y = mean, fill = feed_type)) +
  geom_boxplot() +
  scale_fill_manual(values = level_pal) +
  facet_wrap(facets = vars(country)) +
  prettyplot_300()
```

## Peak day: Northern Hemisphere

The peak/trough days are always on the same day of the production cycle, but on different actual days in the southern and northern hemispheres.

```{r peak-day-hemispheres}
#| code-summary: Get peak days in each hemisphere

pds_N <- daily_vals.total_P_lost_stat %>% 
  qread() %>% 
  filter(prod_t == min(pds)) %>% 
  left_join(farm_locs, by = "farm_ID") %>% 
  distinct(country, t, prod_t)

pds_S <- pds_N %>% 
  filter(country == "Chile") %>% 
  pull(t)

pds_N <- pds_N %>% 
  filter(country == "Canada") %>% 
  pull(t)
```

```{r}

data_formap <- daily_vals.total_P_lost_stat %>% 
  qread() %>% 
  filter(t %in% c(pds_N, pds_S)) %>% 
  mutate(mean = mean %>% set_units("kg") %>% drop_units()) %>% 
  select("farm_ID", "t", "prod_t", "feed_group", "feed_type", "mean") %>% 
  pivot_wider(names_from = feed_group, values_from = mean) %>% 
  mutate(
    MD_PD = MD - PD,
    AD_PD = AD - PD
  ) %>% 
  select(-c(AD, PD, MD)) %>% 
  pivot_longer(
    names_to = "feed_comparison",
    values_to = "diff", cols = c("AD_PD", "MD_PD")
  ) %>% 
  mutate(feed_comparison = factor(feed_comparison, levels = c("MD_PD", "AD_PD"))) %>% 
  left_join(farm_locs, by = "farm_ID")

range(data_formap$diff)

p_bigmap_robinson +
  geom_sf(
    data = data_formap %>% 
      st_as_sf(crs = 4326) %>% 
      st_transform(crs = "+proj=robin"), 
    aes(color = diff, geometry = geometry), 
    size = 2
  ) +
  coord_sf() +
  prettyplot_300() +
  scale_color_viridis_c(option = "turbo", breaks = seq(-500, 500, 150), limits = c(-450, 350)) + 
  facet_grid(rows = vars(t), cols = vars(feed_comparison)) +
  labs(y = "Latitude", x = "Longitude") +
  remove_map_axes() +
  guides(color = guide_colourbar(
    title = expression("Total protein loss (kg)"),
    direction = "horizontal", position = "bottom",
    label.position = "bottom", title.position = "top",
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 1,
    barheight = 1.5, barwidth = 20,
  ))

```
