---
title: "Run Atlantic salmon model"
format: html
editor: source
---

# Introduction

This document runs the aquaculture model for Atlantic salmon farms.

```{r setup, include=F}
#| warning: false
#| message: false

library(tidyverse)
library(arrow)
library(sf)
library(terra)
library(furrr)
library(future)
library(tictoc)
library(fs)
library(conflicted)
library(readxl)
library(units)
library(qs)
library(here)
library(targets)
conflicted::conflicts_prefer(dplyr::filter(), dplyr::select(), .quiet = T)

here("src") %>% 
  list.files(pattern = "\\.R$", full.names = TRUE) %>% 
  str_subset("map", negate = T) %>% 
  walk(source)

remove_unit("g_fish")
remove_unit("kg_fish")
remove_unit("t_fish")
install_unit("g_fish")
install_unit("kg_fish", "1000 g_fish")
install_unit("t_fish", "1000 kg_fish")
```

# Run farms

First, let's test the pipeline to make sure that the feed input data is being processed correctly.

```{r check-farmrun-pipeline}
#| code-summary: Check that the pipeline will run correctly

Sys.setenv(TAR_PROJECT = "project_farmruns")

tar_make(test_reference_feed)
tar_read(test_reference_feed) # if weight < 1000g then feeds aren't processing properly! Fix it!
```

The following code runs the main pipeline.

```{r run-farms-pipeline, eval=F}
#| code-summary: Run the main pipeline

Sys.setenv(TAR_PROJECT = "project_farmruns")

tar_validate()

# Run whole pipeline
tar_make(seconds_meta_append = 300)
```

Load objects from the targets pipeline and save them as .qs files, which are much quicker to load down the line. This takes a few minutes to gather all the targets.

```{r all-farms-results}
#| code-summary: Save pipeline objects as files
#| eval: false

Sys.setenv(TAR_PROJECT = "project_farmruns")

tar_load(farm_results)
qsave(farm_results, file.path(output_path, "total_ends.qs"))

tar_load(total_vals)
qsave(total_vals, file.path(data_analysis_path, "total_vals.qs"))

tar_load(total_props)
qsave(total_props, file.path(data_analysis_path, "total_props.qs"))
```

## Further processing - C and N

To get nitrogen:
- N (g) = protein (g)/6.25

To get carbon:
- From protein: N (mol) = (1/3.7) * C (mol)
- From lipid: assume that lipid is 75% carbon
- From carbohydrate: assume that carb is around 40% carbon

These are from Prasad, S. S., Cameron, D. D., Carter, C. G., Williamson, A., & Blanchard, J. L. (2025). Geometric stoichiometry model yields relevant insights for assessing nutrient-related environmental impacts of aquaculture. Conservation Physiology, 13(1). https://doi.org/10.1093/conphys/coaf066

```{r conversion-to-CN}
get_nitrogen <- function(P) {
  P/6.25
}

get_carbon <- function(P, L, C) {
  carbon_1 <- L * 0.75
  carbon_2 <- C * 0.41
  carbon_3 <- get_nitrogen(P=P)/14.007
  carbon_3 <- (carbon_3 * 3.7) * 12.011
  carbon_1 + carbon_2 + carbon_3
}

CN_vals <- total_vals %>% 
  select(farm_ID, feed_group, measure, feed_type, mean, mean_per_biom) %>% 
  filter(measure %in% c("total_P_lost_stat", "total_C_lost_stat", "total_L_lost_stat", "C_excr_stat", "L_excr_stat", "P_excr_stat")) %>% 
  pivot_wider(names_from = "measure", values_from = c(mean, mean_per_biom), names_sep = ".") %>% 
  mutate(
    mean.N_excr = get_nitrogen(mean.P_excr_stat),
    mean.C_excr = get_carbon(P = mean.P_excr_stat, L = mean.L_excr_stat, C = mean.C_excr_stat),
    mean_per_biom.N_excr = get_nitrogen(mean.P_excr_stat),
    mean_per_biom.C_excr = get_carbon(P = mean_per_biom.P_excr_stat, L = mean_per_biom.L_excr_stat, C = mean_per_biom.C_excr_stat),
    mean.N_total_lost = get_nitrogen(mean.total_P_lost_stat),
    mean.C_total_lost = get_carbon(L = mean.total_L_lost_stat, C = mean.total_C_lost_stat, P = mean.total_P_lost_stat),
    mean_per_biom.N_total_lost = get_nitrogen(mean_per_biom.total_P_lost_stat),
    mean_per_biom.C_total_lost = get_carbon(P = mean_per_biom.total_P_lost_stat, L = mean_per_biom.total_L_lost_stat, C = mean_per_biom.total_C_lost_stat)
  ) %>% 
  select(-contains("stat")) %>% 
  pivot_longer(
    cols = c(mean.N_excr, mean.C_excr, mean_per_biom.N_excr, mean_per_biom.C_excr, mean.N_total_lost, mean.C_total_lost, mean_per_biom.N_total_lost, mean_per_biom.C_total_lost),
    names_to = c(".value", "measure"),
    names_sep = "\\."
  ) %>% 
  mutate(
    nutrient = case_when(str_detect(measure, "C_") ~ "C", T ~ "N"),
    measure = str_remove(measure, "C_|N_") %>% as.factor(),
    nutrient = as.factor(nutrient)
  )

qsave(CN_vals, file.path(data_analysis_path, "total_CN_vals.qs"))
```

# Run sensitivities

```{r check-sensitivities-pipeline}
#| code-summary: Check that the pipeline will run correctly

Sys.setenv(TAR_PROJECT = "project_sensitivities")

tar_make(test_reference_feed)
tar_read(test_reference_feed) # if weight < 1000g then feeds aren't processing properly! Fix it!
```

```{r}
Sys.setenv(TAR_PROJECT = "project_sensitivities")

# Run whole pipeline
tar_make(seconds_meta_append = 300)
```

```{r sens-results}
tar_load(sens_results_pop)
tar_load(sens_results_spec)

sens <- rbind(sens_results_pop, sens_results_spec)
qsave(sens, file.path(output_sens_data_path, "sens_all.qs"))
```
