---
title: "Step 9: Data analysis"
subtitle: "Local to global salmon aquaculture waste modelling"
author: 
  - "Tormey Reimer"
date: last-modified
execute:
  eval: true
---

[← Previous: Run Atlantic salmon model](docs-working/08_run_atlantic_salmon_model.qmd)

# Overview {#sec-overview}

This document does some additional data analysis on the outputs before presenting numbers needed for the manuscript. In particular, we fit a series of mixed-effects models to determine how much of the variance can be attributed to fixed effects (different feeds), random effects (different farms) and residual effects (Monte-Carlo sampled individual fish).

```{r setup}
#| cache: false
#| include: false

library(tidyverse)
library(sf)
library(conflicted)
library(units)
library(qs)
library(here)
library(gt)
library(lme4)
library(MuMIn)

conflicted::conflicts_prefer(
  dplyr::filter(),
  dplyr::select(),
  .quiet = T
)
```

```{r source-scripts}
#| code-summary: "Source R scripts and set up custom units"
#| output: false

here("src") %>%
  list.files(pattern = "\\.R$", full.names = TRUE) %>%
  str_subset("pipeline", negate = T) %>%
  walk(source)

remove_unit("g_fish")
remove_unit("kg_fish")
remove_unit("t_fish")
install_unit("g_fish")
install_unit("kg_fish", "1000 g_fish")
install_unit("t_fish", "1000 kg_fish")
```

```{r}
#| label: location-globals-farm
#| code-summary: "Get locations of farms for mapping"

farm_locs <- file.path(input_farm_coords_path, "farm_static_data.qs") %>% 
  qread() %>% 
  rename(farm_ID = farm_id) %>%
  mutate(
    longitude = st_coordinates(geometry)[, 1],
    latitude = st_coordinates(geometry)[, 2]
  ) %>% 
  mutate(
    country = as.factor(country),
    region = case_when(
      country %in% c("Canada", "United States of America") & longitude < -100 ~ "North America (WC)", 
      country %in% c("Canada", "United States of America") & longitude < -50 ~ "North America (EC)", 
      country %in% c("Ireland", "United Kingdom") ~ "UK & Ireland",
      country %in% c("Iceland", "Faroe Islands") ~ "Iceland & FI",
      country %in% c("Norway", "Finland") ~ "Scandinavia",
      country == "Australia" ~ "Australia",
      country == "Chile" ~ "Chile", 
      country == "Russian Federation" ~ NA
    ) %>% as.factor()
  )
```

# Variance partitioning {#sec-variance-partitioning}

```{r}
#| label: get-farm-endsdata
#| code-summary: "Get all data"

fish_ends_nitrogen <- file.path(output_path, "model_run_outputs", "fish_ends_total_nitrogen.qs") %>% 
  qread() %>% 
  convert_feeds_to_levels()
fish_ends_carbon <- file.path(output_path, "model_run_outputs", "fish_ends_total_carbon.qs") %>% 
  qread() %>% 
  convert_feeds_to_levels()
```

```{r}
#| label: get-farm-ends-data
#| code-summary: "Get all data"

farm_ends_nitrogen <- file.path(output_path, "model_run_outputs", "farm_ends_total_nitrogen.qs") %>% 
  qread() %>% 
  convert_feeds_to_levels()
farm_ends_carbon <- file.path(output_path, "model_run_outputs", "farm_ends_total_carbon.qs") %>% 
  qread() %>% 
  convert_feeds_to_levels()

farm_ends_food_prov <- file.path(output_path, "model_run_outputs", "farm_ends_food_prov.qs") %>% 
  qread() %>% 
  convert_feeds_to_levels()
farm_ends_total_uneat <- file.path(output_path, "model_run_outputs", "farm_ends_total_uneat.qs") %>% 
  qread() %>% 
  convert_feeds_to_levels()
farm_ends_total_excr <- file.path(output_path, "model_run_outputs", "farm_ends_total_excr.qs") %>% 
  qread() %>% 
  convert_feeds_to_levels()
```

For total nitrogen and carbon, we use the total value per biomass rather than total value. This is because total value is highly dependent on farm population, which is implied rather than measured and gives far too much weight to the random effect of farm location.

```{r}
#| label: choose-model
#| code-summary: "Check if we should be accounting for all random slope effects"

set.seed(42)
fish_ends_nitrogen_subset <- fish_ends_nitrogen %>% 
  filter(farm_ID %in% sample(unique(fish_ends_nitrogen$farm_ID), 75)) %>% 
  mutate(farm_ID = droplevels(farm_ID)) %>% 
  group_by(farm_ID, feed, feed_level) %>% 
  slice_sample(n = 25) %>% 
  ungroup()

# model1 <- lmer(value_perbiom ~ feed/feed_level + (1 + feed/feed_level | farm_ID), 
#                data = fish_ends_nitrogen_subset, REML = FALSE,
#                control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))
# qsave(model1, file.path(data_analysis_path, "VP_model1_total_nitrogen.qs"))

# model2 <- lmer(value_perbiom ~ feed/feed_level + (1 + feed/feed_level || farm_ID),
#                 data = fish_ends_nitrogen_subset, REML = FALSE,
#                 control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))
# qsave(model2, file.path(data_analysis_path, "VP_model2_total_nitrogen.qs"))

# model3 <- lmer(value_perbiom ~ feed/feed_level + (1 + feed | farm_ID),
#                data = fish_ends_nitrogen_subset, REML = FALSE,
#                control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))
# qsave(model3, file.path(data_analysis_path, "VP_model3_total_nitrogen.qs"))

# model4 <- lmer(value_perbiom ~ feed/feed_level + (1 | farm_ID),
#                 data = fish_ends_nitrogen_subset, REML = FALSE,
#                 control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))
# qsave(model4, file.path(data_analysis_path, "VP_model4_total_nitrogen.qs"))

model1 <- file.path(data_analysis_path, "VP_model1_total_nitrogen.qs") %>% qread()
model2 <- file.path(data_analysis_path, "VP_model2_total_nitrogen.qs") %>% qread()
model3 <- file.path(data_analysis_path, "VP_model3_total_nitrogen.qs") %>% qread()
model4 <- file.path(data_analysis_path, "VP_model4_total_nitrogen.qs") %>% qread()

print("Singular fit models:")
isSingular(model1)
isSingular(model2)
isSingular(model3)
isSingular(model4)

print("Compare models 4 and 3")
anova(model4, model3)

print("Compare models 4 and 2")
anova(model4, model2)
```

Successively removing random effects shows that farm_ID is the only one that really matters wihtout overcomplicating the model structure. In the final model:

- Feed and nested feed_level effects are fixed effects
- Farm_ID is controlled as a random intercept
- The residual represents individual fish variation

```{r}
#| label: fit-model-value-nitrogen
#| code-summary: "Fit actual model for total_nitrogen"

VP_nitrogen <- file.path(data_analysis_path, "VP_modelfinal_total_nitrogen.qs")
if (!file.exists(VP_nitrogen)) {
  VP_nitrogen <- lmer(
    value_perbiom ~ feed/feed_level + (1 | farm_ID), 
    data = fish_ends_nitrogen, REML = TRUE,
    control = lmerControl(optimizer = "bobyqa",optCtrl = list(maxfun = 2e5))
  )
  qsave(VP_nitrogen, file.path(data_analysis_path, "VP_modelfinal_total_nitrogen.qs"))
} else {
  VP_nitrogen <- qread(VP_nitrogen)
}

# Random effects variance (location + residual)
vc <- as.data.frame(VarCorr(VP_nitrogen))
vc$pct <- vc$vcov / sum(vc$vcov) * 100
vc

# Fixed effects variance — how much do feed and variant explain?
# Use r.squaredGLMM for marginal (fixed, R2m) vs conditional (fixed + random, R2c) R²
r2 <- as.data.frame(r.squaredGLMM(VP_nitrogen))
r2$farm_ID <- r2$R2c - r2$R2m
r2$ind_fish <- 1 - r2$R2c
r2*100

rm(VP_nitrogen)
```

```{r}
#| label: fit-model-value-carbon
#| code-summary: "Fit actual model for total_carbon"

VP_carbon <- file.path(data_analysis_path, "VP_modelfinal_total_carbon.qs")
if (!file.exists(VP_carbon)) {
  VP_carbon <- lmer(
    value_perbiom ~ feed/feed_level + (1 | farm_ID), 
    data = fish_ends_carbon, 
    REML = TRUE,
    control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
  )
  qsave(VP_carbon, file.path(data_analysis_path, "VP_modelfinal_total_carbon.qs"))
} else {
  VP_carbon <- qread(VP_carbon)
}

# Random effects variance (location + residual)
vc <- as.data.frame(VarCorr(VP_carbon))
vc$pct <- vc$vcov / sum(vc$vcov) * 100
vc

# Fixed effects variance — how much do feed and variant explain?
# Use r.squaredGLMM for marginal (fixed, R2m) vs conditional (fixed + random, R2c) R²
r2 <- as.data.frame(r.squaredGLMM(VP_carbon))
r2$farm_ID <- r2$R2c - r2$R2m
r2$ind_fish <- 1 - r2$R2c
r2*100

rm(VP_carbon)
```

# Shifts in latitude {#sec-shifts-in-latitude}

From Julia: Can we say  With mean digestibolities there is X% more carbon/nitrogen loading per ton of fish produced per degree (either lat or temperature) for the reference MD feed. Shifts towards PD result in an equal/different? X % more C across all regions? with only X% more /less for PD (all based on mean digestibilty). Then N results follow slightly different  pattern with AD performing least best. (& MD best in terms of lowest N loadings). Then state whether the differences vary/d not vary across space.

```{r explore-latitude-shifts}
#| code-summary: "Explore percentage change in carbon loading per biomass relative to MD feed"
farm_ends_carbon %>%
  mutate(mean_value_perbiom = mean_value_perbiom %>% set_units("g g_fish-1") %>% set_units("kg t_fish-1")) %>% 
  filter(feed_level == "mean") %>% 
  select(-mean_value, -sd_value, -sd_per_biom, -n) %>% 
  pivot_wider(names_from = feed, values_from = mean_value_perbiom) %>% 
  mutate(
    PD_change = (marine_dominant_biomar-plant_dominant_biomar)/marine_dominant_biomar,
    AD_change = (marine_dominant_biomar-animal_dominant_biomar)/marine_dominant_biomar
  ) %>% 
  select(-contains("biomar")) %>% 
  pivot_longer(cols = contains("change"), names_to = "feed", values_to = "perc_change") %>% 
  ggplot(aes(x = farm_ID, y = perc_change, colour = feed)) + geom_point()
```

# Numbers for paper {#sec-numbers-for-paper}

This section prints numbers to be used in the final manuscript.

```{r}
#| label: farm-info-1
#| code-summary: "Get number of farms from static farm data"

n_farms <- file.path(input_farm_coords_path, "farm_static_data.qs") %>% 
  qread() %>% 
  rename(farm_ID = farm_id) %>% 
  pull(farm_ID) %>% unique() %>% length()

sprintf("Number of farms: %f", n_farms)
```

```{r}
#| label: farm-info-2
#| code-summary: "Get number of farms from temporal farm data"

farm_temp <- file.path(input_farm_coords_path, "farm_temporal_data.qs") %>% 
  qread() %>% 
  rename(farm_ID = farm_id) %>% 
  pull(farm_ID) %>% unique() %>% length()

sprintf("Number of farms: %f", n_farms)
```

```{r}
#| label: mean-harvest-size
#| code-summary: "Calculate range and mean harvest size"

harvest_size <- file.path(gendata_path, "harvest_sizes", "farm_harvest_sizes.qs") %>% 
  qread() %>% 
  mutate(weight = weight/1000)

range(harvest_size$weight)
mean(harvest_size$weight)
```

```{r total-waste}
#| code-summary: "Calculate total waste (uneaten + excreted) and compare across feeds"

total_waste <- rbind(
  farm_ends_total_uneat,
  farm_ends_total_excr
  ) %>% 
  mutate(
    sd_value = sd_value/mean_value,
    sd_per_biom = sd_per_biom/mean_value_perbiom
  ) %>% 
  group_by(feed, feed_level, farm_ID) %>% 
  reframe(
    mean_value = sum(mean_value),
    sd_value = sqrt(sum(sd_value^2)) * mean_value
  ) %>% 
  group_by(feed, feed_level) %>% 
  reframe(mean_value = mean(mean_value)) 
  
  
total_waste %>% 
  pivot_wider(names_from = "feed", values_from = "mean_value") %>% 
  mutate(
    PD_MD = 100*(plant_dominant_biomar - marine_dominant_biomar)/marine_dominant_biomar,
    AD_MD = 100*(animal_dominant_biomar - marine_dominant_biomar)/marine_dominant_biomar
  ) %>% 
  select(-animal_dominant_biomar, -marine_dominant_biomar, -plant_dominant_biomar) %>% 
  gt()
```

```{r}
#| label: proportion-uneaten
#| code-summary: "Calculate mean and SD of proportion of feed that is uneaten"

uneat <- rbind(
  farm_ends_food_prov,
  farm_ends_total_uneat
) %>% 
  select(-sd_per_biom, -mean_value, -sd_value) %>% 
  pivot_wider(names_from = measure, values_from = mean_value_perbiom) %>% 
  mutate(prop_uneat = total_uneat/food_prov)

100*mean(uneat$prop_uneat)
100*sd(uneat$prop_uneat)
```

@tbl-total-uneaten shows mean uneaten feed per biomass by feed type and level. @tbl-total-provided shows mean feed provided per unit biomass by feed type and level.

```{r}
#| label: tbl-total-uneaten
#| code-summary: "Summarise and tabulate total uneaten feed per biomass by feed and level"
#| tbl-cap: "Mean uneaten feed per unit biomass (kg per tonne of fish) by feed type and level, with propagated standard deviation."

farm_ends_total_uneat %>% 
  select(-mean_value, -sd_value) %>% 
  mutate(sd_per_biom = sd_per_biom/mean_value_perbiom) %>% 
  group_by(feed, feed_level, measure) %>% 
  reframe(
    mean_perbiom = mean(mean_value_perbiom),
    sd_perbiom = sqrt(sum(sd_per_biom^2)) * mean_perbiom
  ) %>% 
  mutate(
    mean_perbiom = mean_perbiom %>% set_units("g g_fish-1") %>% set_units("kg t_fish-1") %>% round(2),
    sd_perbiom = sd_perbiom %>% set_units("g g_fish-1") %>% set_units("kg t_fish-1") %>% round(2)
  ) %>% 
  gt()
```

```{r}
#| label: tbl-total-provided
#| code-summary: "Summarise and tabulate total feed provided per biomass by feed and level"
#| tbl-cap: "Mean feed provided per unit biomass (kg per tonne of fish) by feed type and level, with propagated standard deviation."

farm_ends_food_prov %>% 
  select(-mean_value, -sd_value) %>% 
  mutate(sd_per_biom = sd_per_biom/mean_value_perbiom) %>% 
  group_by(feed, feed_level, measure) %>% 
  reframe(
    mean_perbiom = mean(mean_value_perbiom),
    sd_perbiom = sqrt(sum(sd_per_biom^2)) * mean_perbiom
  ) %>% 
  mutate(
    mean_perbiom = mean_perbiom %>% set_units("g g_fish-1") %>% set_units("kg t_fish-1") %>% round(2),
    sd_perbiom = sd_perbiom %>% set_units("g g_fish-1") %>% set_units("kg t_fish-1") %>% round(2)
  ) %>% 
  gt()
```

```{r nitrogen-digestibility-range}
#| code-summary: "Calculate percentage difference in nitrogen loading between mean and max digestibility for AD feed"

tmp <- file.path(output_path, "model_run_outputs", "farm_ends_total_nitrogen.qs") %>%
  qread() %>% 
  convert_feeds_to_levels() %>% 
  mutate(
    sd_per_biom = set_units(sd_per_biom/mean_value_perbiom, "g g_fish-1"),
  ) %>% 
  group_by(feed, feed_level) %>% 
  reframe(
    mean_value_perbiom = mean(mean_value_perbiom) %>% set_units("g g_fish-1") %>% set_units("kg t_fish-1"),
    sd_per_biom = sqrt(sum(sd_per_biom^2)) %>% set_units("g g_fish-1") %>% set_units("kg t_fish-1")
    ) %>% 
      filter(feed == "animal_dominant_biomar")

(tmp$mean_value_perbiom[tmp$feed_level == "mean"] - tmp$mean_value_perbiom[tmp$feed_level == "max"]) / 
  tmp$mean_value_perbiom[tmp$feed_level == "mean"] * 100
```

[Next: Figures for paper →](docs-working/10_figures_for_paper.qmd)
