---
title: "Figures for manuscript"
author: 
  - "Tormey Reimer"
date: last-modified
execute:
  eval: true
---

# Introduction

```{r setup}
#| cache: false
#| include: false

library(tidyverse)
library(arrow)
library(sf)
library(terra)
library(magrittr)
library(furrr)
library(future)
library(tictoc)
library(fs)
library(conflicted)
library(readxl)
library(units)
library(qs)
library(here)
library(scales)
library(targets)
library(ggforce)

conflicted::conflicts_prefer(
  dplyr::filter(), 
  dplyr::select(), 
  scales::rescale(), 
  .quiet = T
)

here("src") %>% 
  list.files(pattern = "\\.R$", full.names = TRUE) %>% 
  walk(source)

remove_unit("g_fish")
remove_unit("kg_fish")
remove_unit("t_fish")
install_unit("g_fish")
install_unit("kg_fish", "1000 g_fish")
install_unit("t_fish", "1000 kg_fish")
```

# Get outputs

```{r location-globals-farm}
#| code-summary: Get locations of farms for mapping

farm_locs <- file.path(input_farm_coords_path, "farm_static_data.qs") %>% 
  qread() %>% 
  rename(farm_ID = farm_id) %>%
  mutate(
    longitude = st_coordinates(geometry)[, 1],
    latitude = st_coordinates(geometry)[, 2]
  ) %>% 
  mutate(
    country = as.factor(country),
    region = case_when(
      country %in% c("Canada", "United States of America") & longitude < -100 ~ "North America (WC)", 
      country %in% c("Canada", "United States of America") & longitude < -50 ~ "North America (EC)", 
      country %in% c("Ireland", "United Kingdom") ~ "UK & Ireland",
      country %in% c("Iceland", "Faroe Islands") ~ "Iceland & FI",
      country %in% c("Norway", "Finland") ~ "Scandinavia",
      country == "Australia" ~ "Australia",
      country == "Chile" ~ "Chile", 
      country == "Russian Federation" ~ NA
    ) %>% as.factor()
  )
```

## Total (summed) values

The total values describe selected output variables (measures) summed over the entire period of a salmon growing period (548 days). The data were generated from a suite of model runs which cover several dimensions:

- 2705 unique farms
- 3 different feed groups, representing different formulations of salmon feed: MD (marine-dominant), AD (animal-dominant) and PD (plant-dominant)
- 3 different feed types representing levels of digestibility of each of the three different feeds: mean digestibility, minimum digestibility, and maximum digestibility

The output values are expressed per biomass of salmon produced.

```{r total-values}
#| code-summary: Get total values data. These data give the total (summed) values of each response metric over 1 complete production period.

total_vals <- file.path(data_analysis_path, "total_vals.qs") %>% 
  qread()
total_props <- file.path(data_analysis_path, "total_props.qs") %>% 
  qread()
CN_vals <- file.path(data_analysis_path, "total_CN_vals.qs") %>% 
  qread()
```

The structure of these data are as follows:

```{r}
total_vals %>% 
  str()
```

```{r}
total_props %>% 
  str()
```

```{r}
CN_vals %>% 
  str()
```

# Main text figures

```{r fig-total-CN-lost}
#| fig-cap: Total carbon (a) and nitrogel (b) loading to the environment with different feeds and digestibility levels.

nutrient_labels <- data.frame(
  nutrient = levels(CN_vals$nutrient),
  label = letters[1:2]
)

CN_vals %>%
  filter(measure == "total_lost") %>% 
  mutate(mean_per_biom = mean_per_biom %>% 
    set_units("g g_fish-1") %>% 
    set_units("kg t_fish-1") %>% 
    drop_units()
  ) %>% 
  ggplot(aes(x = feed_type, y = mean_per_biom, fill = feed_group)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = feed_pal) +
  facet_grid(
    rows = vars(nutrient), 
    scales = "free"
  ) +
  geom_text(
    data = nutrient_labels, 
    aes(x = -Inf, y = Inf, label = label),
    hjust = -1, vjust = 1.5,
    size = 6, fontface = "bold",
    inherit.aes = FALSE
  ) +
  labs(
    x = "Feed digestibility",
    y = expression("Nutrient loading (kg"~t[fish]^{-1}*")")
  ) +
  prettyplot_600() +
  theme(
    legend.position = "none", 
    strip.text = element_blank(), 
    strip.background = element_blank()
  )
```

```{r fig-CN-ratio}
#| fig-cap: C:N ratio of excreted waste1
#| fig-height: 4.5

CN_ratio_vals <- CN_vals %>% 
  left_join(farm_locs, by = "farm_ID") %>% 
  filter(measure %in% c("total_lost", "excr")) %>% 
  select(farm_ID, feed_group, feed_type, measure, mean, nutrient, country, region) %>% 
  pivot_wider(names_from = c(nutrient, measure), values_from = mean, names_sep = ".") %>% 
  mutate(
    N.excr = N.excr/14.007,
    N.total_lost = N.total_lost/14.007,
    C.excr = C.excr/12.011,
    C.total_lost = C.total_lost/12.011,
    CN_ratio.excr = C.excr/N.excr,
    CN_ratio.total_lost = C.total_lost/N.total_lost
  ) %>% 
  group_by(feed_group, feed_type) %>% 
  reframe(
    CN_ratio_mean.excr = mean(CN_ratio.excr),
    CN_ratio_sd.excr = sd(CN_ratio.excr),
    CN_ratio_mean.total_lost = mean(CN_ratio.total_lost),
    CN_ratio_sd.total_lost = sd(CN_ratio.total_lost)
  ) %>% 
  pivot_longer(
    cols = contains("CN"),
    names_to = c(".value", "measure"),
    names_sep = "\\."
  ) %>% 
  mutate(measure = factor(measure, levels = c("excr", "total_lost")))

CN_ratio_vals %>% 
  ggplot(aes(x = feed_type, y = CN_ratio_mean, fill = feed_group, shape = measure)) +
  geom_point(size = 4, position = position_dodge(width = 0.9), colour = "black") +
  scale_fill_manual(values = feed_pal) +
  scale_shape_manual(values = c("excr" = 21, "total_lost" = 24)) +
  prettyplot_600() +
  # theme(legend.position = "right") +
  labs(x = "Feed digestibility", y = "C:N")

# CN_ratio_vals %>% 
#   select(-CN_ratio_sd) %>% 
#   pivot_wider(names_from = feed_type, values_from = CN_ratio_mean) %>% 
#   arrange(measure, feed_group)
```

```{r fig-map-total-CN-lost-PD}
#| fig-cap: Map showing spatial distribution of total protein losses for PD feed at mean digestibility.1
#| fig-height: 5

data_basemap_C <- CN_vals %>% 
  filter(feed_type == "Mean", feed_group == "PD") %>% 
  merge(farm_locs, by = "farm_ID") %>% 
  mutate(
    mean = set_units(mean, "g") %>% set_units("kg") %>% drop_units(),
    mean_per_biom = set_units(mean_per_biom, "g g_fish-1") %>% set_units("kg t_fish-1") %>% drop_units()
  ) %>% 
  filter(measure == "total_lost" & nutrient == "C") %>% 
  st_as_sf(crs = 4326) %>% 
  st_transform(crs = "+proj=robin")

p_bigmap_robinson +
# p_bigmap_robinson_boxes +
  geom_sf(
    data = data_basemap_C, 
    aes(color = mean_per_biom, geometry = geometry), 
    size = 2
  ) +
  coord_sf() +
  scale_color_viridis_c(option = "turbo", breaks = seq(75, 115, 10), limits = c(75, 116)) + 
  guides(color = guide_colourbar(
    title = expression(atop("C loading", paste("(kg"~t[fish]^{-1}*")"))),
    direction = "vertical", position = "right",
    label.position = "right", title.position = "top",
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 1,
    barwidth = 1.5, barheight = 12,
  )) +
  labs(y = "Latitude", x = "Longitude") +
  prettyplot_600() +
  theme(legend.position = "bottom") +
  remove_map_axes()

data_basemap_N <- CN_vals %>% 
  filter(feed_type == "Mean", feed_group == "PD") %>% 
  merge(farm_locs, by = "farm_ID") %>% 
  mutate(
    mean = set_units(mean, "g") %>% set_units("kg") %>% drop_units(),
    mean_per_biom = set_units(mean_per_biom, "g g_fish-1") %>% set_units("kg t_fish-1") %>% drop_units()
  ) %>% 
  filter(measure == "total_lost" & nutrient == "N") %>% 
  st_as_sf(crs = 4326) %>% 
  st_transform(crs = "+proj=robin")

p_bigmap_robinson +
# p_bigmap_robinson_boxes +
  geom_sf(
    data = data_basemap_N, 
    aes(color = mean_per_biom, geometry = geometry), 
    size = 2
  ) +
  coord_sf() +
  scale_color_viridis_c(option = "turbo", limits = c(9, 15)) + 
  guides(color = guide_colourbar(
    title = expression(atop("N loading", paste("(kg"~t[fish]^{-1}*")"))),
    direction = "vertical", position = "right",
    label.position = "right", title.position = "top",
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 1,
    barwidth = 1.5, barheight = 12,
  )) +
  labs(y = "Latitude", x = "Longitude") +
  prettyplot_600() +
  theme(legend.position = "bottom") +
  remove_map_axes()
```

```{r fig-mapping-total-CN-lost-chile}
#| fig-cap: Total N lost across all feeds and digestibility levels in Chile.1
#| fig-height: 8.5

data_formap <- CN_vals %>% 
  filter(measure == "total_lost" & nutrient == "N") %>% 
  merge(farm_locs, by = "farm_ID") %>% 
  mutate(
    mean = mean %>% set_units("g") %>% set_units("t") %>% drop_units(),
    feed_type = factor(
      feed_type,
      levels = c("Minimum", "Mean", "Maximum"),
      labels = c("Min. digestibility", "Mean digestibility", "Max. digestibility")
    )
  )

ggplot() +
  geom_sf(data = worldmap_mercator, fill = "white", color = "dimgray") +
  geom_sf(
    data = data_formap, 
    aes(color = mean/10^6, geometry = geometry), 
    size = 2
  ) +
  coord_sf(
    xlim = inset_boxes_sm[["CHI"]][["xlims"]], 
    ylim = inset_boxes_sm[["CHI"]][["ylims"]]
  ) +
  facet_grid(
    rows = vars(feed_type), 
    cols = vars(feed_group)
  ) +
  prettyplot_600() +
  scale_color_viridis_c(
    option = "turbo", 
    breaks = seq(0, 3.5e1, 5),
    limits = c(0, 3.55e1),
  ) + 
  scale_x_continuous(breaks = seq(5, -92.5, -10)) +
  guides(color = guide_colorbar()) +
  guides(color = guide_colourbar(
    title = expression(atop("N loading", paste("(t " %*% " 10"^6*")"))),
    direction = "vertical", position = "right",
    label.position = "right", title.position = "top",
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 1,
    barheight = 20, barwidth = 1.5
  )) +
  labs(y = "Latitude", x = "Longitude") +
  theme(legend.position = "right")

data_formap <- CN_vals %>% 
  filter(measure == "total_lost" & nutrient == "C") %>% 
  merge(farm_locs, by = "farm_ID") %>% 
  mutate(
    mean = mean %>% set_units("g") %>% set_units("t") %>% drop_units(),
    feed_type = factor(
      feed_type,
      levels = c("Minimum", "Mean", "Maximum"),
      labels = c("Min. digestibility", "Mean digestibility", "Max. digestibility")
    )
  )

ggplot() +
  geom_sf(data = worldmap_mercator, fill = "white", color = "dimgray") +
  geom_sf(
    data = data_formap, 
    aes(color = mean/10^6, geometry = geometry), 
    size = 2
  ) +
  coord_sf(
    xlim = inset_boxes_sm[["CHI"]][["xlims"]], 
    ylim = inset_boxes_sm[["CHI"]][["ylims"]]
  ) +
  facet_grid(
    rows = vars(feed_type), 
    cols = vars(feed_group)
  ) +
  prettyplot_600() +
  scale_color_viridis_c(
    option = "turbo", 
    breaks = seq(0, 2.6e2, 50),
    limits = c(0, 2.7e2),
  ) + 
  scale_x_continuous(breaks = seq(5, -92.5, -10)) +
  guides(color = guide_colorbar()) +
  guides(color = guide_colourbar(
    title = expression(atop("C loading", paste("(t " %*% " 10"^6*")"))),
    direction = "vertical", position = "right",
    label.position = "right", title.position = "top",
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 1,
    barheight = 20, barwidth = 1.5
  )) +
  labs(y = "Latitude", x = "Longitude") +
  theme(legend.position = "right")
```

# Supplementary figures

```{r fig-harvest-size}
#| fig-cap: Harvest sizes in the assigned farms.

harvest_size <- file.path(gendata_path, "harvest_sizes", "farm_harvest_sizes.qs") %>% 
  qread()

harvest_size %>% 
  mutate(weight = weight %>% set_units("g") %>% set_units("kg") %>% drop_units()) %>% 
  ggplot(aes(x = weight)) +
  geom_histogram(binwidth = 0.2, colour = "black", fill = "salmon", alpha = 0.5) +
  scale_x_continuous(breaks = seq(0,6,1), limits = c(1,5.5)) +
  prettyplot_600() +
  labs(x = "Harvest weight (kg)", y = "Frequency")
```

```{r fig-sensitivity}
sens <- file.path(output_sens_data_path, "sens_all.qs") %>% 
  qread()

param_names <- tibble::tribble(
  ~name,        ~lab,
  "alpha",      bquote(alpha),
  "epsprot",    bquote(epsilon["P"]),
  "epslip",     bquote(epsilon["L"]),
  "epscarb",    bquote(epsilon["C"]),
  "epsO2",      bquote(epsilon["O"[2]]),
  "pk",         bquote("pk"),
  "k0",         bquote("k"[0]),
  "m",          bquote("m"),
  "n",          bquote("n"),
  "betac",      bquote(beta*"C"),
  "Tma",        bquote("T"["max"]),
  "Toa",        bquote("T"["opt"]),
  "Taa",        bquote("T"["min"]),
  "omega",      bquote(omega),
  "a",          bquote("a"),
  "k",          bquote("k"),
  "eff",        bquote("eff"),
  "meanW",      bquote(bar("W")),
  "deltaW",     bquote(Delta*"W"), 
  "meanImax",   bquote(bar("I"["max"])),
  "deltaImax",  bquote(Delta*"I"["max"]),
  "overFmean",  bquote(bar("F")),
  "overFdelta", bquote(Delta*"F"),
  "mortmyt",    bquote("mort")
)

sens %>% 
  # mutate(adj_param = factor(
  #   adj_param, levels = c("alpha", "epsprot", "epslip", "epscarb", "epsO2", "pk", "k0", "m", "n", "betac", "Tma", "Toa", "Taa", "omega", "a", "k", "eff", "meanW", "deltaW", "meanImax", "deltaImax", "overFmean", "overFdelta", "mortmyt", "nruns") %>% rev()
  # )) %>% 
  filter(measure %in% c("weight_stat", "total_excr_stat") & adj_param != "nruns") %>% 
  mutate(measure = factor(measure, levels = c("weight_stat", "total_excr_stat"), labels = c("Ind. weight", "Total excreted"))) %>% 
  ggplot(aes(x = adj_param, y = mean_sens, ymin = mean_sens-sd_sens, ymax = mean_sens+sd_sens)) +
  geom_col(colour = "black", fill = "salmon", alpha = 0.5) +
  geom_errorbar(width = 0.5) +
  coord_flip() +
  scale_x_discrete(limits = rev(param_names$name), labels = rev(param_names$lab)) +
  facet_wrap(~measure) +
  prettyplot_600() +
  theme(axis.title.x = element_blank())
```