---
title: "Step 10: Figures for manuscript"
subtitle: "Local to global salmon aquaculture waste modelling"
author: 
  - "Tormey Reimer"
date: last-modified
execute:
  eval: true
---

[← Previous: Data analysis](09_data_analysis.qmd)

# Overview {#sec-overview}

This script generates all the figures (main and supplementary) used in the final manuscript.

```{r}
#| label: setup
#| cache: false
#| include: false

library(tidyverse)
library(arrow)
library(sf)
library(terra)
library(magrittr)
library(furrr)
library(future)
library(tictoc)
library(fs)
library(conflicted)
library(readxl)
library(units)
library(qs)
library(here)
library(scales)
library(targets)
library(ggforce)
library(cowplot)

conflicted::conflicts_prefer(
  dplyr::filter(), 
  dplyr::select(), 
  scales::rescale(), 
  .quiet = T
)

here("src") %>% 
  list.files(pattern = "\\.R$", full.names = TRUE) %>% 
  str_subset("pipeline", negate = T) %>% 
  walk(source)

remove_unit("g_fish")
remove_unit("kg_fish")
remove_unit("t_fish")
install_unit("g_fish")
install_unit("kg_fish", "1000 g_fish")
install_unit("t_fish", "1000 kg_fish")
```

# Get data {#sec-get-data}

```{r}
#| label: location-globals-farm
#| code-summary: "Get locations of farms for mapping"

farm_locs <- file.path(input_farm_coords_path, "farm_static_data.qs") %>% 
  qread() %>% 
  rename(farm_ID = farm_id) %>%
  mutate(
    longitude = st_coordinates(geometry)[, 1],
    latitude = st_coordinates(geometry)[, 2]
  ) %>% 
  mutate(
    country = as.factor(country),
    region = case_when(
      country %in% c("Canada", "United States of America") & longitude < -100 ~ "North America (WC)", 
      country %in% c("Canada", "United States of America") & longitude < -50 ~ "North America (EC)", 
      country %in% c("Ireland", "United Kingdom") ~ "UK & Ireland",
      country %in% c("Iceland", "Faroe Islands") ~ "Iceland & FI",
      country %in% c("Norway", "Finland") ~ "Scandinavia",
      country == "Australia" ~ "Australia",
      country == "Chile" ~ "Chile", 
      country == "Russian Federation" ~ NA
    ) %>% as.factor()
  )
```

The total values describe selected output variables (measures) summed over the entire period of a salmon growing period (548 days). The data were generated from a suite of model runs which cover several dimensions:

- 2750 unique farms
- 3 different feed groups, representing different formulations of salmon feed: MD (marine-dominant), AD (animal-dominant) and PD (plant-dominant)
- 3 different feed types representing levels of digestibility of each of the three different feeds: mean digestibility, minimum digestibility, and maximum digestibility

```{r}
#| label: total-values
#| code-summary: "Get farm ends data"

farm_ends_carbon <- file.path(output_path, "model_run_outputs", "farm_ends_total_carbon.qs") %>% 
  qread() %>% 
  mutate(
    mean_value = set_units(mean_value, "g"),
    sd_value = set_units(sd_value, "g"),
    mean_value_perbiom = set_units(mean_value_perbiom, "g g_fish-1"),
    sd_per_biom = set_units(sd_per_biom, "g g_fish-1"),
  )

farm_ends_nitrogen <- file.path(output_path, "model_run_outputs", "farm_ends_total_nitrogen.qs") %>% 
  qread() %>% 
  mutate(
    mean_value = set_units(mean_value, "g"),
    sd_value = set_units(sd_value, "g"),
    mean_value_perbiom = set_units(mean_value_perbiom, "g g_fish-1"),
    sd_per_biom = set_units(sd_per_biom, "g g_fish-1"),
  )

farm_ends_carbon_uneat <- file.path(output_path, "model_run_outputs", "farm_ends_carbon_uneat.qs") %>% qread()
farm_ends_carbon_excr <- file.path(output_path, "model_run_outputs", "farm_ends_carbon_excr.qs") %>% qread()
farm_ends_nitrogen_uneat <- file.path(output_path, "model_run_outputs", "farm_ends_nitrogen_uneat.qs") %>% qread()
farm_ends_nitrogen_excr <- file.path(output_path, "model_run_outputs", "farm_ends_nitrogen_excr.qs") %>% qread()
```

# Main text figures {#sec-main-text-figures}

## Total C and N to environment {#sec-total-cn}

```{r}
#| label: total-CN-data
#| code-summary: "Prepare combined C and N data for plotting"

farm_ends_CN <- rbind(
  farm_ends_carbon %>% mutate(nutrient = as.factor("C")),
  farm_ends_nitrogen %>% mutate(nutrient = as.factor("N"))
) %>% 
  select(-mean_value, -sd_value, -n) %>% 
  mutate(
    mean_value_perbiom = mean_value_perbiom %>% set_units("g g_fish-1") %>% set_units("kg t_fish-1") %>% drop_units(),
    sd_per_biom = mean_value_perbiom %>% set_units("g g_fish-1") %>% set_units("kg t_fish-1") %>% drop_units()
  ) %>% 
  convert_feeds_to_levels()

nutrient_labels <- data.frame(
  nutrient = levels(farm_ends_CN$nutrient),
  label = c("Carbon", "Nitrogen")
)
```

```{r}
#| label: fig-total-CN-lost
#| code-summary: "Plot total C and N loading to environment by feed and digestibility"
#| fig-cap: "Total carbon (a) and nitrogen (b) loading to the environment with different feeds and digestibility levels."

farm_ends_CN %>%
  mutate(
    feed = factor(feed, levels = c("marine_dominant_biomar", "plant_dominant_biomar", "animal_dominant_biomar")),
    feed_level = factor(feed_level, levels = c("min", "mean", "max"), labels = c("Minimum", "Mean", "Maximum"))
  ) %>% 
  ggplot(aes(x = feed_level, y = mean_value_perbiom, fill = feed)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = feed_pal) +
  facet_grid(
    rows = vars(nutrient), 
    scales = "free"
  ) +
  geom_text(
    data = nutrient_labels, 
    aes(x = -Inf, y = Inf, label = label),
    hjust = -0.1, vjust = 1,
    size = 5.5, family = "serif",
    inherit.aes = FALSE
  ) +
  labs(
    x = "Feed digestibility",
    y = expression("Nutrient loading (kg"~t[fish]^{-1}*")")
  ) +
  prettyplot_600() +
  theme(
    legend.position = "none", 
    strip.text = element_blank(), 
    strip.background = element_blank()
  )
```

## C:N ratio to environment {#sec-cn-ratio}

```{r}
#| label: ratio-data
#| code-summary: "Calculate C:N ratios for excreted and uneaten waste fractions"

farm_ends_CN_ratio <- rbind(
  farm_ends_carbon_uneat,
  farm_ends_carbon_excr,
  farm_ends_nitrogen_uneat,
  farm_ends_nitrogen_excr
) %>% 
  select(-c(mean_value, sd_value, sd_per_biom, n)) %>% 
  convert_feeds_to_levels() %>% 
  mutate(
    nutrient = case_when(str_detect(measure, "nitrogen") ~ "N", T ~ "C") %>% as.factor(),
    measure = str_remove_all(measure, "nitrogen_|carbon_"),
    mean_value_perbiom = mean_value_perbiom %>% set_units("g g_fish-1") %>% set_units("kg t_fish-1") %>% drop_units()
  )

farm_ends_CN_ratio <- rbind(
  farm_ends_CN_ratio,
  farm_ends_CN_ratio %>% 
    pivot_wider(names_from = measure, values_from = mean_value_perbiom) %>% 
    mutate(
      measure = "total",
      mean_value_perbiom = uneat + excr
    ) %>% 
    select(-c(uneat, excr))
) %>% 
  pivot_wider(names_from = nutrient, values_from = mean_value_perbiom) %>% 
  mutate(
    C = C/12.011,
    N = N/14.007,
    CN = C/N,
    measure = as.factor(measure)
  ) %>% 
  select(-c(C, N)) 
  
farm_ends_CN_ratio_means <- farm_ends_CN_ratio %>% 
  group_by(feed, feed_level, measure) %>% 
  reframe(
    CN_mean = mean(CN),
    CN_min = min(CN),
    CN_max = max(CN)
  )
```

```{r}
#| label: fig-CN-ratio
#| code-summary: "Plot C:N ratio of excreted, uneaten, and total waste"
#| fig-cap: "C:N ratio of total waste by feed type and digestibility level."
#| fig-height: 4.5

farm_ends_CN_ratio_means %>% 
  mutate(
    feed = factor(feed, levels = c("marine_dominant_biomar", "plant_dominant_biomar", "animal_dominant_biomar")),
    feed_level = factor(feed_level, levels = c("min", "mean", "max"), labels = c("Minimum", "Mean", "Maximum")),
    measure = factor(measure, levels = c("excr", "uneat", "total"), labels = c("Excreted faeces", "Uneaten Feed", "Total"))
  ) %>% 
  ggplot(aes(
    x = feed_level, 
    y = CN_mean, 
    ymin = CN_min,
    ymax = CN_max,
    fill = feed,
    shape = measure
  )) +
  geom_errorbar(linewidth = 1, position = position_dodge(width = 0.75), width = 0) +
  geom_point(size = 4, position = position_dodge(width = 0.75), colour = "black") +
  scale_fill_manual(values = feed_pal) +
  scale_shape_manual(values = c(
    "Excreted faeces" = 21, 
    "Total" = 24, 
    "Uneaten Feed" = 22
  )) +
  facet_wrap(~measure) +
  prettyplot_600() +
  # theme(legend.position = "right") +
  scale_y_continuous(breaks = 1:15, limits = c(6,13.25)) +
  labs(x = "Feed digestibility", y = "C:N")
```

## Global loading of C and N {#sec-global-loading}

```{r}
#| label: global-loading-data
#| code-summary: "Prepare global C and N loading data for mapping"

global_data <- rbind(
  farm_ends_carbon %>% mutate(nutrient = as.factor("C")),
  farm_ends_nitrogen %>% mutate(nutrient = as.factor("N"))
) %>% 
  select(-mean_value, -sd_per_biom, -sd_value, -n) %>% 
  mutate(mean_value_perbiom = mean_value_perbiom %>% set_units("g g_fish-1") %>% set_units("kg t_fish-1") %>% drop_units()) %>% 
  convert_feeds_to_levels() %>% 
  filter(feed_level == "mean", feed == "plant_dominant_biomar") %>% 
  merge(farm_locs, by = "farm_ID") %>% 
  st_as_sf(crs = 4326) %>% 
  st_transform(crs = "+proj=robin") %>% 
  select(farm_ID, feed, measure, mean_value_perbiom, nutrient, feed_level, iso3c, fao_code, country, region)

Cr <- range(global_data$mean_value[global_data$nutrient == "C"])
Nr <- range(global_data$mean_value[global_data$nutrient == "N"])

print(
  sprintf("Range of C: %f - %f Mt, range of N: %f - %f Mt", Cr[1], Cr[2], Nr[1], Nr[2])
)
```

```{r}
#| label: fig-map-total-CN-lost-PD
#| code-summary: "Map global C and N loading for plant-dominant feed at mean digestibility"
#| fig-cap: "Map showing global distribution of total nitrogen and carbon loading for PD feed at mean digestibility."
#| fig-height: 7.5

p1 <- p_bigmap_robinson +
  geom_sf(
    data = filter(global_data, nutrient == "C"), 
    aes(color = mean_value_perbiom, geometry = geometry), 
    size = 2
  ) +
  coord_sf() +
  scale_color_viridis_c(option = "turbo", breaks = seq(60, 90, 5), limits = c(60, 92)
  ) + 
  guides(color = guide_colourbar(
    title = expression(atop("C loading", paste("(kg"~t[fish]^{-1}*")"))),
    direction = "vertical", position = "right",
    label.position = "right", title.position = "top",
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 1,
    barwidth = 1.5, barheight = 15,
  )) +
  annotate(
    "text", label = "Carbon",
    x = -Inf, y = Inf, 
    vjust = 1, hjust = 0,
    size = 5.5, family = "serif",
    inherit.aes = FALSE
    ) +
  labs(y = "Latitude", x = "Longitude") +
  prettyplot_600() +
  theme(legend.position = "bottom") +
  remove_map_axes()

p2 <- p_bigmap_robinson +
  geom_sf(
    data = filter(global_data, nutrient == "N"), 
    aes(color = mean_value_perbiom, geometry = geometry), 
    size = 2
  ) +
  coord_sf() +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  scale_color_viridis_c(option = "turbo", breaks = seq(7, 12, 1), limits = c(7, 12)) + 
  guides(color = guide_colourbar(
    title = expression(atop("N loading", paste("(kg"~t[fish]^{-1}*")"))),
    direction = "vertical", position = "right",
    label.position = "right", title.position = "top",
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 1,
    barwidth = 1.5, barheight = 12,
  )) +
  annotate(
    "text", label = "Nitrogen",
    x = -Inf, y = Inf, 
    vjust = 1, hjust = 0,
    size = 5.5, family = "serif",
    inherit.aes = FALSE
    ) +
  labs(y = "Latitude", x = "Longitude") +
  prettyplot_600() +
  theme(legend.position = "bottom") +
  remove_map_axes()

plot_grid(
  p1, p2,
  ncol = 1,
  align = "v"
)
```

## Chile loading of C and N {#sec-chile-loading}

```{r}
#| label: chile-loading-data
#| code-summary: "Prepare Chile-specific C and N loading data for mapping"

chile_data <- rbind(
  farm_ends_carbon %>% mutate(nutrient = as.factor("C")),
  farm_ends_nitrogen %>% mutate(nutrient = as.factor("N"))
) %>% 
  select(-mean_value_perbiom, -sd_per_biom, -sd_value, -n) %>% 
  convert_feeds_to_levels() %>% 
  mutate(
    mean_value = mean_value %>% set_units("g") %>% set_units("Mt") %>% drop_units(),
    feed_level = factor(
      feed_level,
      levels = c("min", "mean", "max"),
      labels = c("Min. digestibility", "Mean digestibility", "Max. digestibility")
    ),
    feed = factor(
      feed,
      levels = c("marine_dominant_biomar", "plant_dominant_biomar", "animal_dominant_biomar"),
      labels = c("MD", "PD", "AD")
    )
  ) %>% 
  merge(farm_locs, by = "farm_ID") %>% 
  st_as_sf(crs = 4326) %>% 
  st_transform(crs = "+proj=robin") %>% 
  select(farm_ID, feed, measure, mean_value, nutrient, feed_level, iso3c, fao_code, country, region)

Cr <- range(chile_data$mean_value[chile_data$nutrient == "C"])
Nr <- range(chile_data$mean_value[chile_data$nutrient == "N"])
print(
  sprintf("Range of C: %f - %f Mt, range of N: %f - %f Mt", Cr[1], Cr[2], Nr[1], Nr[2])
)
```

```{r}
#| label: fig-mapping-total-CN-lost-chile
#| code-summary: "Map total C and N loading across all feeds and digestibility levels in Chile"
#| fig-cap: "Total N lost across all feeds and digestibility levels in Chile."
#| fig-height: 8.5

p1 <- ggplot() +
  geom_sf(data = worldmap_mercator, fill = "white", color = "dimgray") +
  geom_sf(
    data = filter(chile_data, nutrient == "C"), 
    aes(color = mean_value, geometry = geometry), 
    size = 2
  ) +
  coord_sf(
    xlim = inset_boxes_sm[["CHI"]][["xlims"]], 
    ylim = inset_boxes_sm[["CHI"]][["ylims"]]
  ) +
  facet_grid(
    rows = vars(feed_level), 
    cols = vars(feed)
  ) +
  prettyplot_600() +
  scale_color_viridis_c(
    option = "turbo", 
    breaks = seq(0, 300, 50),
    limits = c(0, 285),
  ) + 
  scale_x_continuous(breaks = seq(5, -92.5, -10)) +
  guides(color = guide_colorbar()) +
  guides(color = guide_colourbar(
    title = expression(atop("C loading", paste("(t " %*% " 10"^6*")"))),
    direction = "vertical", position = "right",
    label.position = "right", title.position = "top",
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 1,
    barheight = 20, barwidth = 1.5
  )) +
  labs(y = "Latitude", x = "Longitude") +
  theme(legend.position = "right")

p2 <- ggplot() +
  geom_sf(data = worldmap_mercator, fill = "white", color = "dimgray") +
  geom_sf(
    data = filter(chile_data, nutrient == "N"), 
    aes(color = mean_value, geometry = geometry), 
    size = 2
  ) +
  coord_sf(
    xlim = inset_boxes_sm[["CHI"]][["xlims"]], 
    ylim = inset_boxes_sm[["CHI"]][["ylims"]]
  ) +
  facet_grid(
    rows = vars(feed_level), 
    cols = vars(feed)
  ) +
  prettyplot_600() +
  scale_color_viridis_c(
    option = "turbo", 
    breaks = seq(0, 45, 5),
    limits = c(0, 41.5),
  ) + 
  scale_x_continuous(breaks = seq(5, -92.5, -10)) +
  guides(color = guide_colorbar()) +
  guides(color = guide_colourbar(
    title = expression(atop("N loading", paste("(t " %*% " 10"^6*")"))),
    direction = "vertical", position = "right",
    label.position = "right", title.position = "top",
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 1,
    barheight = 20, barwidth = 1.5
  )) +
  labs(y = "Latitude", x = "Longitude") +
  theme(legend.position = "right")

p1
p2
```

# Supplementary figures {#sec-supplementary}

## Harvest sizes {#sec-harvest-sizes}

```{r}
#| label: fig-harvest-sizes
#| code-summary: "Plot distribution of individual harvest weights across farms"
#| fig-cap: "Individual harvest weights (plant-dominant diet)."

harvest_size <- file.path(gendata_path, "harvest_sizes", "farm_harvest_sizes.qs") %>% 
  qread()

harvest_size %>% 
  ggplot(aes(x = weight/1000)) +
  geom_histogram(binwidth = 0.2, colour = "black", fill = "salmon", alpha = 0.5) +
  scale_y_continuous(breaks = seq(0, 350, 50), limits = c(0, 375)) + 
  prettyplot_600() +
  labs(x = "Mean individual harvest weight (kg)", y = "Farm count")
```

## Daily values {#sec-daily-values}

```{r}
#| label: get-daily-values
#| code-summary: "Load and summarise daily C and N loading by country"

farm_runs <- map2_dfr(
  c(file.path(output_path, "model_run_outputs", "farm_runs_total_carbon.qs"),
    file.path(output_path, "model_run_outputs", "farm_runs_total_nitrogen.qs")),
  c("C", "N"),
  function(fnm, nutr) {
    fnm %>% 
      qread() %>% 
      filter(feed == "plant_dominant_biomar") %>% 
      select(-mean_value, sd_value) %>% 
      mutate(
        nutrient = as.factor(nutr),
        mean_value_perbiom = mean_value_perbiom %>% set_units("g g_fish-1") %>% set_units("kg t_fish-1") %>% drop_units(),
        sd_per_biom = sd_per_biom %>% set_units("g g_fish-1") %>% set_units("kg t_fish-1") %>% drop_units(),
      )
    }
  ) %>% 
  left_join(farm_locs, by = join_by(farm_ID))

# How many farms per country? 10 seems like enough
farm_runs <- farm_runs %>% 
  group_by(country, prod_t, nutrient) %>% 
  reframe(
    n = n(),
    mean = mean(mean_value_perbiom),
    sd = sqrt(mean(sd_per_biom^2))
  )
```

```{r}
#| label: fig-daily-values
#| code-summary: "Plot daily C loading over the production cycle by country"
#| fig-height: 7.5
#| fig-cap: "Daily carbon loading over the production cycle for the plant-dominant feed, summarised by country."

farm_runs %>% 
  filter(nutrient == "C") %>% 
  ggplot(aes(
    x = prod_t, 
    y = mean, 
    ymin = mean - sd, 
    ymax = mean + sd, 
    colour = country,
    fill = country
  )) +
  geom_line(linewdith = 0.75) +
  geom_ribbon(linewidth = 0.1, alpha = 0.25) +
  facet_wrap(~country, ncol = 2) +
  prettyplot_600() +
  labs(
    x = "Day of production",
    y = expression("Nutrient loading (kg"~t[fish]^{-1}*")")
)
```

## Model sensitivity {#sec-model-sensitivity}

```{r}
#| label: fig-sensitivity
#| code-summary: "Plot parameter sensitivity for individual weight and total excreted nutrients"
#| fig-cap: "Parameter sensitivity for individual weight and total excreted nutrient outputs."
#| fig-height: 6

sens <- file.path(output_sens_data_path, "sens_all.qs") %>% 
  qread()

param_names <- tibble::tribble(
  ~name,        ~lab,
  "alpha",      bquote(alpha),
  "epsprot",    bquote(epsilon["P"]),
  "epslip",     bquote(epsilon["L"]),
  "epscarb",    bquote(epsilon["C"]),
  "epsO2",      bquote(epsilon["O"[2]]),
  "pk",         bquote("pk"),
  "k0",         bquote("k"[0]),
  "m",          bquote("m"),
  "n",          bquote("n"),
  "betac",      bquote(beta*"C"),
  "Tma",        bquote("T"["max"]),
  "Toa",        bquote("T"["opt"]),
  "Taa",        bquote("T"["min"]),
  "omega",      bquote(omega),
  "a",          bquote("a"),
  "k",          bquote("k"),
  "eff",        bquote("eff"),
  "meanW",      bquote(bar("W")),
  "deltaW",     bquote(Delta*"W"), 
  "meanImax",   bquote(bar("I"["max"])),
  "deltaImax",  bquote(Delta*"I"["max"]),
  "overFmean",  bquote(bar("F")),
  "overFdelta", bquote(Delta*"F"),
  "mortmyt",    bquote("mort")
)

sens %>% 
  filter(measure %in% c("weight_stat", "total_excr_stat") & adj_param != "nruns") %>% 
  mutate(
    measure = factor(measure, 
      levels = c("weight_stat", "total_excr_stat"), 
      labels = c("Ind. weight", "Total excreted")
    )) %>% 
  ggplot(aes(x = adj_param, y = mean_sens, ymin = mean_sens-sd_sens, ymax = mean_sens+sd_sens)) +
  geom_col(colour = "black", fill = "salmon", alpha = 0.5) +
  geom_errorbar(width = 0.5) +
  coord_flip() +
  scale_x_discrete(limits = rev(param_names$name), labels = rev(param_names$lab)) +
  facet_wrap(~measure) +
  prettyplot_600() +
  labs(x = "Adjusted parameter", y = "Sensitivity")
```

[Next: TBD →](index.qmd)
