---
title: "Local-to-Global Mariculture Modelling"
subtitle: "Project Documentation - Index of files"
author: "Tormey Reimer"
date: today
execute:
  eval: false
knitr: 
  opts_chunk:
    fig.align: center
    fig.width: 8.5
format:
  html:
    code-fold: true
---

# Table of Contents

All code and processes for this project are linked below. Note that some scripts require access to raw data which is not stored on Github. 

## Main scripts

The main scripts for this project focus on setting up the model to examine the differences in total protein, carbohydrate, lipid and nitrogen losses between feeds. 

| Page | Description |
|---------|:-------------------------------------------------|
| **[Data Tidying](01_tidying_data.html)** | Initial data tidying and preparation |
| **[Extracting Production Locations](02_extracting_production_locations.html)** | Identification and processing of mariculture production site locations |
| **[Assessing Species Potential](03_assessing_species_potential.html)** | Evaluation of species suitability for different locations |
| **[Extracting Temperatures](04_extracting_temperatures.html)** | SST data extraction and processing for model inputs |
| **[Formulating Feeds](05_formulating_feeds.html)** | Feed composition formulation for model runs |
| **[How many runs per farm?](06_how_man_runs_per_farm.html)** | Determining how many Monte-Carlo runs is enough |
| **[Running Atlantic Salmon Model](06_run_atlantic_salmon_model.html)** | Atlantic salmon growth model execution |
| **[Data Analysis](08_data_analysis.html)** | Analysing the outputs of the Atlantic salmon growth model |
| **[Figures for manuscript](09_figures_for_manuscript.html)** | Code used to generate the specific figures used in the manuscript |

```{r setup}
#| cache: false
#| message: false

library(magrittr)
library(units)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ggh4x)
library(here)
library(ggpubr)
library(cowplot)
library(patchwork)
library(qs)
library(arrow)
library(targets)
library(sf)
library(terra)
library(rnaturalearth)
library(rnaturalearthdata)
library(rnaturalearthhires)
library(dtplyr)
library(furrr)
library(purrr)
library(future)
library(readxl)
library(conflicted)
library(lubridate)
library(RColorBrewer)
# library(lme4)
library(car)
library(emmeans)

conflicts_prefer(dplyr::select(), dplyr::filter(), dplyr::intersect())
# This markdown uses TinyTex - install with tinytex::install_tinytex()

here("src") %>% 
  list.files(pattern = "\\.R$", full.names = TRUE) %>% 
  walk(source)

packages <- c("magrittr", "dplyr", "tidyr", "stringr", "ggplot2", "arrow", "targets", "sf", "terra", "rnaturalearth", "furrr", "future", "units", "lubridate", "worrms", "aquamapsdata") %>% sort()
```

```{r}
#| label: global-variables
#| include: false
#| cache: false

remove_unit("g_fish")
remove_unit("kg_fish")
remove_unit("t_fish")
install_unit("g_fish")
install_unit("kg_fish", "1000 g_fish")
install_unit("t_fish", "1000 kg_fish")

species_params <- find_read(output_species_data_path, "species_params")
pop_params <- find_read(output_species_data_path, "pop_params")

farm_locs <- find_read(output_farm_data_path, "geometry") %>% 
  rename(farm_ID = farm_id) %>% 
  mutate(country = as.factor(country))
countries_long <- farm_locs %>%
  mutate(
    longitude = map_dbl(geometry, ~st_coordinates(.x)[1]),
    latitude = map_dbl(geometry, ~st_coordinates(.x)[2])
  ) %>% 
  st_drop_geometry() %>% 
  group_by(country) %>% 
  reframe(latitude = mean(latitude)) %>% 
  arrange(-latitude) %>% 
  pull(country) %>% 
  as.character() 
countries_short <- countries_long %>% 
  str_replace("United States of America", "USA") %>% 
  str_replace("Russian Federation", "Russia") %>% 
  str_replace("United States of America", "USA")
farm_locs <- farm_locs %>% 
  mutate(
    country = factor(
      country, 
      levels = countries_long,
      labels = countries_short
    ))
```

# Results {#sec-results}

```{r}
#| label: load-all-data-analysis

# Biomass produced
bpr1 <- find_read(output_model_farm_path, "biomass_produced") %>% 
  mutate(feed = factor(feed, levels = long_feeds, labels = short_feeds)) %>% 
  group_by(feed) %>% 
  reframe(mean = mean(mean) %>% set_units("g") %>% set_units("t") %>% drop_units() %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = mean)

bpr2 <- find_read(comparisons_path, "biomass_produced") %>% 
  group_by(feed) %>% 
  reframe(diff = (100*mean(mean)) %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = diff)

# Feed provided
fpr1 <- find_read(data_analysis_path, "food_prov_farm_summed") %>% 
  mutate(feed = factor(feed, levels = long_feeds, labels = short_feeds)) %>% 
  group_by(feed) %>% 
  reframe(total = mean(total) %>% set_units("g") %>% set_units("t") %>% drop_units() %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = total)

fpr2 <- find_read(comparisons_path, "food_prov_farm_summed") %>% 
  group_by(feed) %>% 
  reframe(diff = (100*mean(total)) %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = diff)

# Feed provided per biomass
fpr_pb1 <- find_read(data_analysis_path, "food_prov_farm_per_biom_summed") %>% 
  mutate(feed = factor(feed, levels = long_feeds, labels = short_feeds)) %>% 
  group_by(feed) %>% 
  reframe(mean_biom = mean(mean_biom) %>% set_units("g g_fish-1") %>% set_units("kg kg_fish-1") %>% drop_units() %>% round(2)) %>% 
  pivot_wider(names_from = feed, values_from = mean_biom)

fpr_pb2 <- find_read(comparisons_path, "food_prov_farm_per_biom_summed") %>% 
  group_by(feed) %>% 
  reframe(diff = (100*mean(mean_biom)) %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = diff)

# Uneaten feed
tuf1 <- find_read(data_analysis_path, "total_uneat_farm_summed") %>% 
  mutate(feed = factor(feed, levels = long_feeds, labels = short_feeds)) %>% 
  group_by(feed) %>% 
  reframe(total = mean(total) %>% set_units("g") %>% set_units("t") %>% drop_units() %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = total)

tuf2 <- find_read(comparisons_path, "total_uneat_farm_summed") %>% 
  group_by(feed) %>% 
  reframe(diff = (100*mean(total)) %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = diff)

# Excreted waste
tew1 <- find_read(data_analysis_path, "total_excr_farm_summed") %>% 
  mutate(feed = factor(feed, levels = long_feeds, labels = short_feeds)) %>% 
  group_by(feed) %>% 
  reframe(total = mean(total) %>% set_units("g") %>% set_units("t") %>% drop_units() %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = total)

tew2 <- find_read(comparisons_path, "total_excr_farm_summed") %>% 
  group_by(feed) %>% 
  reframe(diff = (100*mean(total)) %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = diff)

# Total waste
tow1 <- find_read(data_analysis_path, "total_mass_inputs_farm_summed") %>% 
  # mutate(feed = factor(feed, levels = long_feeds, labels = short_feeds)) %>% 
  group_by(feed) %>% 
  reframe(total = mean(total) %>% set_units("g") %>% set_units("t") %>% drop_units() %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = total)

tow2 <- find_read(comparisons_path, "total_mass_inputs_farm_summed") %>% 
  group_by(feed) %>% 
  reframe(diff = (100*mean(total)) %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = diff)

# Excretion proportion of waste
epw <- find_read(comparisons_path, "excr_proportion_of_waste") %>% 
  group_by(feed) %>% 
  reframe(diff = (100*mean(total)) %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = diff)

# Protein, lipid, carbs
plc1 <- find_read(data_analysis_path, "PLC_uneat_excr_farm_summed") %>% 
  # mutate(feed = factor(feed, levels = long_feeds, labels = short_feeds)) %>% 
  group_by(feed, measure, macro) %>% 
  reframe(total = mean(total) %>% set_units("g") %>% set_units("t") %>% drop_units() %>% round(1)) %>% 
  pivot_wider(names_from = c(feed, macro), values_from = total)

# Proportion of PLC waste from excretion
plc2 <- find_read(data_analysis_path, "PLC_uneat_excr_farm_summed") %>% 
  # mutate(feed = factor(feed, levels = long_feeds, labels = short_feeds)) %>% 
  group_by(feed, measure, macro) %>% 
  reframe(total = mean(total) %>% set_units("g") %>% set_units("t") %>% drop_units() %>% round(1)) %>% 
  pivot_wider(names_from = c(feed, macro, measure), values_from = total) %>% 
  mutate(
    MD_C = round(100*MD_C_excr/(MD_C_excr+MD_C_uneat),1),
    MD_L = round(100*MD_L_excr/(MD_L_excr+MD_L_uneat),1),
    MD_P = round(100*MD_P_excr/(MD_P_excr+MD_P_uneat),1),
    NI_C = round(100*NI_C_excr/(NI_C_excr+NI_C_uneat),1),
    NI_L = round(100*NI_L_excr/(NI_L_excr+NI_L_uneat),1),
    NI_P = round(100*NI_P_excr/(NI_P_excr+NI_P_uneat),1),
    PD_C = round(100*PD_C_excr/(PD_C_excr+PD_C_uneat),1),
    PD_L = round(100*PD_L_excr/(PD_L_excr+PD_L_uneat),1),
    PD_P = round(100*PD_P_excr/(PD_P_excr+PD_P_uneat),1)
  ) %>% 
  select(-contains(c("excr", "uneat")))

# % change in PLC waste
plc3 <- find_read(comparisons_path, "PLC_uneat_excr_farm_summed") %>% 
  group_by(feed, macro) %>% 
  reframe(diff = (100*mean(total)) %>% round(1)) %>% 
  pivot_wider(names_from = c(feed, macro), values_from = diff)

# N inputs, uneaten
niu1 <- find_read(data_analysis_path, "P_uneat_farm_summed") %>% 
  mutate(feed = factor(feed, levels = long_feeds, labels = short_feeds),
         total = total * 6.25,
         sd = sd * 6.25) %>% 
  group_by(feed) %>% 
  reframe(total = mean(total) %>% set_units("g") %>% set_units("t") %>% drop_units() %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = total)

niu2 <- find_read(comparisons_path, "N_uneat_farm_summed") %>% 
  group_by(feed) %>% 
  reframe(diff = (100*mean(total)) %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = diff)

# N inputs, excreted
nie1 <- find_read(data_analysis_path, "P_excr_farm_summed") %>% 
  mutate(feed = factor(feed, levels = long_feeds, labels = short_feeds),
         total = total * 6.25,
         sd = sd * 6.25) %>% 
  group_by(feed) %>% 
  reframe(total = mean(total) %>% set_units("g") %>% set_units("t") %>% drop_units() %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = total)

nie2 <- find_read(comparisons_path, "N_excr_farm_summed") %>% 
  group_by(feed) %>% 
  reframe(diff = (100*mean(total)) %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = diff)

# N inputs, total
nit1 <- find_read(data_analysis_path, "N_total_farm_summed.qs") %>% 
  # mutate(feed = factor(feed, levels = long_feeds, labels = short_feeds)) %>% 
  group_by(feed) %>% 
  reframe(total = mean(total) %>% set_units("g") %>% set_units("t") %>% drop_units() %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = total)

nit2 <- find_read(comparisons_path, "N_total_farm_summed.qs") %>% 
  group_by(feed) %>% 
  reframe(diff = (100*mean(total)) %>% round(1)) %>% 
  pivot_wider(names_from = feed, values_from = diff)
```

```{r not-in-use}
| Description | Plant-dominant | Marine-dominant | Novel-inclusive | Change from PD to MD (%) | Change from PD to NI (%) |
|:--------------------------|:----------:|:----------:|:----------:|:----------:|:----------:|
| Salmon biomass produced (t)                         | `r bpr1$PD`        | `r bpr1$MD`        | `r bpr1$NI`        | `r coln(bpr2$MD)`    | `r coln(bpr2$NI)`    |
| Feed provided (t)                                   | `r fpr1$PD`        | `r fpr1$MD`        | `r fpr1$NI`        | `r coln(fpr2$MD)`    | `r coln(fpr2$NI)`    |
| Feed provided (kg/kg salmon)                        | `r fpr_pb1$PD`     | `r fpr_pb1$MD`     | `r fpr_pb1$NI`     | `r coln(fpr_pb2$MD)` | `r coln(fpr_pb2$NI)` |
| Uneaten feed (t)                                    | `r tuf1$PD`        | `r tuf1$MD`        | `r tuf1$NI`        | `r coln(tuf2$MD)`    | `r coln(tuf2$NI)`    |
| Excreted waste (t)                                  | `r tew1$PD`        | `r tew1$MD`        | `r tew1$NI`        | `r coln(tew2$MD)`    | `r coln(tew2$NI)`    |
| Total waste (E+U, t)                                | `r tow1$PD`        | `r tow1$MD`        | `r tow1$NI`        | `r coln(tow2$MD)`    | `r coln(tow2$NI)`    |
| Proportion of total waste composed of excretion (%) | `r epw$PD`         | `r epw$MD`         | `r epw$NI`         |                      |                      |
| Total protein waste (t)                             | `r sum(plc1$PD_P)` | `r sum(plc1$MD_P)` | `r sum(plc1$NI_P)` | `r coln(plc3$MD_P)`  | `r coln(plc3$NI_P)`  |
| Total carbohydrate waste (t)                        | `r sum(plc1$PD_C)` | `r sum(plc1$MD_C)` | `r sum(plc1$NI_C)` | `r coln(plc3$MD_C)`  | `r coln(plc3$NI_C)`  |
| Total lipid waste (t)                               | `r sum(plc1$PD_L)` | `r sum(plc1$MD_L)` | `r sum(plc1$NI_L)` | `r coln(plc3$MD_L)`  | `r coln(plc3$NI_L)`  |
| Proportion of protein waste from excretion (%)      | `r plc2$PD_P`      | `r plc2$MD_P`      | `r plc2$NI_P`      |                      |                      |
| Proportion of carbohydrate waste from excretion (%) | `r plc2$PD_C`      | `r plc2$MD_C`      | `r plc2$NI_C`      |                      |                      |
| Proportion of lipid waste from excretion (%)        | `r plc2$PD_L`      | `r plc2$MD_L`      | `r plc2$NI_L`      |                      |                      |
| Total N from uneaten feed (t)                       | `r niu1$PD`        | `r niu1$MD`        | `r niu1$NI`        | `r coln(niu2$MD)`    | `r coln(niu2$NI)`    |
| Total N from excreted waste (t)                     | `r nie1$PD`        | `r nie1$MD`        | `r nie1$NI`        | `r coln(nie2$MD)`    | `r coln(nie2$NI)`    |
| Total N inputs (E+U, t)                             | `r nit1$PD`        | `r nit1$MD`        | `r nit1$NI`        | `r coln(nit2$MD)`    | `r coln(nit2$NI)`    |
: Summary of all results. Unless otherwise specified, values presented are means across all farms, summed across a single production period. In the % change columns, the colour of the text indicates whether there is an increase (red) or a decrease (green). Some cells of this table are blank - sometimes that means it doesn't make sense to fill that cell, but sometimes it means I haven't got to it yet. {#tbl-results-summary}
```

```{r}
#| label: PLC-waste-comparison

plc1 <- find_read(data_analysis_path, "PLC_uneat_excr_farm_summed")

lm_P <- lm(drop_units(total) ~ feed * measure, data = filter(plc1, macro == "P"))
aov_P <- aov(lm_P)
# summary(aov_P)

emm <- emmeans(lm_P, "feed", "measure")
tukey_results <- pairs(emm, adjust = "tukey")
# print(tukey_results)

lm_C <- lm(drop_units(total) ~ feed * measure, data = filter(plc1, macro == "C"))
aov_C <- aov(lm_C)
# summary(aov_C)

emm <- emmeans(lm_C, "feed", "measure")
tukey_results <- pairs(emm, adjust = "tukey")
# print(tukey_results)

lm_L <- lm(drop_units(total) ~ feed * measure, data = filter(plc1, macro == "L"))
aov_L <- aov(lm_L)
# summary(aov_L)

emm <- emmeans(lm_L, "feed", "measure")
tukey_results <- pairs(emm, adjust = "tukey")
# print(tukey_results)
```

```{r}
#| label: fig-PLC-waste-comparison
#| fig-cap: Mean waste lost to the environment in a complete production cycle when farms are fed the MD (blue), PD (green) or NI (red) diet. Errorbars show SD across all farms.

plc1 %>%
  group_by(measure, macro, feed) %>%
  reframe(
    sd = sd(total) %>% set_units("g") %>% set_units("t") %>% drop_units(),
    total = mean(total) %>% set_units("t") %>% drop_units(), 
  ) %>% 
  mutate(
    macro = factor(
      macro, 
      levels = c("P", "C", "L"),
      labels = c("Protein", "Carbohydrate", "Lipid")
    ),
    measure = factor(
      measure,
      levels = c("uneat", "excr"),
      labels = c("Uneaten", "Excreted")
    )
  ) %>% 
  ggplot(aes(
    x = measure, 
    y = total, 
    ymin = total-sd, ymax = total+sd,
    fill = feed
  )) +
  geom_col(position = position_dodge(), width = 0.95, colour = "black") +
  geom_errorbar(position = position_dodge(width = 0.95), width = 0.5) +
  facet_grid(cols = vars(macro)) +
  prettyplot() +
  scale_fill_manual(values = feed_pal) +
  labs(y = "Mean total waste (t)", x = "Source")
```

# N inputs only

```{r}
#| label: N-inputs-comparisons

N_uneat <- find_read(comparisons_path, "N_uneat_farm_per_biom_summed")
N_excr <- find_read(comparisons_path, "N_excr_farm_per_biom_summed")

N_all <- rbind(N_uneat, N_excr) %>% 
  st_drop_geometry() %>% 
  mutate(
    measure = droplevels(measure),
    country = factor(country,
                   levels = countries_long,
                   labels = countries_short)
  )

# ggplot(N_all, aes(x = country, y = mean_biom, fill = feed)) +
#   geom_boxplot() +
#   facet_grid(cols = vars(measure)) +
#   scale_fill_manual(values = feed_pal)
```

# Spatially

```{r}
#| label: fig-spatial-N-inputs
#| fig-height: 12
#| fig-cap: Total N inputs per unit biomass of salmon produced when farms are fed the MD feed (top), the PD feed (middle) or the NI feed (bottom).

N_ins <- rbind(
  find_read(data_analysis_path, "P_excr_farm_per_biom_summed"),
  find_read(data_analysis_path, "P_uneat_farm_per_biom_summed")
) %>% 
  st_drop_geometry() %>% 
  mutate(feed = factor(feed, levels = long_feeds, labels = short_feeds)) %>% 
  group_by(country, farm_ID, feed) %>% 
  reframe(
    mean_biom = sum(mean_biom) %>% set_units("kg t_fish-1") %>% drop_units(), 
    sd_biom = sqrt(sum(sd_biom^2)) %>% set_units("g g_fish-1") %>% set_units("kg t_fish-1") %>% drop_units()
  ) %>% 
  merge(farm_locs, by = c("country", "farm_ID")) %>% 
  st_as_sf()
  
p_bigmap_robinson +
  geom_sf(
    data = N_ins, 
    aes(colour = mean_biom, geometry = geometry), 
    size = 2
  ) +
  coord_sf() +
  scale_color_viridis_c(option = "turbo", breaks = seq(0, 120, 10)) + 
  guides(col = guide_colourbar(
    title = expression("Total N inputs (kg/t salmon)"),
    direction = "horizontal", position = "bottom",
    label.position = "bottom", title.position = "top",
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 1,
    barheight = 1.5, barwidth = 20,
  )) +
  facet_grid(rows = vars(feed)) +
  # prettyplot() +
  theme(strip.background = element_blank(), strip.text = element_blank(),
        axis.title = element_blank(),
        legend.position = "bottom")
```

```{r}
#| label: N-inputs-patchwork-map

p_mercator <- p_bigmap_mercator +
  geom_sf(
    data = N_ins,
    aes(colour = mean_biom, geometry = geometry), 
    size = 2.5
  ) +
  coord_sf() +
  scale_color_viridis_c(option = "turbo", breaks = seq(0, 120, 10)) + #, limits = c(-30, 10)
  guides(col = guide_colourbar(
    title = expression("Total N inputs (kg/t salmon)"),
    direction = "horizontal", position = "bottom",
    label.position = "bottom", title.position = "top",
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 1,
    barheight = 1.5, barwidth = 20,
  )) +
  # facet_wrap(facets = vars(feed), nrow = 2, ncol = 2) +
  prettyplot() +
  theme(
    # strip.background = element_blank(), strip.text = element_blank(),
    legend.position = "bottom"
  ) +
  labs(y = "Latitude", x = "Longitude")

ins <- get_insets_mercator(p_mercator)

ins[[1]] + facet_grid(cols = vars(feed))
ins[[2]] + facet_grid(rows = vars(feed))
ins[[3]] + facet_grid(cols = vars(feed))
ins[[4]] + facet_grid(cols = vars(feed))
ins[[5]] + facet_grid(cols = vars(feed))
```

```{r}
#| label: fig-spatial-N-inputs-change
#| fig-height: 12
#| fig-cap: Change in total N inputs per unit biomass of salmon produced when farms are fed the MD feed (top), or the NI feed (bottom) compared to when farms are fed the PD feed.

N_ins_comp <- rbind(
  find_read(data_analysis_path, "P_excr_farm_per_biom_summed"),
  find_read(data_analysis_path, "P_uneat_farm_per_biom_summed")
  ) %>% 
  st_drop_geometry() %>% 
  select(-sd_biom) %>% 
  mutate(feed = factor(feed, levels = long_feeds, labels = short_feeds)) %>% 
  pivot_wider(
    names_from = feed,
    values_from = mean_biom,
    values_fn = drop_units
  ) %>% 
  group_by(country, farm_ID) %>% 
  reframe(
    PD = sum(PD),
    MD = sum(MD), 
    NI = sum(NI)
  ) %>% 
  mutate(MD = (MD-PD),
         NI = (NI-PD)) %>% 
  select(-PD) %>% 
  pivot_longer(
    names_to = "feed",
    values_to = "diff",
    cols = c(MD, NI)
  ) %>% 
  mutate(
    feed = factor(feed, levels = short_feeds),
    feed = droplevels(feed),
    diff = diff %>% set_units("g g_fish-1") %>% set_units("kg t_fish-1") %>% drop_units()
  ) %>% 
  merge(farm_locs, by = c("country", "farm_ID")) %>% 
  st_as_sf()
  
p_bigmap_robinson +
  geom_sf(
    data = N_ins_comp, 
    aes(colour = diff, geometry = geometry), 
    size = 2
  ) +
  coord_sf() +
  scale_color_viridis_c(option = "turbo", breaks = seq(10, 35, 5), limits = c(10,36)) + 
  guides(col = guide_colourbar(
    title = expression("Difference in N inputs (kg/t salmon)"),
    direction = "horizontal", position = "bottom",
    label.position = "bottom", title.position = "top",
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 1,
    barheight = 1.5, barwidth = 20,
  )) +
  facet_grid(rows = vars(feed)) +
  # prettyplot() +
  theme(strip.background = element_blank(), strip.text = element_blank(),
        axis.title = element_blank(),
        legend.position = "bottom")
```

```{r}
#| label: N-inputs-change-patchwork-map
#| fig-cap: Change in total N inputs per unit biomass of salmon produced when farms are fed the MD feed (A), or the NI feed (B) compared to when farms are fed the PD feed.

p_mercator <- p_bigmap_mercator +
  geom_sf(
    data = N_ins_comp,
    aes(colour = diff, geometry = geometry), 
    size = 2.5
  ) +
  coord_sf() +
  scale_color_viridis_c(option = "turbo", breaks = seq(10, 35, 5), limits = c(10,36)) + 
  guides(col = guide_colourbar(
    title = expression("Difference in N inputs (kg/t salmon)"),
    direction = "horizontal", position = "bottom",
    label.position = "bottom", title.position = "top",
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 1,
    barheight = 1.5, barwidth = 20,
  )) +
  prettyplot() +
  theme(
    strip.background = element_blank(), strip.text = element_blank(),
    legend.position = "bottom"
  ) +
  labs(y = "Latitude", x = "Longitude")

facet_labels <- c(
  "MD" = "A",
  "NI" = "B"
)

ins <- get_insets_mercator(p_mercator)

ins[[1]] + facet_grid(cols = vars(feed), labeller = as_labeller(facet_labels))
ins[[2]] + facet_grid(rows = vars(feed), labeller = as_labeller(facet_labels))
ins[[3]] + facet_grid(cols = vars(feed), labeller = as_labeller(facet_labels))
ins[[4]] + facet_grid(cols = vars(feed), labeller = as_labeller(facet_labels))
ins[[5]] + facet_grid(cols = vars(feed), labeller = as_labeller(facet_labels))
```

# Supplemental methods {#sec-supp-methods}

```{r}
#| label: location-data

farm_locs <- find_read(output_farm_data_path, "geometry") %>% 
  rename(farm_ID = farm_id)

farm_mean_lat <- farm_locs %>%
  st_centroid() %>%
  mutate(latitude = st_coordinates(.)[, "Y"]) %>%
  st_drop_geometry() %>%
  group_by(country) %>%
  summarise(mean_latitude = mean(latitude, na.rm = TRUE)) %>% 
  arrange(-mean_latitude)

countries <- farm_mean_lat$country
countries[str_detect(countries, "United States of America")] <- "USA"
countries[str_detect(countries, "United Kingdom")] <- "UK"
countries[str_detect(countries, "Russian Federation")] <- "Russia"

farm_locs <- farm_locs %>% 
  mutate(country = factor(country, levels = farm_mean_lat$country, labels = countries))
```

## Population and biomass {#sec-population-biomass}

```{r}
#| label: raw biomass and weight data

weight_cohort <- find_read(output_model_cohort_path, "weight_stat_all_farms")
biomass_cohort <- find_read(output_model_cohort_path, "biomass_stat_all_farms")
```

```{r}
#| label: fig-harvest-size
#| fig-cap: |
#|   Frequency distribution of harvest sizes used to create the population timeseries for each farm, as predicted by the fish-growth model. 

harv_size <- find_read(output_farm_data_path, "farm_harvest_size")

harv_size %>% 
  mutate(weight = weight %>% set_units("g") %>% set_units("kg") %>% drop_units()) %>% 
  select(c(farm_ID, weight)) %>% 
  ggplot(aes(x = weight)) +
  geom_histogram(binwidth = 0.25, colour = "black", fill = "salmon", alpha = 0.75) +
  scale_y_continuous(limits = c(0,500)) +
  scale_x_continuous(breaks = seq(0,7,0.5)) +
  labs(y = "Frequency", x = "Harvest weight (kg)") +
  prettyplot() +
  theme(aspect.ratio = 0.75)
```


```{r}
#| label: fig-cohorts
#| fig-cap: | 
#|   Example of how the overlapping cohorts combine into a single farm production curve (in this case biomass). The left plot shows the farm biomass produced by each cohort from initialisation to harvest, while the right plot shows the combined farm biomass over the course of a single production period of 548 days.

farm_biomass <- find_read(output_model_farm_path, "biomass_stat") %>% 
  filter(farm_ID == 722 & feed == "plant_dominant")

cohort_biomass_split <- rbind(
  farm_biomass %>% 
    farm_to_cohort() %>% 
    mutate(cohort = 1),
  farm_biomass %>% 
    farm_to_cohort(time_offset = 365) %>% 
    mutate(cohort = 2),
  farm_biomass %>% 
    farm_to_cohort(time_offset = 730) %>% 
    mutate(cohort = 3)
) %>% 
  mutate(split = 0) %>% 
  relocate(farm_ID, .before = t) %>% 
  relocate(feed, .before = t)

cohort_biomass <- find_read(output_model_cohort_path, "biomass_stat") %>% 
  filter(farm_ID == 722 & feed == "plant_dominant") %>% 
  mutate(cohort = "all", split = 1)

rbind(
  cohort_biomass_split,
  cohort_biomass
  ) %>% 
  mutate(
    t = t-min(t)+1,
    mean = mean %>% set_units("g") %>% set_units("t"),
    sd = sd %>% set_units("g") %>% set_units("t")
  ) %>% 
  ggplot(aes(x = t, y = mean, ymin = mean-sd, ymax = mean+sd, colour = as.factor(cohort))) +
  geom_line(linewidth = 0.75) +
  geom_ribbon(linetype = "dotted", alpha = 0) +
  facet_grid(cols = vars(split)) +
  scale_x_continuous(breaks = seq(0, 1600, 120)) +
  scale_colour_manual(values = cohort_pal) +
  prettyplot() +
  theme(strip.background = element_blank(), strip.text = element_blank()) +
  labs(y = "Farm biomass", x = "Day of production")
```

```{r}
#| label: pop-biom-data

biomass_farm <- find_read(output_model_farm_path, "biomass_stat") %>% 
  group_by(farm_ID, feed) %>% 
  slice_max(t) %>% 
  ungroup()

production <- find_read(output_farm_data_path, "farm_static_data") %>% 
  merge(biomass_farm, by = "farm_ID", all.x = F) %>% 
  mutate(
    mean = set_units(mean, "g"),
    diff = set_units(mean - tonnes_per_farm, "t"),
    perc_diff = diff/tonnes_per_farm
  ) %>% 
  group_by(country, feed) %>% 
  reframe(
    min_percdiff = min(perc_diff),
    max_percdiff = max(perc_diff),
    mean_percdiff = mean(perc_diff)
  )
```


## Feed provision and feeding response to temperature {#sec-model-feeding}


```{r}
#| label: fig-functional-response-to-temperature
#| fig-cap: Temperature changes in potential ingestion (solid line) and amount of food provided (dotted line), relative to maximum potential ingestion. The vertical dotted lines show the actual minimum and maximum daily temperatures experienced by farms represented in the model.

Sys.setenv(TAR_PROJECT = "project_farmruns")
farm_ts_data <- tar_read(farm_ts_data_chunked) %>% bind_rows()
min_T <- min(farm_ts_data$temp_c)
max_T <- max(farm_ts_data$temp_c)

all_params <- c(
  find_read(output_species_data_path, "species_params"),
  find_read(output_species_data_path, "pop_params")
)

func_T_resp <- data.frame(water_temp = seq(-5, 26, 0.25))
weight <- 1000

# Relative feeding rate with temperature
func_T_resp$feeding_rate <- sapply(
  FUN = feeding_rate, 
  X = func_T_resp$water_temp, 
  species_params = all_params
)

# Maximum potential ingestion (g) at temperature
func_T_resp$ing_pot <- all_params['meanImax'] * (weight^all_params['m']) * func_T_resp$feeding_rate

# Food provided is based on calculated potential ingestion BUT if temperature is too low they are still provided with the food that would be required at lowest temperature (when feeding rate is ~18.3%)
func_T_resp$food_prov <- food_prov_rate(
  water_temp = func_T_resp$water_temp, 
  ing_pot = func_T_resp$ing_pot,
  ing_pot_min = all_params['meanImax'] * (weight^all_params['m']) * feeding_rate(all_params['Taa'], all_params),
  species_params = all_params,
  pop_params = all_params
)

func_T_resp$food_enc <- all_params['eff'] * func_T_resp$food_prov
func_T_resp <- func_T_resp %>% 
  mutate(ing_act = pmin(food_enc, ing_pot))

max_ing <- max(func_T_resp$ing_act)

func_T_resp %>% 
  ggplot(aes(x = water_temp, y = ing_act/max_ing)) +
  geom_line(linewidth = 0.75) +
  geom_line(aes(y = food_prov/max_ing), linewidth = 0.75, linetype = "dotted") +
  geom_vline(xintercept = min_T, linetype = "dotted") +
  geom_vline(xintercept = max_T, linetype = "dotted") +
  # scale_y_continuous(breaks = seq(0,1.5,0.25)) +
  scale_x_continuous(breaks = seq(-5,30,5), limits = c(-5,26)) +
  labs(x = expression("Water temperature ("*degree*"C)"), y = "Feedd ingested/provided") +
  prettyplot()
```

```{r}
#| label: patchwork-map-mercator
#| eval: false
#| fig-cap: |
#|   Example of a patchwork
#| fig-height: 9
#| fig-width: 9.5

p_bigmap_feeding <- p_bigmap_mercator +
  geom_sf(data = feeding, aes(color = prop_feed_eaten, geometry = geometry), size = 2) +
  scale_color_viridis_c(option = "turbo") +
  guides(col = guide_colourbar(
    title = "Proportion of provided feed eaten",
    direction = "vertical", position = "right",
    label.position = "right", title.position = "top", 
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 2,
    barwidth = 1.5, barheight = 20
  )) +
  prettyplot() +
  labs(y = "Latitude", x = "Longitude")

patchwork_mercator(p_bigmap_feeding)
```

## Modelled experimental feeds {#sec-feeds}

```{r}
#| label: feed-details

feed_params <- find_read(output_species_data_path, "feed_params")

feed_summary <- map_dfr(names(feed_params), function(feed) {
  df <- feed_params[[feed]]
  map_dfr(c("Proteins", "Lipids", "Carbohydrates"), function(macro) {
    contrib <- df[[macro]]$proportion * df[[macro]]$macro
    total <- sum(contrib)
    digest <- sum(df[[macro]]$digest * contrib) / total
        tibble(
      feed = feed,
      macro = macro,
      total = total,
      digest = digest
    )
  })
}) %>%
  mutate(macro = factor(macro, levels = c("Proteins", "Lipids", "Carbohydrates"), labels = c("P", "L", "C")),
         feed = as.factor(feed),
         total_r = total * 100 %>% round(1),
         digest_r = digest * 100 %>% round(1))
```

## Model sensitivity to parameter changes {#sec-param-sensitivity}

```{r}
#| label: salmon-param-names

params <- find_read(output_species_data_path, "sens_params")

param_names <- tibble::tribble(
  ~name,        ~lab,
  "alpha",      bquote(alpha),
  "epsprot",    bquote(epsilon["P"]),
  "epslip",     bquote(epsilon["L"]),
  "epscarb",    bquote(epsilon["C"]),
  "epsO2",      bquote(epsilon["O"[2]]),
  "pk",         bquote("pk"),
  "k0",         bquote("k"[0]),
  "m",          bquote("m"),
  "n",          bquote("n"),
  "betac",      bquote(beta*"C"),
  "Tma",        bquote("T"["max"]),
  "Toa",        bquote("T"["opt"]),
  "Taa",        bquote("T"["min"]),
  "omega",      bquote(omega),
  "a",          bquote("a"),
  "k",          bquote("k"),
  "eff",        bquote("eff"),
  "meanW",      bquote(bar("W")),
  "deltaW",     bquote(Delta*"W"), 
  "meanImax",   bquote(bar("I"["max"])),
  "deltaImax",  bquote(Delta*"I"["max"]),
  "overFmean",  bquote(bar("F")),
  "overFdelta", bquote(Delta*"F"),
  "mortmyt",    bquote("mort")
)
```

```{r}
#| label: fig-sensitivity-results
#| fig-cap: Sensitivity of modelled individual fish weight (blue) and total farm excretion across a full production period (red) to a 10% increase in each model parameter. Bars show mean values, error bars show standard deviation between farms. 
#| fig-width: 8

sens_results <- find_read(output_sens_data_path, "sens_results") %>% 
  filter(measure %in% c("total_excr", "weight") & adj_param != "nruns") %>% 
  mutate(measure = droplevels(measure))

sens_results %>% 
  ggplot(aes(x = adj_param, y = mean_sens, ymin = mean_sens-sd_sens, ymax = mean_sens+sd_sens, fill = measure)) +
  geom_col(position = position_dodge(), width = 0.95, alpha = 0.65, colour = "black") +
  geom_errorbar(position = position_dodge(width = 0.95), width = 0.5) +
  coord_flip() +
  prettyplot() +
  scale_x_discrete(breaks = param_names$name, labels = param_names$lab) +
  scale_y_continuous(breaks = seq(-10, 30, 5)) +
  scale_fill_brewer(palette = "Set1") +
  labs(y = "Sensitivity", x = "Adjusted parameter")
```

# Code and packages used {#sec-code}

```{r not-in-use-cite}
# All analysis was conducted in R version `r R.version[['major']]`.`r R.version[['minor']]` "`r R.version[['nickname']]`" [@R_base]. 
# Major packages used include `r cite_packages(packages)`.

# For a full list of R packages used see the lockfile on [Github](https://www.github.com).
# This manuscript was written in Quarto [@quarto] using TinyTex [@tinytex] and the acronyms extension [@acronyms]. 
```
