---
title: "Local to Global Mariculture Modelling"
subtitle: "Project Documentation"
author: "Dr Tormey Reimer"
date: today
execute:
  eval: true
  echo: false
  warning: false
  message: false
  cache: true
  freeze: auto
knitr: 
  opts_chunk:
    fig.align: center
    fig.width: 8.5
---

# Table of Contents

This project contains a comprehensive analysis of mariculture modelling from local to global scales. Below you'll find links to all the main documentation pages:

| Page | Description |
|---------|:-------------------------------------------------|
| **[Data Tidying](01_tidying_data.html)** | Initial data cleaning and preparation procedures |
| **[Extracting Production Locations](02_extracting_production_locations.html)** | Identification and processing of mariculture production site locations |
| **[Assessing Species Potential](03_assessing_species_potential.html)** | Evaluation of species suitability for different locations |
| **[Extracting Temperatures](04_extracting_temperatures.html)** | Temperature data extraction and processing for model inputs |
| **[Formulating Feeds](05_formulating_feeds.html)** | Feed composition analysis and formulation procedures |
| **[Running Atlantic Salmon Model](06_run_atlantic_salmon_model.html)** | Atlantic salmon growth model implementation and execution |
| **[Species Layers](07_species_layers.html)** | Spatial analysis of species distribution layers |
| **[Relative N Inputs](08_relative_N_inputs.html)** | Analysis of nitrogen input patterns and their relative impacts |

# Data analysis

```{r setup}
#| include: false
#| eval: true
#| cache: false
#| message: false

library(magrittr)
library(units)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ggh4x)
library(here)
library(ggpubr)
library(cowplot)
library(patchwork)
library(qs)
library(arrow)
library(targets)
library(sf)
library(terra)
library(rnaturalearth)
library(rnaturalearthdata)
library(rnaturalearthhires)
library(dtplyr)
library(furrr)
library(purrr)
library(future)
library(readxl)
library(conflicted)
library(lubridate)
library(RColorBrewer)
conflicts_prefer(dplyr::select(), dplyr::filter(), dplyr::intersect(), .quiet = T)
# This markdown uses TinyTex - install with tinytex::install_tinytex()

here("src") %>% list.files(pattern = "\\.R$", full.names = TRUE) %>% walk(source)

packages <- c("magrittr", "dplyr", "tidyr", "stringr", "ggplot2", "arrow", "targets", "sf", "terra", "rnaturalearth", "furrr", "future", "units", "lubridate", "worrms", "aquamapsdata") %>% sort()
```

```{r global-variables}
#| include: false

remove_unit("g_fish")
remove_unit("kg_fish")
remove_unit("t_fish")
install_unit("g_fish")
install_unit("kg_fish", "1000 g_fish")
install_unit("t_fish", "1000 kg_fish")

# Filenames
# species_params_excel <- c(file = file.path(input_species_param_path, "Species.xlsx"), sheet = "Atlantic salmon")
# pop_params_excel <- c(file = file.path(input_species_param_path, "Population.xlsx"))

species_params <- find_read(output_species_data_path, "species_params")
pop_params <- find_read(output_species_data_path, "pop_params")
  
feed_pal <- c("past" = "#E41A1C", "reference" = "#377EB8", "future" = "#4DAF4A")  
cohort_pal <- c("1" = "#E41A1C", "2" = "#377EB8", "3" = "#4DAF4A", "all" = "black")
macro_pal <- c(
  "protein" = "darkred", "P" = "darkred", 
  "carbohydrates" = "steelblue", "C" = "steelblue", 
  "lipids" = "darkgoldenrod1", "L" = "darkgoldenrod1"

# Model functioning

## Population and biomass

```{r raw biomass and weight data}
weight_cohort <- find_read(output_model_cohort_path, "weight_stat_all_farms")
biomass_cohort <- find_read(output_model_cohort_path, "biomass_stat_all_farms")
```
)
```



