---
title: "Run aquaculture model for Atlantic salmon"
format: html
editor: source
---

# Introduction

This document runs the aquaculture model for Atlantic salmon farms. It processes farm location data, species parameters, and conducts sensitivity analyses to understand the impact of various parameters on fish and farm growth measures.

```{r setup, include=F}
#| warning: false
#| message: false

library(tidyverse)
library(arrow)
library(sf)
library(terra)
library(furrr)
library(future)
library(tictoc)
library(fs)
library(conflicted)
library(readxl)
library(units)
library(qs)
library(here)
library(targets)
conflicted::conflicts_prefer(dplyr::filter(), dplyr::select(), .quiet = T)

here("src") %>% 
  list.files(pattern = "\\.R$", full.names = TRUE) %>% 
  str_subset("map", negate = T) %>% 
  walk(source)

remove_unit("g_fish")
remove_unit("kg_fish")
remove_unit("t_fish")
install_unit("g_fish")
install_unit("kg_fish", "1000 g_fish")
install_unit("t_fish", "1000 kg_fish")
```

```{r filenames}
#| code-summary: Set up global filenames

farm_script <- here("targets/_targets_farmruns.R")
farm_store <- here("targets/_targets_farmruns")
sens_script <- here("targets/_targets_sensitivities.R")
sens_store <- here("targets/_targets_sensitivities")

# Filenames
# farm_locations_parquet <- file.path(input_farm_coords_path, "farm_coords.parquet")
# farm_coords_file <- file.path(output_farm_data_path, "farm_coords.qs")
# farm_geometry_file <- file.path(output_farm_data_path, "farm_geometry.qs")
# sens_params_file <- file.path(output_species_data_path, "sens_params.qs")
# feed_params_file <- file.path(output_species_data_path, "feed_params.qs")
```

# Data for targets pipelines

Much of the actual analysis is run through targets pipelines. Therefore, we need to make sure that the files going into those pipelines are correct and up to date.

```{r farm-coordinates, eval=F}
#| code-summary: Load and process farm coordinate data with appropriate timing parameters for Northern and Southern hemisphere farms

times_N <- c("t_start" = 121, "t_end" = 121+547, "dt" = 1)
times_S <- c("t_start" = 274, "t_end" = 274+547, "dt" = 1)

farm_static_data <- file.path(input_farm_coords_path, "farm_static_data.qs") %>% 
  qread() %>% 
  mutate(
    t_start = case_when(hemisphere == "N" ~ times_N['t_start'], TRUE ~ times_S['t_start']) %>% unname() %>% as.integer(), 
    t_end = case_when(hemisphere == "N" ~ times_N['t_end'], TRUE ~ times_S['t_end']) %>% unname() %>% as.integer(),
    ) %>% 
  rename(farm_ID = farm_id)

qsave(farm_static_data, file.path(output_farm_data_path, "farm_static_data.qs"))
```

## Farm Time Series Data

Process Sea Surface Temperature (SST) data for each farm location.

```{r farm-sst-data, eval=F}
#| code-summary: Process Sea Surface Temperature (SST) data for each farm location.

farm_ts_data <- file.path(input_farm_coords_path, "farm_temporal_data.qs") %>% 
  qread() %>% 
  mutate(
    t_start = case_when(hemisphere == "N" ~ times_N['t_start'], TRUE ~ times_S['t_start']) %>% unname() %>% as.integer(), 
    t_end = case_when(hemisphere == "N" ~ times_N['t_end'], TRUE ~ times_S['t_end']) %>% unname() %>% as.integer(),
  ) %>% 
  filter(day >= t_start & day <= t_end) %>% 
  rename(farm_ID = farm_id) %>% 
  select(c(farm_ID, day, temp_c)) %>% 
  arrange(farm_ID, day)

qsave(farm_ts_data, file.path(output_farm_data_path, "farm_ts_data.qs"))
```

## Species and population parameters

```{r species-parameters}
#| code-summary: Load species-specific parameters from Excel file
#| cache: true

species_params <- readxl::read_excel(
  path = c(file = file.path(input_species_param_path, "Species.xlsx"), sheet = "Atlantic salmon")["file"], 
  sheet = c(file = file.path(input_species_param_path, "Species.xlsx"), sheet = "Atlantic salmon")["sheet"]
)
vals <- species_params$Value
names(vals) <- species_params$Quantity
species_params <- vals[!is.na(vals)]
qsave(species_params, file.path(output_species_data_path, "species_params.qs"))
```

```{r population-parameters}
#| code-summary: Load population-specific parameters
#| cache: true

pop_params <- readxl::read_excel(path = c(file = file.path(input_species_param_path, "Population.xlsx"))["file"])
vals <- pop_params$Value
names(vals) <- pop_params$Quantity
pop_params <- vals[!is.na(vals)]
qsave(pop_params, file.path(output_species_data_path, "pop_params.qs"))
```

## Feed digestibility

```{r variable-digestibility}
#| code-summary: Show the variability in feed digestibility generated by sampling digestibility values.
#| fig-cap: Whole-feed digestibility of the feeds run in this analysis. 

# tar_make(names = c("feed_params", "feed_names"), script = farm_script, store = farm_store)

tar_load(feed_params, store = farm_store)
tar_load(feed_names, store = farm_store)

feed_params <- feed_params %>% 
  map2_dfr(., feed_names, function(fp, fn) {
    ld <- filter(fp[["Lipids"]], macro != 0)
    ld <- sum(ld$proportion * ld$digest)/sum(ld$proportion)

    cd <- filter(fp[["Carbohydrates"]], macro != 0)
    cd <- sum(cd$proportion * cd$digest)/sum(cd$proportion)

    pd <- filter(fp[["Proteins"]], macro != 0)
    pd <- sum(pd$proportion * pd$digest)/sum(pd$proportion)

    data.frame(
      protein_digestibility = pd,
      carb_digestibility = cd,
      lipid_digestibility = ld,
      feed_name = fn
    )
  }) %>% 
  pivot_longer(
    cols = contains("digestibility"),
    names_to = "macro", 
    values_to = "digestibility"
  ) %>% 
  mutate(
    macro = factor(
      macro, 
      levels = c("protein_digestibility", "lipid_digestibility", "carb_digestibility"),
      labels = c("protein", "lipid", "carb")
    ),
    feed_name = as.factor(feed_name)
  )

feed_params %>% 
  ggplot(aes(x = macro, y = digestibility, fill = feed_name)) +
  geom_col(position = position_dodge()) +
  theme_classic()
```

# Farm Harvest Size Calculations

First, let's test the pipeline to make sure that the feed input data is being processed correctly.

```{r check-farmrun-pipeline}
#| code-summary: Check that the pipeline will run correctly

Sys.setenv(TAR_PROJECT = "project_farmruns")

tar_validate()
# tar_prune()

tar_make(test_reference_feed)

tar_load_globals()
reference_feed_name
tar_read(test_reference_feed) # if weight < 125g then feeds aren't processing properly
```

Calculations of expected harvest sizes for each farm (using default feed and no Monte-Carlo variation) are done in the targets pipeline. This sets a "target population" for each farm and grounds the growth somewhat.

```{r farm-harvest-size, eval=F}
#| code-summary: Run limited pipeline for harvest sizes and testing

Sys.setenv(TAR_PROJECT = "project_farmruns")

# Get harvest sizes for all farms
tar_make(names = contains(c("harvest_size", "feed_params")))
```

```{r save-harvest-size}
#| code-summary: Save harvest size data

harvest_size <- tar_read(harvest_size) %>% 
  bind_rows()

ggplot(harvest_size, aes(x = weight)) +
  geom_histogram(fill = "salmon", alpha = 0.75, colour = "black")

qsave(harvest_size, file.path(output_farm_data_path, "farm_harvest_size.qs"))
```

# Sensitivity analysis

Run the sensitivity runs using the targets pipeline - note that this can take a long time. The species parameters (n=19) are run for a single fish in  a sample number of farms (n=272) and each factor (3). The population parameters (n=6) are run for 1000 fish in a sample number of farms (271) and each factor (3). 

```{r check-sensitivity-pipeline}
#| code-summary: Check that the sensitivities pipeline is running correctly

Sys.setenv(TAR_PROJECT = "project_sensitivities")
tar_validate()
# tar_visnetwork(targets_only = T)

tar_make(test_reference_feed)
tar_read(test_reference_feed) # if weight ~ 0 then feeds aren't processing properly, fix before proceeding
```

```{r sensitivity-run}
#| code-summary: Run sensitivity pipeline

Sys.setenv(TAR_PROJECT = "project_sensitivities")

tar_make(seconds_meta_append = 300)
tar_prune()
```

## Process sensitivity results

Combine and visualise sensitivity analysis results.

```{r sensitivity-results}
#| code-summary: Plot and save sensitivity results

tar_load(sens_results_spec, store = sens_store)
tar_load(sens_results_pop, store = sens_store)

sens_results <- rbind(
  sens_results_pop,
  sens_results_spec
)

qsave(sens_results, file.path(output_sens_data_path, paste0("sens_results_all.qs")))

sens_measures <- levels(sens_results$measure)
sens_results_figfiles <- file.path(output_sens_data_path, paste0("sens_plot_", sens_measures, ".png"))

for (sm in seq_along(sens_measures)) {
  p <- sens_results %>% 
    filter(measure == sens_measures[sm]) %>% 
    ggplot(aes(x = adj_param, y = mean_sens, ymin = mean_sens-sd_sens, ymax = mean_sens+sd_sens)) +
    geom_col(fill = "salmon", alpha = 0.65, colour = "black") +
    geom_errorbar(width = 0.5) +
    coord_flip() +
    theme_classic()
  ggsave(sens_results_figfiles[sm])
}
```

# Run farms

The following code runs the main pipeline in two stages. The pre-run prepares all the pre-requisites for the main results targets, to reduce load. The main run shortcuts past all the pre-run targets and does the main model runs and results processing. 

```{r run-farms-pipeline, eval=F}
#| code-summary: Run the main pipeline

Sys.setenv(TAR_PROJECT = "project_farmruns")

tar_validate()

# Run whole pipeline
tar_make(seconds_meta_append = 300)
```

Load objects from the targets pipeline and save them as .qs files, which are much quicker to load down the line. This takes a few minutes to gather all the targets.

```{r all-farms-cohorts-results}
#| code-summary: Save pipeline objects as files
#| eval: false

Sys.setenv(TAR_PROJECT = "project_farmruns")

tar_load(farm_results)
qsave(farm_results, file.path(output_path, "total_ends.qs"))

# tar_load(cohort_results)

# for (m in levels(cohort_results$measure)) {
#   df <- cohort_results %>% 
#     filter(measure == m)
#   qsave(
#     df, 
#     file.path(
#       output_path, 
#       str_c("cohort_results_", m,".qs")
#     )
#   )
# }

# rm(farm_results, cohort_results, df)
# gc()
```

# Further processing - C and N

```{r total-values}
#| code-summary: Get total values data. These data give the total (summed) values of each response metric over 1 complete production period.

# Change level names, setting units
total_vals <- file.path(output_path, "total_ends.qs") %>% 
  qread() %>% 
  mutate(
    feed_group = factor(
      str_remove(feed, "_min|_max"), 
      levels = orig_feeds,
      labels = short_feeds
      ),
    feed_group = factor(feed_group, levels = c("MD","PD","AD")), # reorder the levels
    feed_type = case_when(
      str_detect(feed, "_min") ~ "min",
      str_detect(feed, "_max") ~ "max",
      T ~ "mean"
      ),
    feed_type = factor(
      feed_type, 
      levels = c("min", "mean", "max"), 
      labels = c("Minimum", "Mean", "Maximum")
    ),
    mean = set_units(mean, "g"),
    sd = set_units(sd, "g"),
    mean_biomass = set_units(mean_biomass, "g_fish"),
    sd_biomass = set_units(sd_biomass, "g_fish"),
    mean_per_biom = set_units(mean_per_biom, "g g_fish-1"),
    sd_per_biom = set_units(sd_per_biom, "g g_fish-1")
  ) %>% 
  select(-feed)

# Sum some measures - mean_per_biom
total_vals_1 <- total_vals %>% 
  filter(measure %in% c(
    "food_prov_stat", "total_excr_stat", "total_uneat_stat", 
    "C_excr_stat", "C_uneat_stat",
    "L_excr_stat", "L_uneat_stat",
    "P_excr_stat", "P_uneat_stat"
    )) %>% 
  mutate(measure = droplevels(measure)) %>% 
  select(farm_ID, measure, mean_per_biom, feed_group, feed_type) %>% 
  pivot_wider(
    names_from = measure,
    values_from = mean_per_biom
  ) %>% 
  mutate(
    total_waste_stat = total_excr_stat + total_uneat_stat,
    total_P_lost_stat = P_excr_stat + P_uneat_stat,
    total_C_lost_stat = C_excr_stat + C_uneat_stat,
    total_L_lost_stat = L_excr_stat + L_uneat_stat
  ) %>% 
  pivot_longer(
    cols = contains("stat"),
    names_to = "measure", names_transform = as.factor,
    values_to = "mean_per_biom"
  )

# Sum some measures - mean
total_vals_2 <- total_vals %>% 
  filter(measure %in% c(
    "food_prov_stat", "total_excr_stat", "total_uneat_stat", 
    "C_excr_stat", "C_uneat_stat",
    "L_excr_stat", "L_uneat_stat",
    "P_excr_stat", "P_uneat_stat"
    )) %>% 
  mutate(measure = droplevels(measure)) %>% 
  select(farm_ID, measure, mean, feed_group, feed_type) %>% 
  pivot_wider(
    names_from = measure, 
    values_from = mean
  ) %>% 
  mutate(
    total_waste_stat = total_excr_stat + total_uneat_stat,
    total_P_lost_stat = P_excr_stat + P_uneat_stat,
    total_C_lost_stat = C_excr_stat + C_uneat_stat,
    total_L_lost_stat = L_excr_stat + L_uneat_stat
  ) %>% 
  pivot_longer(
    cols = contains("stat"),
    names_to = "measure", names_transform = as.factor,
    values_to = "mean"
  )

# Rejoin mean and mean_per_biom
total_vals <- total_vals_1 %>% 
  left_join(total_vals_2, by = c("farm_ID", "feed_group", "feed_type", "measure"))

# Save
qsave(total_vals, file.path(data_analysis_path, "total_vals.qs"))

# Now do the same but as proportions rather than numbers
total_props <- total_vals %>% 
  pivot_wider(
    names_from = measure, names_sep = ".",
    values_from = c(mean, mean_per_biom)
  ) %>% 
  mutate(
    prop_excr_stat = mean.total_excr_stat/mean.total_waste_stat, # The proportion of total waste (feed+faeces) that comes from excreted faeces
    prop_uneat_stat = mean.total_uneat_stat/mean.food_prov_stat, # The proportion of feed provided that is lost to the environment
    prop_P_stat = mean.P_excr_stat/mean.total_P_lost_stat, # The proportion of total protein lost (feed+faeces) that is lost through excretion
    prop_C_stat = mean.C_excr_stat/mean.total_C_lost_stat, # The proportion of total carbohydrate lost (feed+faeces) that is lost through excretion
    prop_L_stat = mean.L_excr_stat/mean.total_L_lost_stat, # The proportion of total lipid lost (feed+faeces) that is lost through excretion
    prop_comp_excr_P_stat = mean.P_excr_stat/mean.total_excr_stat, # The composition of faeces, ie how much of faeces is protein
    prop_comp_excr_C_stat = mean.C_excr_stat/mean.total_excr_stat, # The composition of faeces, ie how much of faeces is carbohydrate
    prop_comp_excr_L_stat = mean.L_excr_stat/mean.total_excr_stat, # The composition of faeces, ie how much of faeces is lipid
  ) %>% 
  select(farm_ID, feed_group, feed_type, contains("prop"))
  
total_props <- total_props %>% 
  select(farm_ID, feed_group, feed_type, contains("prop")) %>% 
  pivot_longer(
    cols = contains("prop"), 
    names_to = c("measure"), names_transform = as.factor,
    values_to = "mean_prop", values_transform = drop_units
  )

qsave(total_props, file.path(data_analysis_path, "total_props.qs"))
```

To get nitrogen:
- N (g) = protein (g)/6.25

To get carbon:
- From protein: N (mol) = (1/3.7) * C (mol)
- From lipid: assume that lipid is 75% carbon
- From carbohydrate: assume that carb is around 40% carbon

```{r conversion factors}

CN_vals <- total_vals %>% 
  select(farm_ID, feed_group, measure, feed_type, mean, mean_per_biom) %>% 
  filter(measure %in% c("total_P_lost_stat", "total_C_lost_stat", "total_L_lost_stat", "C_excr_stat", "L_excr_stat", "P_excr_stat")) %>% 
  pivot_wider(names_from = "measure", values_from = c(mean, mean_per_biom), names_sep = ".") %>% 
  mutate(
    mean.N_excr = mean.P_excr_stat/6.25,
    mean.C_excr = (mean.L_excr_stat * 0.75) + 
      (mean.C_excr_stat * 0.4) + 
      ((3.7 * (mean.N_excr/14.007)) * 12.011),
    mean_per_biom.N_excr = mean_per_biom.P_excr_stat/6.25,
    mean_per_biom.C_excr = (mean_per_biom.L_excr_stat * 0.75) + 
      (mean_per_biom.C_excr_stat * 0.4) + 
      ((3.7 * (mean_per_biom.N_excr/14.007)) * 12.011),
    mean.N_total_lost = mean.total_P_lost_stat/6.25,
    mean.C_total_lost = (mean.total_L_lost_stat * 0.75) + 
      (mean.total_C_lost_stat * 0.4) + 
      ((3.7 * (mean.N_total_lost/14.007)) * 12.011),
    mean_per_biom.N_total_lost = mean_per_biom.total_P_lost_stat/6.25,
    mean_per_biom.C_total_lost = (mean_per_biom.total_L_lost_stat * 0.75) + 
      (mean_per_biom.total_C_lost_stat * 0.4) + 
      ((3.7 * (mean_per_biom.N_total_lost/14.007)) * 12.011)
  ) %>% 
  select(-contains("stat")) %>% 
  pivot_longer(
    cols = c(mean.N_excr, mean.C_excr, mean_per_biom.N_excr, mean_per_biom.C_excr, mean.N_total_lost, mean.C_total_lost, mean_per_biom.N_total_lost, mean_per_biom.C_total_lost),
    names_to = c(".value", "measure"),
    names_sep = "\\."
  ) %>% 
  mutate(
    nutrient = case_when(str_detect(measure, "C_") ~ "C", T ~ "N"),
    measure = str_remove(measure, "C_|N_") %>% as.factor(),
    nutrient = as.factor(nutrient)
  )

qsave(CN_vals, file.path(data_analysis_path, "total_CN_vals.qs"))
```
