---
title: "Run Atlantic salmon model"
format: html
editor: source
---

# Introduction

This document runs the aquaculture model for Atlantic salmon farms.

```{r setup, include=F}
#| warning: false
#| message: false

library(tidyverse)
library(arrow)
library(sf)
library(terra)
library(furrr)
library(future)
library(progressr)
library(tictoc)
library(fs)
library(conflicted)
library(readxl)
library(units)
library(qs)
library(here)
library(targets)
conflicted::conflicts_prefer(dplyr::filter(), dplyr::select(), .quiet = T)

here("src") %>% 
  list.files(pattern = "\\.R$", full.names = TRUE) %>% 
  str_subset("map", negate = T) %>% 
  walk(source)

# remove_unit("g_fish")
# remove_unit("kg_fish")
# remove_unit("t_fish")
# install_unit("g_fish")
# install_unit("kg_fish", "1000 g_fish")
# install_unit("t_fish", "1000 kg_fish")
```

# Run farms

The pipeline was having difficulty with the large files, so I've changed it to do the bulk of the processing here.

```{r set-up-decomposed-runs}
overwrite <- F
inds_per_farm <- 2500
farm_sample <- 25

# The decomposed runs are stored in one of three different potential places. The function uses this to check all possible filepaths before creating a new file.
potential_paths <- c(
  file.path(output_path, "decomposed_runs"),
  file.path(output_path, "decomposed_runs", "processed"),
  file.path(bigdata_path, "outputs", "decomposed_runs", "processed")
)

# Find out which farms are already done
already_done <- list.files(file.path(output_path, "farm_ends")) %>% 
  str_remove_all(".qs") %>% 
  str_split_i("_", 3) %>% 
  str_remove_all("ID") %>% 
  unique() %>% 
  as.numeric()

# Sample remaining farms
farm_IDs <- file.path(input_farm_coords_path, "farm_static_data_w_stocking.qs") %>% 
  qread() %>% 
  rename(farm_ID = farm_id) %>% 
  distinct(farm_ID) %>% 
  pull(farm_ID) %>% 
  setdiff(., already_done) %>% 
  sample(farm_sample)

farm_static_data <- file.path(input_farm_coords_path, "farm_static_data_w_stocking.qs") %>% 
  qread() %>% 
  sf::st_drop_geometry() %>% 
  rename(farm_ID = farm_id) %>% 
  filter(farm_ID %in% farm_IDs) %>% 
  group_by(farm_ID) %>% 
  group_split()

farm_temporal_data <- file.path(input_farm_coords_path, "farm_temporal_data.qs") %>% 
  qread() %>% 
  rename(farm_ID = farm_id) %>% 
  filter(farm_ID %in% farm_IDs) %>% 
  arrange(farm_ID, day) %>% 
  group_by(farm_ID) %>% 
  group_split()

# Salmon params
all_params <- c(
  file.path(output_species_data_path, "species_params.qs") %>% qread(),
  file.path(output_species_data_path, "pop_params.qs") %>% qread()
)

# Add population from stocking n
farm_temporal_data <- purrr::map2(
  farm_static_data, 
  farm_temporal_data, 
  function(static, ts) {
    cbind(
      ts %>% 
        filter(day >= static$t_start & day <= static$t_end),
      data.frame(Npop = generate_pop(
          harvest_n = static$harvest_n,
          mort = all_params['mortmyt'],
          times = c(t_start = static$t_start, t_end = static$t_end, dt = 1)
      ))
    )
  }
)

# Feed params
feed_params <- c(
  file.path(output_species_data_path, "feed_params_MD.qs") %>% qread(), 
  file.path(output_species_data_path, "feed_params_PD.qs") %>% qread(), 
  file.path(output_species_data_path, "feed_params_AD.qs") %>% qread()
)
```

```{r bulk-decomposed-runs}
plan(multisession, workers = 5)
handlers(global = TRUE)  # auto-enables progress reporting
handlers("cli")   

# For testing
# ts <- ts_data[[1]]
# static <- static_data[[1]]

# Run!
with_progress({
  p <- progressor(steps = length(feed_params)*length(farm_IDs))

  for (i in seq_along(feed_params)) {
    feed <- names(feed_params)[i]
    if (!str_detect(feed, "_min|_max")) {
      feed <- paste(feed, "mean", sep = "_")
    }

    cat("Starting", feed, "now...")

    future_map2(
      farm_static_data,
      farm_temporal_data,
      function(static, ts) {
        p()
        ID <- static$farm_ID
        filename <- file.path(output_path, "decomposed_runs", str_c(feed, "_ID", fixnum(ID), "_", inds_per_farm, "_inds.qs"))

        if (all(!file.exists(file.path(potential_paths, basename(filename)))) | overwrite) {
          fg <- farm_growth_decomposed(
              pop_params = all_params,
              species_params = all_params,
              water_temp = ts$temp_c,
              feed_params = feed_params[[i]],
              times = c(t_start = static$t_start, t_end = static$t_end, dt = 1),
              N_pop = ts$Npop,
              nruns = inds_per_farm
            ) %>% 
            mutate(
              farm_ID = as.integer(ID),
              feed = as.factor(feed)
            )
          qsave(fg, filename)
        }
      },
      .options = furrr_options(seed = 123)
    )
  }
})

plan(sequential)
gc()
```

## Processing

```{r determine-new-files}
last_run_file <- file.path(output_path, "processing_timestamp.qs")
last_run <- if (file.exists(last_run_file)) qs::qread(last_run_file) else as.POSIXct(0, origin = "2026-02-25")

decom_files <- data.frame(
  filename = list.files(file.path(output_path, "decomposed_runs"), full.names = TRUE, pattern = ".qs")
  ) %>%
  mutate(
    ID = str_split_i(basename(filename), "_", 5),
    last_modified = file.info(filename)$mtime
  ) %>%
  group_by(ID) %>%
  nest()

modified_files <- decom_files %>%
  filter(
    map_lgl(data, ~ any(.x$last_modified > last_run)),
    map_lgl(data, ~ nrow(.x) == 9)
  )
```

```{r processing-functions}
get_fish_ends <- function(df) {
  fe_weight <- df %>% 
    filter(measure == "weight" & prod_t == max(prod_t)) %>% 
    mutate(value = value * N_pop) %>% 
    select(-measure, -N_pop, -t, -prod_t) %>% 
    rename(weight = value)

  fe_outcomes <- df %>%
    filter(prod_t != max(prod_t) & measure %in% c("total_excr", "total_uneat", "Nitrogen_excr", "Nitrogen_uneat", "Carbon_excr", "Carbon_uneat", "total_Carbon", "total_Nitrogen")) %>% 
    mutate(value = value * N_pop) %>% 
    group_by(farm_ID, feed, fish, measure) %>% 
    reframe(value = sum(value))

  full_join(fe_weight, fe_outcomes, by = join_by(farm_ID, feed, fish)) %>% 
    mutate(measure = droplevels(measure)) %>% 
    mutate(value_perbiom = value/weight) %>% 
    select(-weight)
}

get_farm_run <- function(df) {
  fr_weight <- df %>% 
    filter(measure == "weight" & prod_t != max(prod_t)) %>% 
    mutate(value = value * N_pop) %>% 
    select(-measure, -N_pop) %>% 
    rename(weight = value)

  fr_outcomes <- df %>%
    filter(prod_t != max(prod_t) & measure %in% c("weight", "total_excr", "total_uneat", "Nitrogen_excr", "Nitrogen_uneat", "Carbon_excr", "Carbon_uneat", "total_Carbon", "total_Nitrogen")) %>% 
    mutate(value = value * N_pop) %>% 
    group_by(farm_ID, feed, fish, measure, t, prod_t) %>% 
    reframe(value = sum(value))

  full_join(fr_weight, fr_outcomes, by = join_by(farm_ID, feed, fish, t, prod_t)) %>% 
    mutate(measure = droplevels(measure)) %>% 
    mutate(value_perbiom = value/weight) %>% 
    select(-weight) %>% 
    group_by(farm_ID, feed, prod_t, t, measure) %>% 
    reframe(
      mean_value = mean(value),
      sd_value = sd(value),
      mean_value_perbiom = mean(value_perbiom),
      sd_value_perbiom = sd(value_perbiom)
    )
}

get_farm_ends <- function(df) {
  fa_weight <- df %>% 
    filter(measure == "weight" & prod_t == max(prod_t)) %>% 
    mutate(value = value * N_pop) %>% 
    select(-measure, -N_pop, -t, -prod_t) %>% 
    rename(weight = value)

  fa_outcomes <- df %>%
    filter(measure %in% c("total_excr", "total_uneat", "Nitrogen_excr", "Nitrogen_uneat", "Carbon_excr", "Carbon_uneat", "total_Carbon", "total_Nitrogen")) %>% 
    mutate(value = value * N_pop) %>% 
    group_by(farm_ID, feed, fish, measure) %>% 
    reframe(value = sum(value))

  full_join(fa_weight, fa_outcomes, by = join_by(farm_ID, feed, fish)) %>% 
    mutate(measure = droplevels(measure)) %>% 
    mutate(value_perbiom = value/weight) %>% 
    select(-weight) %>% 
    group_by(farm_ID, feed, measure) %>% 
    reframe(
      mean_value = mean(value),
      sd_value = sd(value),
      mean_value_perbiom = mean(value_perbiom),
      sd_value_perbiom = sd(value_perbiom)
    )
}
```

```{r save-intermedia-outputs}
# For testing
# i <- 1
# fnms <- modified_files$data[[i]]
# ID <- modified_files$ID[i]

plan(multisession, workers = 2)
overwrite <- F

walk2(
  modified_files$data,
  modified_files$ID,
  function(fnms, ID) {

    fish_ends_file <- file.path(output_path, "fish_ends", str_c("fish_ends_", ID, ".qs"))
    farm_run_file <- file.path(output_path, "farm_run", str_c("farm_run_", ID, ".qs"))
    farm_ends_file <- file.path(output_path, "farm_ends", str_c("farm_ends_", ID, ".qs"))

    if (any(!file.exists(c(fish_ends_file, farm_run_file, farm_ends_file))) | overwrite) {
      # Read in the raw files
      fish_runs <- future_map(fnms$filename, function(fnm) {qread(fnm)}, .options = furrr_options(seed = 123))

      if (!file.exists(fish_ends_file) | overwrite) {
        # Keep fish seperate but take only the ends (sums) of measures, for statistics
        fish_ends <- future_map_dfr(fish_runs, get_fish_ends, .options = furrr_options(seed = 123))
        qsave(fish_ends, fish_ends_file)
        rm(fish_ends)
      }

      if (!file.exists(farm_run_file) | overwrite) {
        # Combine all fish into farm mean and sd, per t
        farm_run <- future_map_dfr(fish_runs, get_farm_run, .options = furrr_options(seed = 123))
        qsave(farm_run, farm_run_file)
        rm(farm_run)
      }

      if (!file.exists(farm_ends_file) | overwrite) {
        # Combine all fish into farm mean and sd
        farm_ends <- future_map_dfr(fish_runs, get_farm_ends, .options = furrr_options(seed = 123))
        qsave(farm_ends, farm_ends_file)
        rm(farm_ends)
      }

      rm(fish_runs)
      gc()

    }
  },
  .progress = T
)
```

```{r update-progress-and-move-files}
qs::qsave(Sys.time(), last_run_file)

# Move all files processed into a new folder
modified_files %>%
  unnest(data) %>%
  mutate(
    new_path = file.path(output_path, "decomposed_runs", "processed", basename(filename)),
    file.rename(filename, new_path)
  )

gc()
```

Some random plots for the supplementaries.

```{r random-plots}
farm_run_files <- file.path(output_path, "farm_run") %>% 
  list.files(full.names = T) %>% 
  sample(5)

dfs <- map_dfr(farm_run_files, function(fnm) {
  qread(fnm)
})

ggplot(dfs, aes(x = prod_t, y = mean_value, colour = as.factor(feed))) +
  geom_line() +
  theme_classic() +
  facet_grid(rows = vars(farm_ID), cols = vars(measure), scales = "free")
```

# Pipeline

The following code runs the main pipeline.

```{r run-farms-pipeline, eval=F}
#| code-summary: Run the main pipeline

Sys.setenv(TAR_PROJECT = "project_farmruns")

# tar_validate()

# Run whole pipeline
tar_make(lm_tests)
```

Load objects from the targets pipeline and save them as .qs files, which are much quicker to load down the line. This takes a few minutes to gather all the targets.

To get nitrogen:
- N (g) = protein (g)/6.25

To get carbon:
- From protein: N (mol) = (1/3.7) * C (mol)
- From lipid: lipid is 75% carbon
- From carbohydrate: carbohydrate is around 40% carbon

These are from Prasad, S. S., Cameron, D. D., Carter, C. G., Williamson, A., & Blanchard, J. L. (2025). Geometric stoichiometry model yields relevant insights for assessing nutrient-related environmental impacts of aquaculture. Conservation Physiology, 13(1). https://doi.org/10.1093/conphys/coaf066

```{r all-farms-results}
#| code-summary: Save pipeline objects as files
#| eval: false

Sys.setenv(TAR_PROJECT = "project_farmruns")

# These are the raw results - they're huge and take a couple minutes to load/save
tar_load(farm_run_MD_chunked)
qsave(farm_run_MD_chunked, file.path(output_path, "farm_run_MD_chunked.qs"))
tar_load(farm_run_AD_chunked)
qsave(farm_run_AD_chunked, file.path(output_path, "farm_run_AD_chunked.qs"))
tar_load(farm_run_PD_chunked)
qsave(farm_run_PD_chunked, file.path(output_path, "farm_run_PD_chunked.qs"))

tar_load(farm_results)
qsave(farm_results, file.path(data_analysis_path, "total_vals.qs"))

tar_load(total_props)
qsave(total_props, file.path(data_analysis_path, "total_props.qs"))

tar_load(CN_vals)
qsave(CN_vals, file.path(data_analysis_path, "total_CN_vals.qs"))
```

# Run sensitivities

```{r check-sensitivities-pipeline}
#| code-summary: Check that the pipeline will run correctly

Sys.setenv(TAR_PROJECT = "project_sensitivities")

tar_make(test_reference_feed)
tar_read(test_reference_feed) # if weight < 1000g then feeds aren't processing properly! Fix it!
```

```{r}
Sys.setenv(TAR_PROJECT = "project_sensitivities")

# Run whole pipeline
tar_make(seconds_meta_append = 300)
```

```{r sens-results}
tar_load(sens_results_pop)
tar_load(sens_results_spec)

sens <- rbind(sens_results_pop, sens_results_spec)
qsave(sens, file.path(output_sens_data_path, "sens_all.qs"))
```
